### 2016-09-03 paper.tex
made some changes

\documentclass[twoside]{article}
\usepackage{aistats2015}

% If your paper is accepted, change the options for the package
% aistats2015 as follows:
%
%\usepackage[accepted]{aistats2015}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{xcolor}
%\usepackage[authordate,bibencoding=auto,strict,backend=biber,natbib]{biblatex-chicago}
\usepackage[round]{natbib}   % omit 'round' option if you prefer square brackets
\bibliographystyle{plainnat}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand{\Q}{\mathcal{Q}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\w}{\theta}
\newcommand{\wkl}{\hat\w^{kl}}
\newcommand{\wave}{\hat\w^{ave}}
\newcommand{\waver}{\hat\w^{ave,r}}
\newcommand{\wboot}{\hat\w^{boot}}
\newcommand{\wmle}{\hat\w^{mle}}
\newcommand{\wmler}{\hat\w^{mle,r}}
\newcommand{\wstar}{{\w^{*}}}
\newcommand{\wq}{\hat\w^{q}}
\newcommand{\wqstar}{\hat\w^{q^*}}
\newcommand{\dist}{\mathcal{D}}

\newcommand{\normal}[2]{\ensuremath{\mathcal{N}\left({{#1}},{{#2}}\right)}}
\newcommand{\trans}[1]{\ensuremath{{#1}^{\mathsf{T}}}}
\newcommand{\ltwo}[1]{{\lVert {#1} \rVert}_2}
\newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
\newcommand{\lzero}[1]{{\lVert {#1} \rVert}_0}

\DeclarePairedDelimiterX{\infdivx}[2]{(}{)}{%
  #1\;\delimsize\|\;#2%
}
\newcommand{\kl}{\text{KL}\infdivx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{filecontents}{paper.bib}

@inproceedings{merugu2003privacy,
  title={Privacy-preserving distributed clustering using generative models},
  author={Merugu, Srujana and Ghosh, Joydeep},
  booktitle={Data Mining, 2003. ICDM 2003. Third IEEE International Conference on},
  pages={211--218},
  year={2003},
  organization={IEEE}
}

@inproceedings{mcdonald2009efficient,
  title={Efficient large-scale distributed training of conditional maximum entropy models},
  author={McDonald, Ryan and Mohri, Mehryar and Silberman, Nathan and Walker, Dan and Mann, Gideon S},
  booktitle={Advances in Neural Information Processing Systems},
  pages={1231--1239},
  year={2009}
}

@inproceedings{zinkevich2010parallelized,
  title={Parallelized stochastic gradient descent},
  author={Zinkevich, Martin and Weimer, Markus and Li, Lihong and Smola, Alex J},
  booktitle={Advances in neural information processing systems},
  pages={2595--2603},
  year={2010}
}

@inproceedings{zhang2012communication,
  title={Communication-efficient algorithms for statistical optimization},
  author={Zhang, Yuchen and Wainwright, Martin J and Duchi, John C},
  booktitle={Advances in Neural Information Processing Systems},
  pages={1502--1510},
  year={2012}
}

@article{liu2012distributed,
  title={Distributed parameter estimation via pseudo-likelihood},
  author={Liu, Qiang and Ihler, Alexander},
  journal={arXiv preprint arXiv:1206.6420},
  year={2012}
}

@inproceedings{zhang2013information,
  title={Information-theoretic lower bounds for distributed statistical estimation with communication constraints},
  author={Zhang, Yuchen and Duchi, John and Jordan, Michael I and Wainwright, Martin J},
  booktitle={Advances in Neural Information Processing Systems},
  pages={2328--2336},
  year={2013}
}

@inproceedings{zhang2013divide,
  title={Divide and Conquer Kernel Ridge Regression.},
  author={Zhang, Yuchen and Duchi, John C and Wainwright, Martin J},
  booktitle={COLT},
  year={2013}
}

@inproceedings{zhang2012communication,
  title={Communication-efficient algorithms for statistical optimization},
  author={Zhang, Yuchen and Wainwright, Martin J and Duchi, John C},
  booktitle={Advances in Neural Information Processing Systems},
  pages={1502--1510},
  year={2012}
}

@article{vw,
  author={Langford, John},
  title={Vowpal Wabbit open source project},
  %url={https://github.com/JohnLangford/vowpal_wabbit}
}

@article{agarwal2014reliable,
  title={A reliable effective terascale linear learning system.},
  author={Agarwal, Alekh and Chapelle, Olivier and Langford, John},
  journal={Journal of Machine Learning Research},
  volume={15},
  number={1},
  pages={1111--1133},
  year={2014}
}

@inproceedings{liu2014distributed,
  title={Distributed estimation, information loss and exponential families},
  author={Liu, Qiang and Ihler, Alexander T},
  booktitle={Advances in Neural Information Processing Systems},
  pages={1098--1106},
  year={2014}
}

@inproceedings{shamir2014fundamental,
  title = {Fundamental Limits of Online and Distributed Algorithms for Statistical Learning and Estimation},
  author = {Shamir, Ohad},
  booktitle = {Advances in Neural Information Processing Systems 27},
  pages = {163--171},
  year = {2014}
}

@article{duchi2014optimality,
  title={Optimality guarantees for distributed statistical estimation},
  author={Duchi, John C and Jordan, Michael I and Wainwright, Martin J and Zhang, Yuchen},
  journal={arXiv preprint arXiv:1405.0782},
  year={2014}
}

@inproceedings{garg2014communication,
  title={On communication cost of distributed statistical estimation and dimensionality},
  author={Garg, Ankit and Ma, Tengyu and Nguyen, Huy},
  booktitle={Advances in Neural Information Processing Systems},
  pages={2726--2734},
  year={2014}
}

@article{braverman2015communication,
  title={Communication lower bounds for statistical estimation problems via a distributed data processing inequality},
  author={Braverman, Mark and Garg, Ankit and Ma, Tengyu and Nguyen, Huy L and Woodruff, David P},
  journal={arXiv preprint arXiv:1506.07216},
  year={2015}
}

@article{han2016bootstrap,
  title={Bootstrap Model Aggregation for Distributed Statistical Learning},
  author={Han, Jun and Liu, Qiang},
  journal={arXiv preprint arXiv:1607.01036},
  year={2016}
}
\end{filecontents}
\immediate\write18{bibtex paper}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

\twocolumn[

\aistatstitle{Instructions for paper submissions to AISTATS 2015}

\aistatsauthor{ Anonymous Author 1 \And Anonymous Author 2 \And Anonymous Author 3 }

\aistatsaddress{ Unknown Institution 1 \And Unknown Institution 2 \And Unknown Institution 3 } ]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% one-shot distributed learning
% communication efficient
% non-interactive

%Used by Vowpal Wabbit \cite{vw,agarwal2014reliable}.

\section{INTRODUCTION}
We consider the problem of distributed Maximum Likelihood Estimation (MLE) with only a single round of communication.

%In a Generalized Linear Model,
%\begin{equation}
%\begin{aligned}
%y &= g^{-1}(\trans\x\w) + \epsilon
%,~~~
%\epsilon \sim \normal{0}{\sigma}
%\end{aligned}
%\end{equation}

\section{PROBLEM STATEMENT}

%Let $\dist$ be a distribution over $\mathbb{R}^d$,
%and $X$ a dataset of $N$ points sampled from $\dist$.
%The maximum likelihood estimator for the GLM is
%\begin{equation}
%\wstar=\argmax_\w \int_\dist f(\trans\x\w) + g(\w) dx
%\end{equation}
Let $\X$ be a probability space
and $\Theta$ be the space of model parameters.
Let $f : \X \times \Theta \to \mathbb{R}$ be the likelihood of the data given the model parameters.
Then the true model parameters are given by
\begin{equation}
\wstar = \argmax_\w \int_\X f(\x;\w) d\x
\end{equation}
Since the distribution of $\X$ is typically unknown, $\wstar$ cannot be calculated directly.
Instead, given a sample $X\in\X$, we calculate the MLE by
\begin{equation}
\wmle=\argmax_\w \sum_{\x\in X} f(\x;\w)
\end{equation}
When the dataset $X$ is too large to fit on a single machine,
standard optimization methods cannot calculate $\wmle$ directly.
A number of distributed optimization methods have been created,
but these typically require one round of communication for each iteration of the optimization.
This communication is the main bottleneck of the optimization.

Our goal is to approximate $\wstar$ with only a single round of communication.
Split the dataset $X$ into $m$ disjoint datasets $X_1,...,X_m$ each of size $n=\frac{N}{m}$.
On each machine, solve the \emph{local} maximum likelihood estimate
\begin{equation}
\wmle_i = \argmax_\w \sum_{\x\in X_i} f(\x;\w)
\end{equation}
Solving for $\wmle_i$ requires no communication with other machines.
Next, we need a function $q : \Theta^m \to \Theta$ that combines the parameter estimates from each machine.
Define the resulting parameter estimate as
\begin{equation}
\wq = q(\wmle_1, \wmle_2, ..., \wmle_m)
\end{equation}
Let $\Q$ be the set of all functions from $\Theta^m\to\Theta$,
and $Q$ be a subset of $\Q$.
Our goal is to find the function
\begin{equation}
q^* = \argmin_{q\in Q} \ltwo{\wqstar - \wstar}
\end{equation}

\section{RELATED WORK}

Related work can be divided into two categories:
alternative $q$ estimators,
and bounds on the communication complexity of distributed learning.

\subsection{Alternative estimators}
The simplest and most popular communication efficient estimator is the averaging estimator
\begin{equation}
\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
\end{equation}
\cite{mcdonald2009efficient}
 gave the first analysis of this estimator wherein they showed that the variance of the estimator reduces as $O(m^{-1/2})$, but that the bias of $\wave$ is the same as $\wmle_i$.
\cite{zhang2012communication} showed that the Mean Squared Error (MSE) $\ltwo{\wave-\wstar}^2$ decays as $O((nm)^{-1} + n^{-2})$.
This is slower than the optimal sequential algorithm which has a MSE of $O((mn)^{-2})$.
Surprisingly, \cite{zhang2013divide} showed that in the special case of kernel ridge regression, a reduction in bias was not needed to have the MSE decay at the optimal sequential rate.
By a careful choice of regularization parameter, they caused $\wmle_i$ to have lower bias but higher variance, so that the final estimate of $\wave$ had both reduced bias and variance.

Other research has focused on modifications to the $\wave$ estimator to reduce bias.
\cite{zinkevich2010parallelized} showed that if the training sets partially overlap each other (instead of being disjoint), then the resulting estimator will have lower bias.
\cite{zhang2012communication} also provided a bootstrap average estimator,
which works as follows.
Let $r\in(0,1)$, and $X_i^r$ be a bootstrap sample of $X_i$ with $rn$ data points.
Then the bootstrap average estimator is
\begin{equation}
\begin{aligned}
\wmler_i &= \argmax_\w \sum_{\x\in X_i^r} f(x;\w)
\\
\waver &= \frac{1}{m}\sum_{i=1}^m \wmler_i
\\
\wboot & = \frac{\wave-r\waver}{1-r}
\end{aligned}
\end{equation}
This estimator enjoys a MSE decaying as $O((nm)^{-1}+n^{-3})$.

%An alternative definition of the $\wave$ estimator is
%\begin{equation}
%\wave = \argmin_\w \frac{1}{m}\sum_{i=1}^m \ltwo{\wmle_i-\w}^2
%\end{equation}
%It is easy to show that the two definitions are equivalent with standard calculus.

\cite{liu2014distributed} propose a more Bayesian approach inspired by \cite{merugu2003privacy}.
Instead of averaging the model's parameters,
they directly ``average the models'' with the following KL-average estimator.
\begin{equation}
\wkl = \argmin_\w \sum_{i=1}^m \kl[\bigg]{p(\cdot;\wmle_i)}{p(\cdot;\w)}
\end{equation}
The minimization is performed via a bootstrap sample from the smaller models.
This method has two main advantages.
First, it is robust to reparameterizations of the model.
Second, it is statistically optimal for the class of non-interactive optimization methods.
(We show in the next section that their bound does not apply to our procedure due to our semi-interactive setting.)
The main downside of the KL-average is that the minimization has a prohibitively high computational cost.
Let $n^{kl}$ be the size of the bootstrap sample.
Then the original implementation's MSE shrinks as $O((nm)^{-1}+(nn^{kl})^{-1})$.
This implies that the bootstrap procedure requires as many samples as the original problem to get a MSE that shrinks at the same rate as the averaging estimator.
\cite{han2016bootstrap} show a method to reduce this rate to $O((nm)^{-1}+(n^2n^{kl})^{-1})$ using control variates, but the procedure remains prohibitively expensive.
Their experiments show the procedure scaling only to datasets of size $nm\approx10^4$,
whereas our experiments involve a dataset of size $nm\approx10^8$

\subsection{Performance bounds}

Performance bounds come in two flavors: statistical and information theoretic.
On the statistical side, \cite{liu2014distributed} showed that for an arbitrary function $q$ \emph{which does not depend on the data},
$\ltwo{\wq-\wmle}^2$ decays as $\Omega(\gamma^2_\wstar I^{-1}_\wstar/n^2)$.
Here $\gamma_\wstar$ is the statistical curvature of the model and $I_\wstar$ is the Fisher information.
Furthermore, they show that $\wkl$ matches this bound.
This bound is not relevant for our estimator because our merge function depends on the data.
Furthermore, we directly optimize $\ltwo{\wq-\wstar}^2$ which is a more useful quantity in practice.

\cite{shamir2014fundamental}, \cite{zhang2013information}, and \cite{garg2014communication} all provide information theoretic lower bounds on the sample complexity of non-interactive learning problems.
As above, however, there results are not applicable in our semi-interactive setting.
%provides general information theoretic bounds for what he terms learning with information constraints.
%As a special case of Shamir's analysis is the non-interactive communication model where $q$ does not depend on the data.
%When $q$ is allowed to depend on the data, his model no longer applies.
There is one information theoretic lower bound that does apply to us.
Let the true parameter vector $\wstar$ be $k$-sparse.
That is, $\lzero{\wstar} \le k$.
%Then \cite{zhang2013information} showed that at least $\Omega(mk/\log m)$ bits of communication are required in the non-interactive setting,
%and \cite{garg2014communication} improved this lower bound to $\Omega(mk)$.
Surprisingly, \cite{braverman2015communication} showed that the minimax optimal error rate for least squares regression requires $\Omega(m\cdot\min\{n,d\})$ bits of communication even in the fully interactive setting.
This is important because sparsity does not reduce the amount of communication required, and this bound does apply in our setting.

%\section{OUR ALGORITHM}
%
%\section{ANALYSIS}
%
%\section{EXPERIMENTS}
%
%\section{CONCLUSION}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{plainnat}
\bibliography{paper}

\end{document}

### 2016-09-08 paper.tex
lots of changes

@@
 \usepackage{amssymb}
+\usepackage{amsthm}
 \usepackage{mathtools}
@@
 \usepackage{xcolor}
+\usepackage[inline]{enumitem}
 %\usepackage[authordate,bibencoding=auto,strict,backend=biber,natbib]{biblatex-chicago}
@@
 
+\newtheorem{cor}{Corallary}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+
 \DeclareMathOperator*{\argmin}{arg\,min}
 \DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\vecspan}{span}
+\DeclareMathOperator*{\affspan}{aff}
+\DeclareMathOperator*{\tr}{tr}
 
+\newcommand{\zero}{\text{\textbf{0}}}
 \newcommand{\Q}{\mathcal{Q}}
+\newcommand{\Y}{\mathcal{Y}}
 \newcommand{\X}{\mathcal{X}}
+\newcommand{\W}{\mathcal{\hat W}}
+\newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
+\newcommand{\Wtave}{{\mathcal{W}^{ave,*}}}
+\newcommand{\V}{\mathcal{V}}
+\newcommand{\A}{\mathcal{A}}
+\newcommand{\D}{\mathcal{D}}
+\newcommand{\E}{\mathbb{E}}
 \newcommand{\x}{\mathbf{x}}
@@
 \newcommand{\wkl}{\hat\w^{kl}}
+\newcommand{\wreopt}{\hat\w^{reopt}}
 \newcommand{\wave}{\hat\w^{ave}}
+\newcommand{\wtave}{\hat\w^{ave,*}}
 \newcommand{\waver}{\hat\w^{ave,r}}
@@
 
+\newcommand{\tbias}{t_{\text{\textit{bias}}}}
+\newcommand{\tvar}{t_{\text{\textit{var}}}}
+
+\newcommand{\I}{I}
+\newcommand{\Iinv}{I^{-1}}
+\newcommand{\law}{\ensuremath{\xrightarrow{L}}}
 \newcommand{\normal}[2]{\ensuremath{\mathcal{N}\left({{#1}},{{#2}}\right)}}
 \newcommand{\trans}[1]{\ensuremath{{#1}^{\mathsf{T}}}}
-\newcommand{\ltwo}[1]{{\lVert {#1} \rVert}_2}
+\newcommand{\ltwo}[1]{{\left\lVert {#1} \right\rVert}_2}
 \newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
 \newcommand{\lzero}[1]{{\lVert {#1} \rVert}_0}
+\newcommand{\proj}[1]{\pi_{{#1}}}
+\newcommand{\prob}[1]{\Pr\left[{#1}\right]}
 
@@
 
+\newcommand{\ignore}[1]{}
+\newcommand{\fixme}[1]{\textbf{FIXME:} {#1}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{filecontents}{paper.bib}
+
+@book{lehmann1999elements,
+  title={Elements of large-sample theory},
+  author={Lehmann, Erich Leo},
+  year={1999},
+  publisher={Springer Science \& Business Media}
+}
 
@@
   organization={IEEE}
+}
+
+@article{dasgupta2003elementary,
+  title={An elementary proof of a theorem of {J}ohnson and {L}indenstrauss},
+  author={Dasgupta, Sanjoy and Gupta, Anupam},
+  journal={Random Structures \& Algorithms},
+  volume={22},
+  number={1},
+  pages={60--65},
+  year={2003},
+  publisher={Wiley Online Library}
+}
+
+@article{matouvsek2008variants,
+  title={On variants of the {J}ohnson--{L}indenstrauss lemma},
+  author={Matou{\v{s}}ek, Ji{\v{r}}{\'\i}},
+  journal={Random Structures \& Algorithms},
+  volume={33},
+  number={2},
+  pages={142--156},
+  year={2008},
+  publisher={Wiley Online Library}
 }
@@
   year={2010}
+}
+
+@article{spokoiny2012parametric,
+  title={Parametric estimation. Finite sample theory},
+  author={Spokoiny, Vladimir},
+  journal={The Annals of Statistics},
+  volume={40},
+  number={6},
+  pages={2877--2909},
+  year={2012},
+  publisher={Institute of Mathematical Statistics}
 }
@@
   pages={1502--1510},
+  year={2012}
+}
+
+@article{hsu2012tail,
+  title={A tail inequality for quadratic forms of subgaussian random vectors},
+  author={Hsu, Daniel and Kakade, Sham M and Zhang, Tong and others},
+  journal={Electron. Commun. Probab},
+  volume={17},
+  number={52},
+  pages={1--6},
   year={2012}
@@
   year={2014}
+}
+
+@book{shalev2014understanding,
+  title={Understanding machine learning: From theory to algorithms},
+  author={Shalev-Shwartz, Shai and Ben-David, Shai},
+  year={2014},
+  publisher={Cambridge University Press}
 }
@@
 \section{INTRODUCTION}
-We consider the problem of distributed Maximum Likelihood Estimation (MLE) with only a single round of communication.
+We consider the problem of distributed Maximum Likelihood Estimation (MLE).
 
@@
 
-\section{PROBLEM STATEMENT}
+\subsection{Problem Statement}
 
-%Let $\dist$ be a distribution over $\mathbb{R}^d$,
-%and $X$ a dataset of $N$ points sampled from $\dist$.
-%The maximum likelihood estimator for the GLM is
+%Let $\Theta\subseteq\mathbb{R}^d$ be the parameter space,
+%$\X\subseteq\mathbb{R}^d$ be the covariate space,
+%and $\Y\subseteq\mathbb{R}$ be the response space.
+%Let $Z\subset\X\times\Y$ be a set of $mn$ i.i.d. datapoints.
+%Then the centralized Maximum Likelihood Estimator (MLE) is defined as
 %\begin{equation}
-%\wstar=\argmax_\w \int_\dist f(\trans\x\w) + g(\w) dx
+%\wmle = \argmax_\w \sum_{(\x,y)\in Z} f(y-\trans\x\w)
 %\end{equation}
-Let $\X$ be a probability space
-and $\Theta$ be the space of model parameters.
-Let $f : \X \times \Theta \to \mathbb{R}$ be the likelihood of the data given the model parameters.
-Then the true model parameters are given by
+
+Let $\Y\subseteq\mathbb{R}$ be the space of response variables,
+$X\subseteq\mathbb{R}^d$ be the space of the covariates,
+and $\Theta\subseteq\mathbb{R}^d$ be the parameter space.
+We assume a linear model.
+The log-likelihood of data point $(\x,y)\in\X\times\Y$ given the model's true parameter $\wstar\in\Theta$ is denoted by $f(y,\trans\x\wstar)$.
+%\begin{equation}
+%y = g^{-1}(\trans\x\wstar)
+%\end{equation}
+%Here, $g : \mathbb{R} \to \mathbb{R}$ is the link function and $\wstar\in\Theta\subseteq\mathbb{R}^d$ is the true parameter of our model.
+Let $Z\subset\X\times\Y$ be a dataset of $mn$ i.i.d. observations.
+Then the maximum likelihood estimator is
 \begin{equation}
-\wstar = \argmax_\w \int_\X f(\x;\w) d\x
+%\wmle=\argmax_\w \sum_{i=1}^{nm} g(y_i-\trans\x_i\w)
+\wmle=\argmax_\w \sum_{(\x,y)\in Z} f(y,\trans\x\w)
 \end{equation}
-Since the distribution of $\X$ is typically unknown, $\wstar$ cannot be calculated directly.
-Instead, given a sample $X\in\X$, we calculate the MLE by
+Assume that $Z$ has been split up onto $m$ machines so that each machine $i$ has dataset $Z_i$ of size $n$, and all the $Z_i$ are disjoint.
+Then each machine calculates the local estimator
 \begin{equation}
-\wmle=\argmax_\w \sum_{\x\in X} f(\x;\w)
-\end{equation}
-When the dataset $X$ is too large to fit on a single machine,
-standard optimization methods cannot calculate $\wmle$ directly.
-A number of distributed optimization methods have been created,
-but these typically require one round of communication for each iteration of the optimization.
-This communication is the main bottleneck of the optimization.
-
-Our goal is to approximate $\wstar$ with only a single round of communication.
-Split the dataset $X$ into $m$ disjoint datasets $X_1,...,X_m$ each of size $n=\frac{N}{m}$.
-On each machine, solve the \emph{local} maximum likelihood estimate
-\begin{equation}
-\wmle_i = \argmax_\w \sum_{\x\in X_i} f(\x;\w)
+\wmle_i = \argmax_\w \sum_{(\x,y) \in Z_i} f(y,\trans\x\w)
 \end{equation}
 Solving for $\wmle_i$ requires no communication with other machines.
-Next, we need a function $q : \Theta^m \to \Theta$ that combines the parameter estimates from each machine.
-Define the resulting parameter estimate as
+Our goal is to provide a function $q : \Theta^m \to \Theta$ that ``merges'' these estimates.
+The resulting merged estimate should have error as close to $\wmle$ as possible.
+
+%Next, we need a function $q : \Theta^m \to \Theta$ that combines the parameter estimates from each machine.
+%Define the resulting parameter estimate as
+%\begin{equation}
+%\wq = q(\wmle_1, \wmle_2, ..., \wmle_m)
+%\end{equation}
+%Let $\Q$ be the set of all functions from $\Theta^m\to\Theta$,
+%and $Q$ be a subset of $\Q$.
+%Our goal is to find the function
+%\begin{equation}
+%q^* = \argmin_{q\in Q} \ltwo{\wqstar - \wstar}
+%\end{equation}
+
+\subsection{Proposed Solution}
+
+The averaging estimator is defined as
 \begin{equation}
-\wq = q(\wmle_1, \wmle_2, ..., \wmle_m)
+\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
 \end{equation}
-Let $\Q$ be the set of all functions from $\Theta^m\to\Theta$,
-and $Q$ be a subset of $\Q$.
-Our goal is to find the function
+Define the matrix $\hat W : \mathbb{R}^{d\times m}$ to have $i$th column equal $\wmle_i-\wave$.
+Then our estimator is defined as
 \begin{equation}
-q^* = \argmin_{q\in Q} \ltwo{\wqstar - \wstar}
+\begin{aligned}
+\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x (\hat W\alpha + \wave)\right)
+\\
+\wreopt &= \hat W \hat\alpha + \wave
+\end{aligned}
 \end{equation}
+where $Z^{reopt}$ is a small dataset of size $O(m)$.
+In the analysis of this estimator, we assume that $Z^{reopt}$ is independent of the $\wmle_i$;
+that is, the data points in $Z^{reopt}$ were not used in the single machine optimizations.
+In practice, however, this restriction is not necessary.
+
+This estimator has the following intuitive interpretation.
+We use the $\wmle_i$ estimates from the first round of communication to create a subspace $\Wave=\affspan\{\wave\}\cup{\{\wmle\}}_{i=1}^m$.
+Then in the second round of optimization, we project all of our data points down into this subspace.
+Because this subspace is small, we need only a small number of data points to get a good estimate of the optimum.
+Because of the way we constructed $\Wave$,
+the optimum in $\Wave$ will be close to the true optimum.
+%The true parameter $\wstar$ should be close to this subspace.
+%In the first round of optimization (determining the $\wmle_i$s),
+%we're actually finding a subspace
+%The estimator $\wreopt$ lives in a subs
+
+\newpage
+\section{ANALYSIS}
+
+\subsection{Notation}
+
+\subsection{Few data points are needed}
+\label{sec:vcdim}
+
+Define the space
+$\Wave=\affspan\{\wave_i\}_{i=1}^m\cup\{\wave\}$.
+Define $L : \mathbb{R} \to \mathbb{R}$ to be the 0-1 loss over the true data distribution.
+
+\begin{theorem}
+Let $\w^\Wave$ be an arbitrary point in $\Wave$ and $\delta\in(0,1)$.
+Then with probability at least $1-\delta$,
+\begin{equation}
+L(\wreopt) \le L(\w^\Wave) + O\left(\sqrt{\frac{m + \log(1/\delta)}{n^{reopt}}}\right)
+\end{equation}
+\end{theorem}
+\begin{proof}
+For each machine $i$, fix the set of data points $Z_i$.
+This in turn fixes the $\wmle_i$ and the space $\Wave$.
+The hypothesis class defined by $\Wave$ has VC dimension $m$ because it is linear with $m$ parameters.
+Since each of the data points in $Z^{reopt}$ is independent of $\Wave$,
+the result follows immediately from standard VC theory.
+See for example Chapter 6 of \cite{shalev2014understanding}.
+%
+%\fixme{
+    %Need to rework the definitions in Section 1 so that the data points in the second optimization don't depend on the first optimization.
+%}
+\end{proof}
+
+\subsection{In the Normal Regime}
+\label{sec:normal}
+
+\subsubsection{Assumptions}
+
+We assume that for every estimator $\hat\w$ trained on $n$ data points,
+\begin{equation}
+\label{eq:assumption}
+\hat\w - \E\hat\w \sim \frac{1}{\sqrt n}\normal{\zero}{I^{-1}_{\E\hat\w}}
+%\hat\w \sim \normal{\E\hat\w}{\frac{1}{n}I^{-1}_{\E\hat\w}}
+\end{equation}
+%
+%Throughout Section \ref{sec:normal} we assume that $\wave_i$ is normally distributed with mean $\wtave$ and covariance $\frac{1}{n}I^{-1}_{\wtave}$.
+%\begin{equation}
+%\label{eq:assumption}
+%\wmle_i \sim \normal{\E\wmle_i}{\frac{1}{n}I^{-1}_{\E\wmle_i}}
+%\end{equation}
+%Here, $I_{\wtave}$ is the Fisher information matrix at point $\wtave$.
+Here, $I_{\E\hat\w}$ is the Fisher information matrix for parameter $\E\hat\w$.
+This assumption is motivated by the central limit theorem for maximum likelihood estimation.
+%There are many versions of this theorem.
+Theorem 7.5.2 of \cite{lehmann1999elements} is particularly simple version and reproduced below.
+\begin{theorem}
+\label{thm:clt}
+Assume the following mild regularity conditions:
+\begin{enumerate}[noitemsep,topsep=0pt]%,label=(\itshape\roman*)]
+\item all data points in $Z_i$ are i.i.d.;
+\item the parameter space $\Theta$ is open;
+\item the model is identifiable and consistent;
+\item the matrix $\I_{\wtave}$ is non-singular;
+\item the support set $A = \{(y,\x) : f(y,\x;\theta) > 0\}$ is independent of $\theta$;
+and \item for all $(y,\x)\in A$, $f(y,\x;\theta)$ is three times differentiable with respect to $\theta$,
+and the third derivative is continuous.
+\end{enumerate}
+Then in the limit as $n\to\infty$,
+\begin{equation*}
+\wave_i \law \wtave + \frac{1}{\sqrt n} \normal{0}{\Iinv_\wstar}
+\end{equation*}
+Here the notation $\law$ denotes convergence in law,
+which is the standard notion of convergence for a sequence of probability distributions.
+\end{theorem}
+
+Note that Theorem \ref{thm:clt} does not require the likelihood $f$ to be convex.
+
+There are many generalizations of Theorem \ref{thm:clt}.
+For example, Theorem 4.2 in \cite{spokoiny2012parametric} provides finite sample bounds with explicit dependence on the degree of model misspecification, optimization error, and the lack of independence in the data variables.
+Because Spokoiny's theorem is non-asymptotic, the theorem that the estimator is a sub-Gaussian random variable.
+
+\subsubsection{Analyzing $\wave$}
+
+We provide a simple bound that shows that averaging does not improve the bias of our estimator,
+but reduces the variance at the rate $O((nm)^{-1})$.
+This bound is well known, but our analysis is significantly simpler and provides a sharp, easy to interpret constant factor.
+The main idea of our technique is to show that the variance is normally distributed.
+The ideas in this proof will be used in Section \ref{sec:analreopt} where we show that the variance of $\wreopt$ is the same as $\wave$, but that the bias is reduced.
+
+\begin{theorem}
+Let $t>0$.
+Then with probability at least $1-\exp(-t)$,
+\begin{equation}
+\ltwo{\wstar-\wave}^2 \le \ltwo{\wstar-\E\wmle_i}^2 + \frac{v_t}{mn}
+\end{equation}
+where
+\begin{equation}
+v_t =
+\tr{\Iinv_{\wtave}}
++ 2\sqrt{\tr ({\I^{-2}_{\wtave}})t}
++ 2\ltwo{\Iinv_{\wtave}}t
+\end{equation}
+\end{theorem}
+\begin{proof}
+It is well known that the MSE of an estimator is the sum of its bias and variance.
+That is,
+\begin{equation}
+\ltwo{\wstar-\wave}^2 = \ltwo{\wstar-\E\wave}^2 + \ltwo{\E\wave-\wave}^2
+\label{eq:biasvar}
+\end{equation}
+By the linearity of expectation, we have that
+\begin{align}
+\E\wave
+&=
+\E\frac{1}{m}\sum_{i=1}^m\wmle_i
+%\\&=
+=
+\frac{1}{m}\sum_{i=1}^m\E\wmle_i
+=
+\E\wmle_i
+\label{eq:expwave}
+\end{align}
+and so the bias
+$\ltwo{\wstar-\E\wave}
+=
+\ltwo{\wstar-\E\wmle_i}
+$.
+Next, we see that the variance is a Gaussian random variable.
+We have by Equations \ref{eq:expwave} and \ref{eq:assumption} that
+\begin{align}
+\wave-\E\wave
+&=
+\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wave\right)
+\\&=
+\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)
+\\&\sim
+\frac{1}{m}\sum_{i=1}^m\left(\frac{1}{\sqrt n}\normal\zero{I^{-1}_{\E\wave}}\right)
+\\&=
+\frac{1}{\sqrt {nm}}\normal\zero{I^{-1}_{\E\wave}}
+%\\&\sim
+%\frac{1}{m}\normal{m\E\wmle_i}{\frac{m}{n}I^{-1}}
+%\\&\sim
+%\normal{\E\wmle_i}{\frac{1}{nm}I^{-1}}
+%\\
+%\wave-\E\wmle_i
+%&\sim
+%\normal{\zero}{\frac{1}{nm}I^{-1}}
+%\\
+%\wave-\E\wave
+%&\sim
+\end{align}
+%Centering the distribution and applying Equation \ref{eq:expwave} gives
+%\begin{equation}
+%\E\wave-\wave
+%\sim
+%\normal{\zero}{\frac{1}{nm}I^{-1}}
+%\end{equation}
+So bounding the variance reduces to bounding the norm of a Gaussian.
+Applying the bound in Proposition 1.1 from \cite{hsu2012tail} gives that with probability at least $1-\exp(-t)$,
+\begin{equation}
+\ltwo{\E\wave-\wave}^2 \le \frac{v_t}{\sqrt{nm}}
+\label{eq:ct1}
+\end{equation}
+%where
+%\begin{equation}
+%\begin{aligned}
+%v_t'
+   %= \tr\left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)
+%+ 2&\sqrt{\tr \left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)^2t}&
+%\\+ 2\ltwo{\frac{1}{nm}\I^{-1}_{\wtave}}t&
+%\end{aligned}
+%\end{equation}
+%Then by the linearity of trace and scalability of norms,
+%\begin{equation}
+%v_t' =
+%\frac{1}{mn}\left(
+%\tr\I^{-1}_{\wtave}
+%+ 2\sqrt{\tr \I^{-2}_{\wtave}t}
+%+ 2\ltwo{\I^{-1}_{\wtave}}t
+%\right)
+%\label{eq:ct2}
+%\end{equation}
+Substituting Equations \ref{eq:expwave} and \ref{eq:ct1} into Equation \ref{eq:biasvar} gives the stated result.
+
+\end{proof}
+
+%\begin{theorem}
+%Let $t>0$.
+%Then with probability at least $1-\exp(-t)$,
+%\begin{equation}
+%\ltwo{\wave-\wtave} \le \sqrt{\frac{v_t}{mn}}
+%\end{equation}
+%where
+%\begin{equation}
+%v_t =
+%\tr{\Iinv_{\wtave}}
+%+ 2\sqrt{\tr ({\I^{-2}_{\wtave}})t}
+%+ 2\ltwo{\Iinv_{\wtave}}t
+%\end{equation}
+%\end{theorem}
+
+%\begin{proof}
+%By our assumption above, we have that
+%\begin{equation}
+%\wave \sim \wtave+\normal{0}{\frac{1}{nm}I^{-1}_{\wtave}}
+%\end{equation}
+%which implies
+%\begin{equation}
+%\wave - \wtave \sim \normal{0}{\frac{1}{nm}I^{-1}_{\wtave}}
+%\end{equation}
+%So our problem reduces to bounding the norm of a Gaussian random variable.
+%Applying the bound in Proposition 1.1 from \cite{hsu2012tail} gives the stated result.
+%\end{proof}
+
+%\ignore{
+%We introduce the following lemma due to \cite{hsu2012tail} for bounding the norm of a Gaussian random variable.
+%\begin{lemma}
+%\label{lemma:hsu}
+%Let $\x\in\mathbb{R}^d$ be an isotropic multivariate Gaussian random vector with zero mean.
+%Let $\trans A A = \Sigma$ be the covariance matrix of random vector $A\x$.
+%For all $t>0$,
+%\begin{equation}
+%\prob{\ltwo{A\x}^2 > \tr{\Sigma} + 2\sqrt{\tr (\Sigma^2)t} + 2\ltwo{\Sigma}t}
+%\le
+%\exp(-t)
+%\end{equation}
+%\end{lemma}
+%A straightforward application of this lemma gives us that with probability at least $1 - \exp(-\tvar)$,
+%\begin{equation}
+%\begin{split}
+%%\prob{
+%&\ltwo{\wave-\wtave}
+%\\
+%&\le
+%\sqrt{
+    %\frac{1}{m}
+    %\left(\tr{\Iinv_{\wtave}}
+        %+ 2\sqrt{\tr ({\Iinv_{\wtave}}^2)\tvar}
+        %+ 2\ltwo{\Iinv_{\wtave}}\tvar
+    %\right)
+%}
+%%}
+%\end{split}
+%\end{equation}
+%}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Analyzing $\wreopt$}
+\label{sec:analreopt}
+
+We now extend the analysis of the previous section to the $\wreopt$ estimator.
+We show that its variance is bounded by the variance of $\wave$,
+but that the bias shrinks as $O(1-m/d)$.
+
+Recall from Section \ref{sec:vcdim} that the estimator $\wreopt$ has low error with high probability compared to any parameters in the space $\Wave$.
+Denote by $\proj\Wave$ the projection operator onto the space $\Wave$.
+In this section, we analyze the bias and variance of $\proj\Wave\wstar$.
+
+\begin{theorem}
+Let $t>0$.
+Then we have that
+\begin{equation}
+\ltwo{\wstar-\proj\Wave\wstar}
+\le
+b_t\sqrt{\left(1-\frac{m}{d}\right)}
++
+\sqrt\frac{v_{t}}{nm}
+\end{equation}
+where
+\begin{equation}
+b_t = \ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
+\end{equation}
+with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
+and $v_t$ is defined as in Theorem \ref{thm:}, Equation \ref{eq:} with probability at least $1-\exp(-t)$.
+\end{theorem}
+%Define the spaces
+%\begin{align}
+%\W&=\vecspan\{\wmle_i - \wave\}_{i=1}^m\\
+%\Wave&=\affspan\{\wave_i\}_{i=1}^m\cup\{\wave\}\\
+%\Wtave&=\affspan\{\wave_i\}_{=1}^m\cup\{\wtave\}
+%%\Wave&=\vecspan\{\wave_i-\wave\}_{i=1}^m\\
+%%\Wtave&=\vecspan\{\wave_i-\wtave\}_{i=1}^m
+%\end{align}
+%Denote by $\proj\A$ to be the projection operator onto the space $\A$.
+%Denote by $I_\w$ the Fischer information matrix at the point $\w$.
+
+\ignore{
+\begin{theorem}
+%Assume that $n$ is large enough that $\wmle_i$ is distributed according to $\normal{\wtave}{\Iinv_{\wtave}}$.
+%Let $\delta\in(0,1)$.
+%Then with probability at least $1-\delta$,
+We have that
+\begin{equation}
+\ltwo{\wstar-\proj\Wave\wstar}
+\le \epsilon_{bias} + \epsilon_{var}
+\end{equation}
+where for all $\tbias>0$,
+\begin{equation*}
+%\begin{split}
+%\prob{
+    \epsilon_{bias}
+    \le
+    \ltwo{\wstar-\wtave}\sqrt{(\tbias + 1)\left(1-\frac{m}{d}\right)}
+%}
+%\\
+%\ge
+%1-\exp \bigg(-\frac{b}{2} (\beta-\ln(\beta+1))\bigg)~~&
+%\end{split}
+\end{equation*}
+with probability at least
+$1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$
+and for all $\tvar>0$,
+\begin{equation*}
+\begin{split}
+&\epsilon_{var}
+\\
+&\le
+\sqrt{
+    \frac{1}{m}
+    \left(\tr{\Iinv_{\wtave}}
+        + 2\sqrt{\tr ({I^{-2}_{\wtave}})\tvar}
+        + 2\ltwo{\Iinv_{\wtave}}\tvar
+    \right)
+}
+\end{split}
+\end{equation*}
+with probability at least $1-\exp(-\tvar)$.
+
+%\begin{equation}
+%\begin{split}
+%&\ltwo{\wstar-\proj\Wave\wstar}
+%\\
+%&\le
+%\ltwo{\wstar-\wtave}\sqrt{(\beta + 1)\left(1-\frac{m}{d}\right)}
+%\\
+%&+
+%\sqrt{
+    %\frac{1}{m}
+    %\left(\tr{\Iinv_{\wtave}}
+        %+ 2\sqrt{\tr ({\Iinv_{\wtave}}^2)t}
+        %+ 2\ltwo{\Iinv_{\wtave}}t
+    %\right)
+%}
+%\le
+%\exp \bigg(-\frac{b}{2} (\beta-\ln(\beta+1))\bigg)~~&
+%\end{split}
+%\end{equation}
+\end{theorem}
+}
+
+\begin{proof}
+Note that the space $\Wave$ depends on the data.
+Therefore, we can think of $\proj\Wave\wstar$ as an estimator of $\wstar$.
+As in the proof of Theorem \ref{thm:},
+we begin by decomposing the MSE into its bias and variance components.
+\begin{equation}
+\begin{split}
+\ltwo{\wstar-\proj\Wave\wstar}^2
+&=
+\ltwo{\wstar-\E\proj\Wave\wstar}^2
+\\&+
+\ltwo{\E\proj\Wave\wstar-\proj\Wave\wstar}^2
+\end{split}
+\end{equation}
+%Define $\Wtave = \vecspan\{\wmle_i-\wtave\}_{i=1}^m$.
+%Then,
+We have that
+\begin{align}
+\ltwo{\wstar-\proj\Wave\wstar}
+&\le
+\ltwo{\wstar-\proj\Wtave\wstar}
++
+\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
+\\
+&=
+\ltwo{\wstar-\proj\Wtave\wstar}
++
+\ltwo{\wave-\wtave}
+\label{eq:errdecomp}
+\end{align}
+The first line follows from the triangle inequality.
+The second line follows because every point in $\Wtave$ is a translation of a point in $\Wave$ by the vector $\wave-\wtave$.
+
+%\begin{lemma}
+%Let $a$ and $b$ be integers with $b<a$.
+%Let $\x$ be a random vector on the $a$ dimensional unit sphere.
+%Let $T : \mathbb{R}^a \to \mathbb{R}^b$ be the projection operator that preserves the first $b$ coordinates of a vector and drops the rest.
+%Then for all $\beta>0$,
+%\begin{equation}
+%\prob{\ltwo{T\x} \ge \sqrt{(\beta + 1)\frac{b}{a}}}
+%\le
+%\exp \left(\frac{b}{2}(-\beta+\ln(\beta+1))\right)
+%\end{equation}
+%\end{lemma}
+We will now bound the two terms in Equation \ref{eq:errdecomp} separately.
+To bound the left term, we introduce the following lemma.
+It is adapted from Lemma 2.2 in \cite{dasgupta2003elementary}.
+It has been simplified to ignore cases irrelevant for our purpose.
+\begin{lemma}
+Let $a$ and $b$ be integers with $b<a$.
+Let $\x\in\mathbb{R}^a$ be an arbitrary vector.
+Let $T : \mathbb{R}^a \to \mathbb{R}^b$ be projection onto a subspace of size $b$ distributed uniformly at random.
+Then for all $\beta>0$,
+\begin{equation}
+\begin{split}
+\prob{\ltwo{T\x} \ge \ltwo\x\sqrt{(\beta + 1)\frac{b}{a}}}&
+\\
+\le
+\exp \bigg(-\frac{b}{2} & (\beta-\ln(\beta+1))\bigg)
+\end{split}
+\end{equation}
+\end{lemma}
+To apply this lemma, note that $\proj\Wtave$ is not a linear projection but an affine projection.
+We first apply a translation so that we are working with the linear projection $\proj\W$.
+\begin{align}
+\ltwo{\wstar-\proj\Wtave\wstar}
+&=
+\ltwo{(\wstar-\wtave)-\proj\W(\wstar-\wtave)}
+\\
+&=
+\ltwo{(I-\proj\W)(\wstar-\wtave)}
+\end{align}
+Now, $(I-\proj\W)$ is a random projection onto a subspace of dimension $d-m$.
+(\fixme{This projection is not chosen uniformly at random because the covariance of $\wmle_i$ is not the identity.})
+This gives us that for all $\tbias>0$,
+\begin{equation}
+\begin{split}
+\prob{\ltwo{\wstar-\proj\Wtave\wstar} \ge \ltwo{\wstar-\wtave}\sqrt{(\tbias + 1)\left(1-\frac{m}{d}\right)}}&
+\\
+\le
+\exp \bigg(-\frac{b}{2} (\tbias-\ln(\tbias+1))\bigg)~~&
+\end{split}
+\end{equation}
+
+\end{proof}
+
+%The following lemma is adapted from Theorem 3.1 in \cite{matouvsek2008variants}.
+%It has been simplified to ignore cases irrelevant for our purpose.
+%\begin{lemma}
+%Let $a$ and $b$ be integers with $a>b$.
+%Let $T\in\mathbb{R}^{a\times b}$ be a random matrix with entries drawn i.i.d. from the standard normal distribution.
+%Let $\x\in\mathbb{R}^a$ be an arbitrary vector.
+%Then there exists a constant $c$ such that for all $\delta\in(0,1)$,
+%\begin{equation*}
+%\prob{\ltwo{T\x} \le \left(\sqrt{b}+c\sqrt{\log\frac{\delta}{2}}\right)\ltwo\x} \le 1-\delta
+%\end{equation*}
+%\end{lemma}
+
+%\begin{cor}
+%%Assume that Conditions (1-5) hold.
+%%Then as $n\to\infty$,
+%%\begin{enumerate}
+%%\item
+%If $f$ is locally $\sigma$-strongly convex,
+%\begin{equation}
+%\ltwo{\wstar-\wreopt} \le
+%\end{equation}
+%%\end{enumerate}
+%\end{cor}
+%
+%\begin{proof}
+%\end{proof}
+
+%\subsection{Proposed Solution}
+%
+%We propose the following solution.
+%Define the matrix $W : \mathbb{R}^{d\times m}$ to have $i$th column equal $\wmle_i$.
+%%That is,
+%%\begin{equation}
+%%W = (\wmle_1, \wmle_2, ..., \wmle_m)
+%%\end{equation}
+%Then our estimator is defined as
+%\begin{equation}
+%\begin{aligned}
+%\alpha &= \argmax_\w \sum _{(\x,y)\in Z_i} f(y,\trans\x W\w)
+%\\
+%\w &= \trans W \alpha
+%\end{aligned}
+%\end{equation}
+%Note that the inner $\w$ above is a vector of dimension $m$, not $d$ (as elsewhere).
+%In words
+%%In words, this estimator takes the data set $Z'$ and projects it onto the subspace spanned by the
+
+\newpage
+
+\clearpage
 
@@
 
-%\section{OUR ALGORITHM}
-%
-%\section{ANALYSIS}
-%
-%\section{EXPERIMENTS}
-%
-%\section{CONCLUSION}
+\newpage
+\ignore{
+\section{ANALYSIS}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+First we justify why only a small number of data points is needed for the second optimization.
+For simplicity, we'll restrict our discussion to the case of binary classification.
+This lets us use standard VC-dimension arguments.
+
+\subsection{Few datapoints are needed in the optimization}
+
+%Recall the following theorem from standard VC theory.
+%\begin{theorem}
+%Let $Z$ be a dataset of $n$ points sampled i.i.d. from distribution $\D$.
+%Let $\wmle$ be the parameters of the maximum likelihood estimator of $Z$.
+%Let $L : \Theta \to \mathbb{R}$ be the true 0-1 loss on the distribution $\D$.
+%Let $d$ be the VC dimension of the hypothesis class $\Theta$.
+%Then with probability at least $1-\delta$,
+%\begin{equation}
+%L(\wmle) \le L(\wstar) + O\left(\sqrt{\frac{d + \log(1/\delta)}{n}}\right)
+%\end{equation}
+%\end{theorem}
+%For an extended discussion of the above theorem,
+%see for example Chapter 6 of \cite{shalev2014understanding}.
+\begin{theorem}
+%Let $\Y=\{0,1\}$.
+Define $\alpha^{*}$ to be the optimal parameter constrained to the space spanned by the columns of $W$.
+In notation,
+\begin{equation}
+\alpha^{*} = \argmax_{\alpha} \int_{\Y\times\X} f(y,\trans\x W\alpha) d(y,\x)
+\end{equation}
+Define $L : \mathbb{R} \to \mathbb{R}$ to be the 0-1 loss over the true data distribution.
+Then with probability at least $1-\delta$,
+\begin{equation}
+L(\alpha^*) \le L(\alpha) + O\left(\sqrt{\frac{m + \log(1/\delta)}{n^{reopt}}}\right)
+\end{equation}
+\end{theorem}
+\begin{proof}
+For each machine $i$, fix the set of data points $Z_i$
+(which in turn fixes the $\wmle_i$).
+Define $\A$ to be the space of parameters
+\begin{equation}
+\A=\{\trans W\w : \w \in \Theta\}
+\end{equation}
+The linear hypothesis class defined by $\A$ has VC dimension $m$.
+Since each of the data points in $Z^{reopt}$ is independent of $W$,
+the result follows immediately from standard VC theory.
+See for example Chapter 6 of \cite{shalev2014understanding}.
+\end{proof}
+
+\begin{theorem}
+
+\end{theorem}
+
+\ignore{
+\subsection{Overall growth}
+
+In this section we assume the following regularity conditions.
+
+\begin{theorem}
+Assume conditions (1-5) above hold.
+%Let $m$ be arbitrary.
+Then in the limit as $n$ approaches $\infty$:
+\begin{enumerate}
+\item
+If $f$ is locally $\sigma$-strongly convex around $\wstar$
+\begin{equation}
+\Pr{a}
+\end{equation}
+\item
+If $f$ is locally $\ell$-Lipschitz around $\wstar$
+(i.e. ),
+then
+\begin{equation}
+\end{equation}
+\end{enumerate}
+\end{theorem}
+
+\begin{lemma}
+\label{lem:normaff}
+Let $W : \mathbb{R}^{d\times m}$ be a random matrix where each entry is distributed as a standard Gaussian.
+Let $\Wave$ be the affine hull of the columns of $W$;
+that is,
+\begin{equation}
+\Wave = \left\{\sum_{i=1}^m c_i\wmle_i : c\in\mathbb{R}, \sum_{i=1}^m c_i = 1 \right\}
+\end{equation}
+Then for all $t$,
+\begin{equation}
+\Pr\left[\ltwo{\proj{W}0}^2 > O(t(1-m/d))\right] \le 1 - \exp(-t)
+\end{equation}
+\end{lemma}
+}
+
+\section{EXPERIMENTS}
+
+\subsection{Synthetic Data}
+
+\subsection{Real World Advertising Data}
+
+\section{CONCLUSION}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\ignore{
+\newpage
+\section*{Appendix: Proof of Lemma \ref{lem:normaff}}
+
+%\cite{hsu2012tail}
+%\begin{lemma}
+%\label{lemma:hsu}
+%Let $\x\in\mathbb{R}^d$ be an isotropic multivariate Gaussian random vector with zero mean.
+%For all $t>0$,
+%\begin{equation}
+%\prob{\ltwo{A\z}^2 > \tr{\Sigma} + 2\sqrt{\tr (\Sigma^2)t} + 2\ltwo{\Sigma}t}
+%\le
+%\exp(-t)
+%\end{equation}
+%\end{lemma}
+
+Define $\V$ to be the vector space spanned by the first $m-1$ columns of $\Wave$.
+That is, $\V=\vecspan\{\wmle_i\}_{i=1}^{m-1}$.
+We have that
+\begin{align}
+\ltwo{\w-\proj{\Wave}\w}
+&\le \ltwo{\w - \proj{\Wave}\proj{\V}\w}
+\\
+%&\le \ltwo{}
+& \le \ltwo{\w - \proj{\V}\w} + \ltwo{\proj{\V}\w - \proj{\Wave}\proj{\V}\w}
+\end{align}
+The first inequality follows by the definition of $\proj\Wave\w$ as the point in $\Wave$ with minimum distance to $\w$,
+and the second follows by the triangle inequality.
+%\begin{align}
+%\ltwo{\w - \proj{H_m}\w}
+%& \le \ltwo{\w - \proj{H_m}\proj{V_{m-1}}\w} & \text{by definition of $\proj{}$}\\
+%& \le \ltwo{\w - \proj{V_{m-1}}\w} + \ltwo{\proj{V_{m-1}}\w - \proj{H_m}\proj{V_{m-1}}\w} &\text{by triangle ineq.}
+%\end{align}
+We will bound each of these terms separately.
+
+The left term above can be bound directly using standard random projection techniques.
+%The following lemma is a restatement of Lemma 2.2 from \cite{dasgupta2003elementary}.
+%It has been simplified to ignore cases irrelevant for our purpose.
+%\begin{lemma}
+%\label{lem:dasgupta}
+%Let $\x\in\mathbb{R}^d$ be an arbitrary vector and $W\in\mathbb{R}^{d\times m}$ be a random matrix with entries drawn i.i.d. from the standard normal distribution.
+%Then for all $t>1$,
+%\begin{equation*}
+%\prob{\ltwo{W\x} \ge \sqrt\frac{tm}{d}} \le \exp\left(\frac{m}{2}(1-t+\ln t)\right)
+%\end{equation*}
+%%Let $\delta\in(0,1)$ and $c$ be a suitable constant.
+%%Then for all $\x\in\mathbb{R}^d$, we have
+%%\begin{equation}
+%%\prob{\ltwo{W\x} \le (1+\epsilon)\ltwo{\x}} \ge 1 - \delta
+%%\end{equation}
+%\end{lemma}
+%To apply Lemma \ref{lem:dasgupta},
+%notice that $\ltwo{\w-\proj{\V}\w} = \ltwo{(I-\proj\V)\w}$.
+%\begin{equation}
+%\prob{\ltwo{\w-\proj{V}\w} \ge \sqrt{t\left(1-\frac{m}{d}\right)}} \le \exp\left(\frac{d-m}{2}(1-t+\ln t)\right)
+%\end{equation}
+%
+The following lemma is adapted from Theorem 3.1 in \cite{matouvsek2008variants}.
+It has been simplified to ignore cases irrelevant for our purpose.
+\begin{lemma}
+Let $a$ and $b$ be integers with $a>b$.
+Let $T\in\mathbb{R}^{a\times b}$ be a random matrix with entries drawn i.i.d. from the standard normal distribution.
+Let $\x\in\mathbb{R}^a$ be an arbitrary vector.
+Then there exists a constant $c$ such that for all $\delta\in(0,1)$,
+\begin{equation*}
+\prob{\ltwo{T\x} \le \left(\sqrt{b}+c\sqrt{\log\frac{\delta}{2}}\right)\ltwo\x} \le 1-\delta
+\end{equation*}
+\end{lemma}
+}
+
 \end{document}

### 2016-09-14 paper.tex
lots of changes

@@
 \documentclass[twoside]{article}
-\usepackage{aistats2015}
+\usepackage{aistats2017}
 
@@
 \usepackage[inline]{enumitem}
+
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+%\usepackage{algorithmicx}
+%\usepackage{algorithm2e}
+
 %\usepackage[authordate,bibencoding=auto,strict,backend=biber,natbib]{biblatex-chicago}
@@
 \bibliographystyle{plainnat}
+
+\usepackage{tikz}
+\usepackage{pgfplots}
+\pgfplotsset{width=7cm,compat=1.8}
 
@@
 \newcommand{\W}{\mathcal{\hat W}}
+\newcommand{\Waff}{\mathcal{\hat A}}
 \newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
@@
 \newcommand{\x}{\mathbf{x}}
+\newcommand{\y}{\mathbf{y}}
 \newcommand{\w}{\theta}
@@
 \newcommand{\wave}{\hat\w^{ave}}
-\newcommand{\wtave}{\hat\w^{ave,*}}
+\newcommand{\wtave}{\E\hat\w^{ave}}
 \newcommand{\waver}{\hat\w^{ave,r}}
@@
 
-\newcommand{\I}{I}
-\newcommand{\Iinv}{I^{-1}}
+\newcommand{\I}{\mathcal I}
+%\newcommand{\I^{-1}}{\I^{-1}}
 \newcommand{\law}{\ensuremath{\xrightarrow{L}}}
@@
 \newcommand{\trans}[1]{\ensuremath{{#1}^{\mathsf{T}}}}
-\newcommand{\ltwo}[1]{{\left\lVert {#1} \right\rVert}_2}
+\newcommand{\ltwo}[1]{{\left\lVert {#1} \right\rVert}}
 \newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
@@
 \section{INTRODUCTION}
-We consider the problem of distributed Maximum Likelihood Estimation (MLE).
-
-%In a Generalized Linear Model,
-%\begin{equation}
-%\begin{aligned}
-%y &= g^{-1}(\trans\x\w) + \epsilon
-%,~~~
-%\epsilon \sim \normal{0}{\sigma}
-%\end{aligned}
-%\end{equation}
+
+Many modern datasets are too large to fit in the memory of a single machine,
+so they must be partitioned onto many machines.
+%These datasets are typically partitioned onto many machines.
+%Statistical learning algorithms are then run in a distributed manner.
+%There are two main strategies for these distributed algorithms.
+To analyze these datasets, we need distributed algorithms.
+Existing distributed algorithms can be classified as either interactive or non-interactive.
+In this paper we propose an algorithm that exhibits the benefits of both types.
+
+\emph{Interactive algorithms} require many rounds of communication between machines.
+These algorithms often resemble standard iterative algorithms where each iteration is followed by a communication step.
+The appeal of interactive algorithms is that they enjoy the same statistical regret bounds as standard sequential algorithms.
+Unfortunately, these algorithms can be too slow in practice because communication is the main bottleneck in modern distributed architectures,
+and these algorithms require many rounds of communication.
+
+\emph{Non-interactive} algorithms require only a single round of communication.
+They are significantly faster than interactive algorithms
+but have worse regret bounds.
+Recent work (discussed in Section \ref{sec:bounds}) has shown that no non-interactive algorithm can achieve regret bounds comparable to an interactive one.
+The popular large scale learning system Vowpal Wabbit cleverly uses both types of algorithms to achieve good real world performance.
+A non-interactive algorithm is run first to provide a good warm start to a subsequent interactive algorithm \citep{vw,agarwal2014reliable}.
+%Synonyms for non-interactive algorithms used in the literature include \emph{one-shot distributed learning} and \emph{communication efficient learning}.
+
+In this paper, we propose a \emph{semi-interactive} distributed algorithm called XXX.
+Our algorithm performs two rounds of communication,
+so it is not subject to the existing regret bounds of non-interactive algorithms.
+The amount of information communicated in the second round is small, however,
+so our algorithm retains the speed advantages of non-interactive algorithms.
+
+%In the remainder of this paper, we first formally describe our problem and proposed solution.
+%We then provide a detailed comparison between our method and existing methods.
+%We then provide a theoretical analysis.
+%We conclude with an empirical analysis.
+
+%\section{PROBLEM AND SOLUTION}
+
+We now formally describe our problem and proposed solution.
 
@@
 The log-likelihood of data point $(\x,y)\in\X\times\Y$ given the model's true parameter $\wstar\in\Theta$ is denoted by $f(y,\trans\x\wstar)$.
-%\begin{equation}
-%y = g^{-1}(\trans\x\wstar)
-%\end{equation}
-%Here, $g : \mathbb{R} \to \mathbb{R}$ is the link function and $\wstar\in\Theta\subseteq\mathbb{R}^d$ is the true parameter of our model.
+Our analysis in Section \ref{sec:anal} places very mild restrictions on $f$.
+In particular, $f$ need not be convex.
+%For our experiments in Section \ref{sec:exp}, however, we restrict ourselves to the likelihood for logistic regression
+%\begin{equation}
+%f(y,\trans\x\w) =
+%\end{equation}
 Let $Z\subset\X\times\Y$ be a dataset of $mn$ i.i.d. observations.
@@
 \end{equation}
-Assume that $Z$ has been split up onto $m$ machines so that each machine $i$ has dataset $Z_i$ of size $n$, and all the $Z_i$ are disjoint.
+Assume that $Z$ has been partitioned onto $m$ machines so that each machine $i$ has dataset $Z_i$ of size $n$, and all the $Z_i$ are disjoint.
 Then each machine calculates the local estimator
@@
 
-The averaging estimator is defined as
-\begin{equation}
-\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
-\end{equation}
-Define the matrix $\hat W : \mathbb{R}^{d\times m}$ to have $i$th column equal $\wmle_i-\wave$.
+%The averaging estimator is defined as
+%\begin{equation}
+%\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
+%\end{equation}
+%Then our estimator is defined as
+%\begin{equation}
+%\begin{aligned}
+%\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x (\hat W\alpha + \wave)\right)
+%\\
+%\wreopt &= \hat W \hat\alpha + \wave
+%\end{aligned}
+%\end{equation}
+%where $Z^{reopt}$ is a small dataset of size $O(m)$.
+Define the matrix $\hat W : \mathbb{R}^{d\times m}$ to have $i$th column equal to $\wmle_i$.
+Let $n^{reopt}$ be an integer less than $n$,
+and $Z^{reopt}_i$ be a subset of $Z_i$ randomly sampled without replacement.
 Then our estimator is defined as
@@
 \begin{aligned}
-\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x (\hat W\alpha + \wave)\right)
+%Z_i^{proj} &= \{(y,\x\hat W) : (y,\x) \in Z_i^{reopt}\}
+%\\
+\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x \hat W\alpha \right)
 \\
-\wreopt &= \hat W \hat\alpha + \wave
+\wreopt &= \hat W \hat\alpha
 \end{aligned}
 \end{equation}
-where $Z^{reopt}$ is a small dataset of size $O(m)$.
-In the analysis of this estimator, we assume that $Z^{reopt}$ is independent of the $\wmle_i$;
-that is, the data points in $Z^{reopt}$ were not used in the single machine optimizations.
-In practice, however, this restriction is not necessary.
-
-This estimator has the following intuitive interpretation.
+%In the analysis of this estimator, we assume that $Z^{reopt}$ is independent of the $\wmle_i$;
+%that is, the data points in $Z^{reopt}$ were not used in the single machine optimizations.
+%In practice, however, this restriction is not necessary.
+
+%Notice that this estimator requires two rounds of communication.
+There are two ways to implement this estimator.
+The simplest is with two rounds of communication.
+In the first round, each machine broadcasts its value of $\wmle_i$ to every other machine.
+In the second round, each machine sends $Z_i^{proj}$ to a predesignated machine.
+This predesignated machine performs the second optimization and outputs the final answer.
+The full procedure is shown in Algorithm \ref{alg:distributed} above.
+There is an alternative implementation if there is a common set of data points $Z^{common}$ possessed by each machine in addition to $Z_i$.
+
+Our estimator has the following intuitive interpretation.
+Let $\W$ be the column space of $\hat W$.
 We use the $\wmle_i$ estimates from the first round of communication to create a subspace $\Wave=\affspan\{\wave\}\cup{\{\wmle\}}_{i=1}^m$.
@@
 
-\newpage
+\begin{figure}
+\hspace{-0.1in}
+\begin{tikzpicture}
+    [
+    dot/.style = {minimum width=0.15cm,inner sep=0pt,line width=0pt,fill,circle,black,font=\small}
+    ]
+\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.5,-0.5) (-1.05,1) (-0.1,-0.1) (0.75,-0.5) (0.45,-0.7) };
+\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.85,-0.65) (-1.3,1.55) (0.25,0) (2.5,-0.5) (1,-0.95)};
+\draw[gray] plot [smooth cycle,tension=1] coordinates {(-1.15,-1) (-1.5,2) (1,0) (5,0) (3,-0.95)};
+
+\draw[thick] (-2.2,2.15) -- (5.5,1);
+
+\node[dot] (wstar) at (-0.25,-0.25) {};
+\node at (0,-0.45) {$\wstar$};
+
+%\node[dot] (wstarproj) at (0.1,1.8) {};
+%\draw (wstar) -- (wstarproj);
+%\draw (0.07,1.65) -- (0.23,1.62) -- (0.26,1.78);
+
+%\node[dot] (wavestar) at (4,0.5) {};
+%\node                 at (4.5,0.3) {$\E\wave$};
+\node[dot] (wave) at (3.8,1.25) {};
+\node at (4.0,1.6) {$\wave$};
+\node[dot] (wave) at (2.9,1.40) {};
+\node at (3.1,1.75) {$\wmle_1$};
+\node[dot] (wave) at (4.7,1.10) {};
+\node at (4.9,1.45) {$\wmle_2$};
+%\node[dot] (wave) at (3.8,1.25) {};
+
+\node[dot] at (-1.35,2.03) {};
+\node at (-1.1,2.4) {$\wreopt$};
+
+\node at (4,-1.25) {$f(y,\trans\x\w)$};
+\node at (5.75,1) {$\W$};
+
+\end{tikzpicture}
+\caption{
+    Our method performs a second round of optimization to find the best parameter vector in $\W$.
+    Since $\wave\in\W$, our estimator $\wreopt$ is guaranteed to have higher true likelihood with high probability.
+}
+\label{fig:contour}
+\end{figure}
+
+\begin{algorithm}[t]
+\caption{Estimation in two rounds}
+\label{alg:distributed}
+\begin{algorithmic}
+\State Each machine $i$ independently
+\State ~~~~~loads dataset $Z_i$
+\State ~~~~~calculates $\wmle_i$
+\State All machines broadcast $\wmle_i$ to all other machines
+\State Each machine $i$ independently
+\State ~~~~~constructs $\hat W$
+\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
+\State ~~~~~calculates $\trans\x\hat W$
+\State All machines broadcast $Z^{reopt}_i$ to a master machine
+\State The master calculates $\wreopt$
+\end{algorithmic}
+%\label{fig:alg}
+\end{algorithm}
+
+\begin{algorithm}[t]
+\caption{Estimation with common data}
+\label{alg:distributed}
+\begin{algorithmic}
+\State Each machine $i$ independently
+\State ~~~~~loads dataset $Z_i$
+\State ~~~~~calculates $\wmle_i$
+\State All machines broadcast $\wmle_i$ to all other machines
+\State Each machine $i$ independently
+\State ~~~~~constructs $\hat W$
+\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
+\State ~~~~~calculates $\trans\x\hat W$
+\State All machines broadcast $Z^{reopt}_i$ to a master machine
+\State The master calculates $\wreopt$
+\end{algorithmic}
+%\label{fig:alg}
+\end{algorithm}
+
+\section{RELATED WORK}
+
+Related work can be divided into two categories:
+alternative $q$ estimators,
+and bounds on the communication complexity of distributed learning.
+
+\subsection{Alternative estimators}
+\label{sec:alt}
+The simplest and most popular communication-efficient estimator is the averaging estimator
+\begin{equation}
+\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
+\end{equation}
+
+Previous analysis of $\wave$ makes a number of limiting assumptions.
+\cite{mcdonald2009efficient} gave the first analysis of this estimator in the special case of L2 regularized maximum entropy models.
+They provide tail bounds on $\ltwo{\wstar-\wave}$, showing that the variance of the estimator reduces as $O((nm)^{-1/2})$ with high probability,
+but they could not show a reduction in bias.
+Their analysis uses a martingale technique that requires the radius of the dataset be independent of the size of the dataset.
+This is a particularly limiting assumption as even the simple case of Gaussian distributed data does not satisfy it.
+\cite{zhang2012communication} provide a more general analysis showing that the Mean Squared Error (MSE) $\E\ltwo{\wstar-\wave}^2$ decays as $O((nm)^{-1} + n^{-2})$.
+This matches the optimal MSE of $\wmle$ whenever $m<n$.
+Their analysis still requires a number of technical assumptions.
+For example, they assume the parameter space $\Theta$ is bounded.
+This assumption does not hold under the standard Bayesian interpretation of L2 regularization as a Gaussian prior of the parameter space.
+They further make strong convexity and 8\emph{th} order smoothness assumptions which guarantee that $\wmle_i$ is a ``nearly unbiased estimator'' of $\wstar$.
+(As we will see in the next section,
+recently developed information theoretic bounds show these assumptions are required to show a reduction in MSE to the single machine rate.)
+Surprisingly, \cite{zhang2013divide} showed that in the special case of kernel ridge regression, a reduction in bias was not needed to have the MSE decay at the optimal sequential rate.
+By a careful choice of regularization parameter, they caused $\wmle_i$ to have lower bias but higher variance, so that the final estimate of $\wave$ had both reduced bias and variance.
+
+Other research has focused on modifications to the $\wave$ estimator to reduce bias.
+\cite{zinkevich2010parallelized} showed that if the training sets partially overlap each other (instead of being disjoint), then the resulting estimator will have lower bias.
+The authors admit their analysis is ``highly technical.''
+\cite{zhang2012communication} also provided a bootstrap average estimator,
+which works as follows.
+Let $r\in(0,1)$, and $X_i^r$ be a bootstrap sample of $X_i$ with $rn$ data points.
+Then the bootstrap average estimator is
+\begin{equation}
+\begin{aligned}
+\wmler_i &= \argmax_\w \sum_{\x\in X_i^r} f(x;\w)
+\\
+\waver &= \frac{1}{m}\sum_{i=1}^m \wmler_i
+\\
+\wboot & = \frac{\wave-r\waver}{1-r}
+\end{aligned}
+\end{equation}
+The intuition behind this estimator is to use the bootstrap sample to directly estimate and correct for the bias.
+This estimator enjoys a MSE decaying as $O((nm)^{-1}+n^{-3})$.
+
+%An alternative definition of the $\wave$ estimator is
+%\begin{equation}
+%\wave = \argmin_\w \frac{1}{m}\sum_{i=1}^m \ltwo{\wmle_i-\w}^2
+%\end{equation}
+%It is easy to show that the two definitions are equivalent with standard calculus.
+
+\cite{liu2014distributed} propose a more Bayesian approach inspired by \cite{merugu2003privacy}.
+Instead of averaging the model's parameters,
+they directly ``average the models'' with the following KL-average estimator.
+\begin{equation}
+\wkl = \argmin_\w \sum_{i=1}^m \kl[\bigg]{p(\cdot;\wmle_i)}{p(\cdot;\w)}
+\end{equation}
+The minimization is performed via a bootstrap sample from the smaller models.
+This method has two main advantages.
+First, it is robust to reparameterizations of the model.
+Second, it is statistically optimal for the class of non-interactive optimization methods.
+(We show in the next section that their bound does not apply to our $\wreopt$ estimator due to our semi-interactive setting.)
+The main downside of the KL-average is that the minimization has a prohibitively high computational cost.
+Let $n^{kl}$ be the size of the bootstrap sample.
+Then the original implementation's MSE shrinks as $O((nm)^{-1}+(nn^{kl})^{-1})$.
+This implies that the bootstrap procedure requires as many samples as the original problem to get a MSE that shrinks at the same rate as the averaging estimator.
+\cite{han2016bootstrap} show a method to reduce this rate to $O((nm)^{-1}+(n^2n^{kl})^{-1})$ using control variates, but the procedure remains prohibitively expensive.
+Their experiments show the procedure scaling only to datasets of size $nm\approx10^4$,
+whereas our experiments involve a dataset of size $nm\approx10^8$
+
+\subsection{Performance bounds}
+\label{sec:bounds}
+
+Performance bounds come in two flavors: statistical and information theoretic.
+On the statistical side, \cite{liu2014distributed} showed that for an arbitrary function $q$ \emph{which does not depend on the data},
+$\ltwo{\wq-\wmle}^2$ decays as $\Omega(\gamma^2_\wstar \I^{-1}_\wstar/n^2)$.
+Here $\gamma_\wstar$ is the statistical curvature of the model and $\I_\wstar$ is the Fisher information.
+Furthermore, they show that $\wkl$ matches this bound.
+This bound is not relevant for our estimator because our merge function depends on the data.
+Furthermore, we directly optimize $\ltwo{\wq-\wstar}^2$ which is a more useful quantity in practice.
+
+\cite{shamir2014fundamental}, \cite{zhang2013information}, and \cite{garg2014communication} all provide information theoretic lower bounds on the sample complexity of non-interactive learning problems.
+As above, however, there results are not applicable in our semi-interactive setting.
+%provides general information theoretic bounds for what he terms learning with information constraints.
+%As a special case of Shamir's analysis is the non-interactive communication model where $q$ does not depend on the data.
+%When $q$ is allowed to depend on the data, his model no longer applies.
+There is one information theoretic lower bound that does apply to us.
+Let the true parameter vector $\wstar$ be $k$-sparse.
+That is, $\lzero{\wstar} \le k$.
+%Then \cite{zhang2013information} showed that at least $\Omega(mk/\log m)$ bits of communication are required in the non-interactive setting,
+%and \cite{garg2014communication} improved this lower bound to $\Omega(mk)$.
+Surprisingly, \cite{braverman2015communication} showed that the minimax optimal error rate for least squares regression requires $\Omega(m\cdot\min\{n,d\})$ bits of communication even in the fully interactive setting.
+This is important because sparsity does not reduce the amount of communication required, and this bound does apply in our setting.
+
 \section{ANALYSIS}
+\label{sec:anal}
 
@@
 \subsection{In the Normal Regime}
-\label{sec:normal}
+%\label{sec:normal}
+%
+%We say that a random vector $X$ is \emph{sub-Gaussian} with mean $\mu$ and covariance $\Sigma$
+%if there exists a constant $c>0$ such that
+%\begin{equation}
+%\prob{\ltwo{\Sigma(X-\mu)} > t} \le c \exp(-t^2)
+%\end{equation}
+%We denote this relationship in symbols as $X \lesssim \normal{\mu}{\Sigma}$.
+%
+%\begin{theorem}
+%We have that with probability at least
+%\begin{equation}
+%\wmle-\wstar \lesssim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
+%\end{equation}
+%In the limit as $n\to\infty$, we have the stronger statement that
+%\begin{equation}
+%\wmle-\wstar \sim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
+%\end{equation}
+%\end{theorem}
 
@@
 \label{eq:assumption}
-\hat\w - \E\hat\w \sim \frac{1}{\sqrt n}\normal{\zero}{I^{-1}_{\E\hat\w}}
+\hat\w - \E\hat\w \sim \frac{1}{\sqrt n}\normal{\zero}{\I^{-1}_{\E\hat\w}}
 %\hat\w \sim \normal{\E\hat\w}{\frac{1}{n}I^{-1}_{\E\hat\w}}
@@
 %Here, $I_{\wtave}$ is the Fisher information matrix at point $\wtave$.
-Here, $I_{\E\hat\w}$ is the Fisher information matrix for parameter $\E\hat\w$.
+Here, $\I_{\E\hat\w}$ is the Fisher information matrix for parameter $\E\hat\w$.
 This assumption is motivated by the central limit theorem for maximum likelihood estimation.
@@
 \begin{equation*}
-\wave_i \law \wtave + \frac{1}{\sqrt n} \normal{0}{\Iinv_\wstar}
+\wave_i \law \wtave + \frac{1}{\sqrt n} \normal{0}{\I^{-1}_\wstar}
 \end{equation*}
@@
 but reduces the variance at the rate $O((nm)^{-1})$.
-This bound is well known, but our analysis is significantly simpler and provides a sharp, easy to interpret constant factor.
+Similar bounds are well known (see Section \ref{sec:alt}).
+Our analysis has the following advantages:
+it requires fewer assumptions, has a simpler proof, and has an easy to interpret constant factor.
 The main idea of our technique is to show that the variance is normally distributed.
@@
 
-\begin{theorem}
-Let $t>0$.
-Then with probability at least $1-\exp(-t)$,
-\begin{equation}
-\ltwo{\wstar-\wave}^2 \le \ltwo{\wstar-\E\wmle_i}^2 + \frac{v_t}{mn}
-\end{equation}
-where
-\begin{equation}
-v_t =
-\tr{\Iinv_{\wtave}}
-+ 2\sqrt{\tr ({\I^{-2}_{\wtave}})t}
-+ 2\ltwo{\Iinv_{\wtave}}t
-\end{equation}
-\end{theorem}
+\ignore{
+\begin{lemma}
+We have that
+\begin{equation}
+\E\ltwo{\wstar-\wave}^2
+\le
+\E\ltwo{\wstar-\E\wmle_i}^2
++
+\frac{1}{m}\E\ltwo{\E\wmle_i-\wmle_i}^2
+\end{equation}
+\end{lemma}
 \begin{proof}
@@
 \begin{equation}
-\ltwo{\wstar-\wave}^2 = \ltwo{\wstar-\E\wave}^2 + \ltwo{\E\wave-\wave}^2
-\label{eq:biasvar}
-\end{equation}
+\E\ltwo{\wstar-\wave}^2
+=
+\E\ltwo{\wstar-\E\wave}^2
++
+\E\ltwo{\E\wave-\wave}^2
+\end{equation}
+First we consider the bias.
 By the linearity of expectation, we have that
@@
 and so the bias
+$
+\E\ltwo{\wstar-\E\wave}^2
+=
+\E\ltwo{\wstar-\E\wmle_i}^2
+$.
+In words, the bias of the averaging estimator is the same as the bias of the estimator on an individual machine.
+
+Next we consider the variance.
+We use the definition of $\wave$ and Equation \ref{eq:expwave} above to conclude that
+\begin{align}
+\wave-\E\wave
+&=
+\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave
+\\&=
+\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wave\right)
+\\&=
+\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)
+%\\&\sim
+%\frac{1}{m}\sum_{i=1}^m\left(\frac{1}{\sqrt n}\normal\zero{I^{-1}_{\E\wave}}\right)
+%\\&=
+%\frac{1}{\sqrt {nm}}\normal\zero{I^{-1}_{\E\wave}}
+\end{align}
+Taking the norm and squaring both sides gives
+\begin{equation}
+\ltwo{\wave-\E\wave}
+=
+\ltwo{
+\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)
+}
+\end{equation}
+\end{proof}
+}
+
+\begin{theorem}
+Let $t>0$.
+Then with probability at least $1-\exp(-t)$,
+\begin{equation}
+\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wmle_i} + \sqrt\frac{v_t}{mn}
+\end{equation}
+where
+\begin{equation}
+v_t =
+\tr{\I^{-1}_{\wtave}}
++ 2\sqrt{\tr \left({\I^{-2}_{\wtave}}\right)t}
++ 2\ltwo{\I^{-1}_{\wtave}}t
+\end{equation}
+\end{theorem}
+\begin{proof}
+We have by the triangle inequality that
+\begin{equation}
+\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wave} + \ltwo{\E\wave-\wave}
+\label{eq:biasvar}
+\end{equation}
+The left term above captures the estimator's bias and the right term the variance.
+First we consider the bias.
+By the linearity of expectation, we have that
+\begin{align}
+\E\wave
+&=
+\E\frac{1}{m}\sum_{i=1}^m\wmle_i
+%\\&=
+=
+\frac{1}{m}\sum_{i=1}^m\E\wmle_i
+=
+\E\wmle_i
+\label{eq:expwave}
+\end{align}
+and so the bias term
 $\ltwo{\wstar-\E\wave}
@@
 $.
-Next, we see that the variance is a Gaussian random variable.
+In words, the bias of the combined averaging estimator is the same as the bias of the estimator on an individual machine.
+
+Now we consider the variance.
 We have by Equations \ref{eq:expwave} and \ref{eq:assumption} that
@@
 &=
+%\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave
+%\\&=
 \frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wave\right)
+\label{eq:var1}
 \\&=
@@
 \\&\sim
-\frac{1}{m}\sum_{i=1}^m\left(\frac{1}{\sqrt n}\normal\zero{I^{-1}_{\E\wave}}\right)
+\frac{1}{m}\sum_{i=1}^m\left(\frac{1}{\sqrt n}\normal\zero{\I^{-1}_{\E\wave}}\right)
 \\&=
-\frac{1}{\sqrt {nm}}\normal\zero{I^{-1}_{\E\wave}}
+\frac{1}{\sqrt {nm}}\normal\zero{\I^{-1}_{\E\wave}}
+\label{eq:var2}
 %\\&\sim
@@
 %\end{equation}
-So bounding the variance reduces to bounding the norm of a Gaussian.
+In the last step we used the fact that adding a centered Gaussian to itself $m$ times is equivalent to multiplying the Gaussian by $\sqrt{m}$.
+%So bounding the variance reduces to bounding the norm of a Gaussian.
+Our last step is to bound the norm of the Gaussian in Equation \ref{eq:var2}.
 Applying the bound in Proposition 1.1 from \cite{hsu2012tail} gives that with probability at least $1-\exp(-t)$,
 \begin{equation}
-\ltwo{\E\wave-\wave}^2 \le \frac{v_t}{\sqrt{nm}}
+\ltwo{\E\wave-\wave} \le \sqrt\frac{v_t}{nm}
 \label{eq:ct1}
@@
 Substituting Equations \ref{eq:expwave} and \ref{eq:ct1} into Equation \ref{eq:biasvar} gives the stated result.
-
 \end{proof}
+
+Note that this reduction in variance is essentially optimal.
+By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
+we can show that the variance of the MLE trained on all of the data is
+\begin{equation}
+\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
+\end{equation}
+The only difference is the location of the Fisher information matrix.
 
@@
 %v_t =
-%\tr{\Iinv_{\wtave}}
+%\tr{\I^{-1}_{\wtave}}
 %+ 2\sqrt{\tr ({\I^{-2}_{\wtave}})t}
-%+ 2\ltwo{\Iinv_{\wtave}}t
+%+ 2\ltwo{\I^{-1}_{\wtave}}t
 %\end{equation}
@@
     %\frac{1}{m}
-    %\left(\tr{\Iinv_{\wtave}}
-        %+ 2\sqrt{\tr ({\Iinv_{\wtave}}^2)\tvar}
-        %+ 2\ltwo{\Iinv_{\wtave}}\tvar
+    %\left(\tr{\I^{-1}_{\wtave}}
+        %+ 2\sqrt{\tr ({\I^{-1}_{\wtave}}^2)\tvar}
+        %+ 2\ltwo{\I^{-1}_{\wtave}}\tvar
     %\right)
@@
 
-We now extend the analysis of the previous section to the $\wreopt$ estimator.
-We show that its variance is bounded by the variance of $\wave$,
-but that the bias shrinks as $O(1-m/d)$.
-
-Recall from Section \ref{sec:vcdim} that the estimator $\wreopt$ has low error with high probability compared to any parameters in the space $\Wave$.
-Denote by $\proj\Wave$ the projection operator onto the space $\Wave$.
-In this section, we analyze the bias and variance of $\proj\Wave\wstar$.
+%We now extend the analysis of the previous section to the $\wreopt$ estimator.
+%We show that its variance is bounded by the variance of $\wave$,
+%but that the bias shrinks as $O(1-m/d)$.
+%
+%Recall from Section \ref{sec:vcdim} that the estimator $\wreopt$ has low error with high probability compared to any parameters in the space $\Wave$.
+%Denote by $\proj\Wave$ the projection operator onto the space $\Wave$.
+%In this section, we analyze the bias and variance of $\proj\Wave\wstar$.
+
+%If $\A$ is an affine space, then denote by $\proj{\A}$ the projection operator onto $\A$.
+We introduce the following notation.
+If $\A$ is an arbitrary space,
+then the projection of a point $\x$ onto $\A$ is denoted by
+\begin{equation}
+\proj{\A}\x = \argmin_{\mathbf{a}\in\A} \ltwo{\mathbf{a}-\x}
+\end{equation}
+Recall that $\W$ is the vector space spanned by the $\wmle_i$s.
+In this section we analyze the estimator $\proj{\W}\wstar$.
 
 \begin{theorem}
+\label{thm:wave}
 Let $t>0$.
-Then we have that
-\begin{equation}
-\ltwo{\wstar-\proj\Wave\wstar}
+Then we have that with probability at least $(1-\exp(-t))^2$,
+\begin{equation}
+\ltwo{\wstar-\proj\W\wstar}
 \le
+%b_t\sqrt{\left(1-\frac{m}{d}\right)}
 b_t\sqrt{\left(1-\frac{m}{d}\right)}
@@
 where
-\begin{equation}
-b_t = \ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
-\end{equation}
-with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
-and $v_t$ is defined as in Theorem \ref{thm:}, Equation \ref{eq:} with probability at least $1-\exp(-t)$.
+%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
+\begin{equation}
+b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
+\end{equation}
+%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
+$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
+and $v_t$ is defined as in Theorem \ref{thm:wave}.
 \end{theorem}
-%Define the spaces
-%\begin{align}
-%\W&=\vecspan\{\wmle_i - \wave\}_{i=1}^m\\
-%\Wave&=\affspan\{\wave_i\}_{i=1}^m\cup\{\wave\}\\
-%\Wtave&=\affspan\{\wave_i\}_{=1}^m\cup\{\wtave\}
-%%\Wave&=\vecspan\{\wave_i-\wave\}_{i=1}^m\\
-%%\Wtave&=\vecspan\{\wave_i-\wtave\}_{i=1}^m
-%\end{align}
-%Denote by $\proj\A$ to be the projection operator onto the space $\A$.
-%Denote by $I_\w$ the Fischer information matrix at the point $\w$.
 
@@
 \begin{theorem}
-%Assume that $n$ is large enough that $\wmle_i$ is distributed according to $\normal{\wtave}{\Iinv_{\wtave}}$.
+%Assume that $n$ is large enough that $\wmle_i$ is distributed according to $\normal{\wtave}{\I^{-1}_{\wtave}}$.
 %Let $\delta\in(0,1)$.
@@
     \frac{1}{m}
-    \left(\tr{\Iinv_{\wtave}}
-        + 2\sqrt{\tr ({I^{-2}_{\wtave}})\tvar}
-        + 2\ltwo{\Iinv_{\wtave}}\tvar
+    \left(\tr{\I^{-1}_{\wtave}}
+        + 2\sqrt{\tr ({\I^{-2}_{\wtave}})\tvar}
+        + 2\ltwo{\I^{-1}_{\wtave}}\tvar
     \right)
@@
     %\frac{1}{m}
-    %\left(\tr{\Iinv_{\wtave}}
-        %+ 2\sqrt{\tr ({\Iinv_{\wtave}}^2)t}
-        %+ 2\ltwo{\Iinv_{\wtave}}t
+    %\left(\tr{\I^{-1}_{\wtave}}
+        %+ 2\sqrt{\tr ({\I^{-1}_{\wtave}}^2)t}
+        %+ 2\ltwo{\I^{-1}_{\wtave}}t
     %\right)
@@
 \begin{proof}
+This proof follows the same form as the proof for Theorem \ref{thm:wave}.
+We first decompose the error into a bias and variance component;
+then we bound each component separately.
+To decompose the error, we need to introduce the following affine space.
+\begin{align}
+\Waff&=\affspan \{\wmle_i\}_{i=1}^m
+\\
+\Waff_{\E}&=\affspan \{\wmle_i - \wave + \E\wave\}_{i=1}^m
+\end{align}
+We now have that
+\begin{align}
+\ltwo{\wstar-\proj\W\wstar}
+&\le
+\ltwo{\wstar-\proj\Waff\wstar}
+\\
+&\le
+\ltwo{\wstar-\proj{\Waff_\E}\wstar}
+%\\&
++
+\ltwo{\proj{\Waff_\E}\wstar-\proj\Waff\wstar}
+%\\&+
+%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
+\end{align}
+The first line above follows because $\Waff\subseteq\W$,
+and the second by the triangle inequality.
+
+%Notice that $\Waff\subseteq\W$, so $\ltwo{\wstar-\proj\W\wstar} \le \ltwo{\wstar-\proj\Waff\wstar}$.
+%\begin{equation}
+%\Wtave=\{\x - \wave + \wtave : \x\in\Wave\}
+%\end{equation}
+%We have by the triangle inequality that
+%\begin{equation}
+%\begin{split}
+%\ltwo{\wstar-\proj\Wave\wstar}
+%&\le
+%\ltwo{\wstar-\proj\Wtave\wstar}
+%\\&+
+%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
+%\end{split}
+%\end{equation}
+The left term above can be thought of as akin to the bias of the estimator $\proj\Wtave\wstar$,
+and the right term can be thought of as akin to the variance.
+
+We begin by bounding the right term.
+We have that with probability at least $1-\exp(-t)$,
+\begin{align}
+&\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
+\\
+&=
+\ltwo{(\proj\Wave\wstar+\wave-\wtave)-\proj\Wave\wstar}
+\\&=
+\ltwo{\wave-\wtave}
+\\&\le
+\sqrt\frac{v_t}{mn}
+\end{align}
+The first equality follows because the space $\Wtave$ is a translation of $\Wave$ by the vector $(\wave-\wtave)$.
+The probabilistic inequality is due to Equation \ref{eq:ct1} in the proof of Theorem \ref{thm:wave}.
+
+We now bound the left term.
+We have that
+%Note that $\proj\Wtave$ is not a linear projection but an affine projection.
+%We first apply a translation so that we are working with the linear projection $\proj\W$.
+\begin{align}
+\ltwo{\wstar-\proj\Wtave\wstar}
+&=
+\ltwo{(\wstar-\wtave)-\proj\W(\wstar-\wtave)}
+\\&=
+\ltwo{(I-\proj\W)(\wstar-\wtave)}
+\end{align}
+In the first step, we translate the vectors so that the affine projection onto $\Wtave$ becomes a linear projection onto $\W$.
+The operator $(I-\proj\W)$ is a random projection onto a $(d-m)$-dimensional space.
+
+%Now, $(I-\proj\W)$ is a random projection onto a subspace of dimension $d-m$.
+%(\fixme{This projection is not chosen uniformly at random because the covariance of $\wmle_i$ is not the identity.})
+%This gives us that for all $\tbias>0$,
+%\begin{equation}
+%\begin{split}
+%\prob{\ltwo{\wstar-\proj\Wtave\wstar} \ge \ltwo{\wstar-\wtave}\sqrt{(\tbias + 1)\left(1-\frac{m}{d}\right)}}&
+%\\
+%\le
+%\exp \bigg(-\frac{b}{2} (\tbias-\ln(\tbias+1))\bigg)~~&
+%\end{split}
+%\end{equation}
+\ignore{
 Note that the space $\Wave$ depends on the data.
 Therefore, we can think of $\proj\Wave\wstar$ as an estimator of $\wstar$.
-As in the proof of Theorem \ref{thm:},
-we begin by decomposing the MSE into its bias and variance components.
+As in the proof of Theorem \ref{thm:wave},
+we begin by decomposing the MSE of $\proj\Wave\wstar$ into its bias and variance components.
 \begin{equation}
@@
 \end{equation}
+We begin with the variance.
+
 %Define $\Wtave = \vecspan\{\wmle_i-\wtave\}_{i=1}^m$.
@@
 
+Define the spaces
+\begin{align}
+\W&=\vecspan\{\wmle_i - \wave\}_{i=1}^m\\
+\Wave&=\affspan\{\wave_i\}_{i=1}^m\cup\{\wave\}\\
+\Wtave&=\affspan\{\wave_i\}_{=1}^m\cup\{\wtave\}
+%\Wave&=\vecspan\{\wave_i-\wave\}_{i=1}^m\\
+%\Wtave&=\vecspan\{\wave_i-\wtave\}_{i=1}^m
+\end{align}
+Denote by $\proj\A$ to be the projection operator onto the space $\A$.
+Denote by $\I_\w$ the Fischer information matrix at the point $\w$.
+
 %\begin{lemma}
@@
 %\end{lemma}
+
 We will now bound the two terms in Equation \ref{eq:errdecomp} separately.
@@
 \end{equation}
+}
 
@@
 %In words
-%%In words, this estimator takes the data set $Z'$ and projects it onto the subspace spanned by the
+%%In words, this estimator takes the dataset $Z'$ and projects it onto the subspace spanned by the
 
@@
 \clearpage
-
-\section{RELATED WORK}
-
-Related work can be divided into two categories:
-alternative $q$ estimators,
-and bounds on the communication complexity of distributed learning.
-
-\subsection{Alternative estimators}
-The simplest and most popular communication efficient estimator is the averaging estimator
-\begin{equation}
-\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
-\end{equation}
-\cite{mcdonald2009efficient}
- gave the first analysis of this estimator wherein they showed that the variance of the estimator reduces as $O(m^{-1/2})$, but that the bias of $\wave$ is the same as $\wmle_i$.
-\cite{zhang2012communication} showed that the Mean Squared Error (MSE) $\ltwo{\wave-\wstar}^2$ decays as $O((nm)^{-1} + n^{-2})$.
-This is slower than the optimal sequential algorithm which has a MSE of $O((mn)^{-2})$.
-Surprisingly, \cite{zhang2013divide} showed that in the special case of kernel ridge regression, a reduction in bias was not needed to have the MSE decay at the optimal sequential rate.
-By a careful choice of regularization parameter, they caused $\wmle_i$ to have lower bias but higher variance, so that the final estimate of $\wave$ had both reduced bias and variance.
-
-Other research has focused on modifications to the $\wave$ estimator to reduce bias.
-\cite{zinkevich2010parallelized} showed that if the training sets partially overlap each other (instead of being disjoint), then the resulting estimator will have lower bias.
-\cite{zhang2012communication} also provided a bootstrap average estimator,
-which works as follows.
-Let $r\in(0,1)$, and $X_i^r$ be a bootstrap sample of $X_i$ with $rn$ data points.
-Then the bootstrap average estimator is
-\begin{equation}
-\begin{aligned}
-\wmler_i &= \argmax_\w \sum_{\x\in X_i^r} f(x;\w)
-\\
-\waver &= \frac{1}{m}\sum_{i=1}^m \wmler_i
-\\
-\wboot & = \frac{\wave-r\waver}{1-r}
-\end{aligned}
-\end{equation}
-This estimator enjoys a MSE decaying as $O((nm)^{-1}+n^{-3})$.
-
-%An alternative definition of the $\wave$ estimator is
-%\begin{equation}
-%\wave = \argmin_\w \frac{1}{m}\sum_{i=1}^m \ltwo{\wmle_i-\w}^2
-%\end{equation}
-%It is easy to show that the two definitions are equivalent with standard calculus.
-
-\cite{liu2014distributed} propose a more Bayesian approach inspired by \cite{merugu2003privacy}.
-Instead of averaging the model's parameters,
-they directly ``average the models'' with the following KL-average estimator.
-\begin{equation}
-\wkl = \argmin_\w \sum_{i=1}^m \kl[\bigg]{p(\cdot;\wmle_i)}{p(\cdot;\w)}
-\end{equation}
-The minimization is performed via a bootstrap sample from the smaller models.
-This method has two main advantages.
-First, it is robust to reparameterizations of the model.
-Second, it is statistically optimal for the class of non-interactive optimization methods.
-(We show in the next section that their bound does not apply to our procedure due to our semi-interactive setting.)
-The main downside of the KL-average is that the minimization has a prohibitively high computational cost.
-Let $n^{kl}$ be the size of the bootstrap sample.
-Then the original implementation's MSE shrinks as $O((nm)^{-1}+(nn^{kl})^{-1})$.
-This implies that the bootstrap procedure requires as many samples as the original problem to get a MSE that shrinks at the same rate as the averaging estimator.
-\cite{han2016bootstrap} show a method to reduce this rate to $O((nm)^{-1}+(n^2n^{kl})^{-1})$ using control variates, but the procedure remains prohibitively expensive.
-Their experiments show the procedure scaling only to datasets of size $nm\approx10^4$,
-whereas our experiments involve a dataset of size $nm\approx10^8$
-
-\subsection{Performance bounds}
-
-Performance bounds come in two flavors: statistical and information theoretic.
-On the statistical side, \cite{liu2014distributed} showed that for an arbitrary function $q$ \emph{which does not depend on the data},
-$\ltwo{\wq-\wmle}^2$ decays as $\Omega(\gamma^2_\wstar I^{-1}_\wstar/n^2)$.
-Here $\gamma_\wstar$ is the statistical curvature of the model and $I_\wstar$ is the Fisher information.
-Furthermore, they show that $\wkl$ matches this bound.
-This bound is not relevant for our estimator because our merge function depends on the data.
-Furthermore, we directly optimize $\ltwo{\wq-\wstar}^2$ which is a more useful quantity in practice.
-
-\cite{shamir2014fundamental}, \cite{zhang2013information}, and \cite{garg2014communication} all provide information theoretic lower bounds on the sample complexity of non-interactive learning problems.
-As above, however, there results are not applicable in our semi-interactive setting.
-%provides general information theoretic bounds for what he terms learning with information constraints.
-%As a special case of Shamir's analysis is the non-interactive communication model where $q$ does not depend on the data.
-%When $q$ is allowed to depend on the data, his model no longer applies.
-There is one information theoretic lower bound that does apply to us.
-Let the true parameter vector $\wstar$ be $k$-sparse.
-That is, $\lzero{\wstar} \le k$.
-%Then \cite{zhang2013information} showed that at least $\Omega(mk/\log m)$ bits of communication are required in the non-interactive setting,
-%and \cite{garg2014communication} improved this lower bound to $\Omega(mk)$.
-Surprisingly, \cite{braverman2015communication} showed that the minimax optimal error rate for least squares regression requires $\Omega(m\cdot\min\{n,d\})$ bits of communication even in the fully interactive setting.
-This is important because sparsity does not reduce the amount of communication required, and this bound does apply in our setting.
 
@@
 }
+}
 
 \section{EXPERIMENTS}
+\label{sec:exp}
 
@@
 \section{CONCLUSION}
-}
 
@@
 %To apply Lemma \ref{lem:dasgupta},
-%notice that $\ltwo{\w-\proj{\V}\w} = \ltwo{(I-\proj\V)\w}$.
+%notice that $\ltwo{\w-\proj{\V}\w} = \ltwo{(\I-\proj\V)\w}$.
 %\begin{equation}

### 2016-09-22 paper.tex
changes

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{assumption}{Condition}
 
@@
 \newcommand{\zero}{\text{\textbf{0}}}
+\newcommand{\Zproj}{Z^{\textit{proj}}}
+\newcommand{\Zreopt}{Z^{\textit{owa}}}
+\newcommand{\nreopt}{n^{\textit{owa}}}
 \newcommand{\Q}{\mathcal{Q}}
@@
 \newcommand{\X}{\mathcal{X}}
-\newcommand{\W}{\mathcal{\hat W}}
+\newcommand{\W}{{\hat \Theta^{\textit{owa}}}}
 \newcommand{\Waff}{\mathcal{\hat A}}
+\newcommand{\WaffE}{{\mathcal{\hat A}_\E}}
 \newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
@@
 \newcommand{\wkl}{\hat\w^{kl}}
-\newcommand{\wreopt}{\hat\w^{reopt}}
+\newcommand{\ahat}{\hat\alpha}
+\newcommand{\afull}{\ahat^{\textit{full}}}
+\newcommand{\wreopt}{\hat\w^{owa}}
+\newcommand{\wreoptfull}{\hat\w^{\textit{owa,full}}}
 \newcommand{\wave}{\hat\w^{ave}}
@@
 
-@article{spokoiny2012parametric,
+@article{spokoiny2012parametricestimation,
   title={Parametric estimation. Finite sample theory},
@@
 \begin{abstract}
+We present a distributed learning algorithm with low communication complexity.
+Our algorithm is an easy to implement modification of the standard averaging method.
+In the averaging method, each machine in a cluster first solves the learning problem on a subset of the data;
+the resulting parameter vectors are then averaged together.
+In our method, we perform a second round of optimization to find the \emph{optimal weighted average} of the parameter vectors.
+This second optimization happens in a low dimensional space,
+and so it adds negligible computational and communication costs.
+%Given $m$ machines (each with $n$ data points of dimension $d$),
+%we show that our algorithm reduces bias at a rate of $O(\frac{1-m/d}{n})$.
+Whereas the averaging method reduces only the variance of the estimator,
+our method reduces both the bias and the variance.
+Our analysis is more general than the analysis of similar communication efficient algorithms that also reduce the bias.
+Notably, we do not require the likelihood function to be convex.
+In practice,
+this bias reduction makes our algorithm significantly less sensitive to the amount of regularization.
+This has the practical benefit of greatly speeding up the model selection process.
 \end{abstract}
@@
 Solving for $\wmle_i$ requires no communication with other machines.
-Our goal is to provide a function $q : \Theta^m \to \Theta$ that ``merges'' these estimates.
-The resulting merged estimate should have error as close to $\wmle$ as possible.
+%Our goal is to provide a function $q : \Theta^m \to \Theta$ that ``merges'' these estimates.
+%The resulting merged estimate should have error as close to $\wmle$ as possible.
+A non-interactive distributed estimator merges these $\wmle_i$ into a single improved estimate.
+The simplest example is the averaging estimator
+\begin{equation}
+\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
+\end{equation}
+This estimator is well studied, and in Section \ref{sec:anal} we comprae this previous work to our own.
+Here we briefly note that $\wave$ is known to reduce the variance of an estimator but not its bias.
+Our goal is to design an estimator that reduces both variance and bias.
 
@@
 
-%The averaging estimator is defined as
-%\begin{equation}
-%\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
-%\end{equation}
-%Then our estimator is defined as
-%\begin{equation}
-%\begin{aligned}
-%\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x (\hat W\alpha + \wave)\right)
-%\\
-%\wreopt &= \hat W \hat\alpha + \wave
-%\end{aligned}
-%\end{equation}
-%where $Z^{reopt}$ is a small dataset of size $O(m)$.
+In this section we present a modification to the averaging estimator called the \emph{Optimal Weighted Average} (OWA).
+OWA uses a second round of optimization to calculate the optimal linear combination of the $\wmle_i$s.
+%Using these optimal weights, we reduce bias at a rate of $O(\sqrt{1-m/d})$.
+This second optimization occurs over a small fraction of the dataset,
+so its computational and communication cost is negligible.
+
+To motivate our estimator, we first describe a simpler estimator that uses the entire dataset for the second round of optimization.
+%The idea of our proposed solution is to find the best linear combination of the $\wmle_i$s using a second round of optimization.
 Define the matrix $\hat W : \mathbb{R}^{d\times m}$ to have $i$th column equal to $\wmle_i$.
-Let $n^{reopt}$ be an integer less than $n$,
-and $Z^{reopt}_i$ be a subset of $Z_i$ randomly sampled without replacement.
-Then our estimator is defined as
-\begin{equation}
-\begin{aligned}
-%Z_i^{proj} &= \{(y,\x\hat W) : (y,\x) \in Z_i^{reopt}\}
-%\\
-\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x \hat W\alpha \right)
-\\
-\wreopt &= \hat W \hat\alpha
-\end{aligned}
-\end{equation}
-%In the analysis of this estimator, we assume that $Z^{reopt}$ is independent of the $\wmle_i$;
-%that is, the data points in $Z^{reopt}$ were not used in the single machine optimizations.
-%In practice, however, this restriction is not necessary.
-
-%Notice that this estimator requires two rounds of communication.
-There are two ways to implement this estimator.
-The simplest is with two rounds of communication.
-In the first round, each machine broadcasts its value of $\wmle_i$ to every other machine.
-In the second round, each machine sends $Z_i^{proj}$ to a predesignated machine.
-This predesignated machine performs the second optimization and outputs the final answer.
-The full procedure is shown in Algorithm \ref{alg:distributed} above.
-There is an alternative implementation if there is a common set of data points $Z^{common}$ possessed by each machine in addition to $Z_i$.
-
-Our estimator has the following intuitive interpretation.
-Let $\W$ be the column space of $\hat W$.
-We use the $\wmle_i$ estimates from the first round of communication to create a subspace $\Wave=\affspan\{\wave\}\cup{\{\wmle\}}_{i=1}^m$.
-Then in the second round of optimization, we project all of our data points down into this subspace.
-Because this subspace is small, we need only a small number of data points to get a good estimate of the optimum.
-Because of the way we constructed $\Wave$,
-the optimum in $\Wave$ will be close to the true optimum.
-%The true parameter $\wstar$ should be close to this subspace.
-%In the first round of optimization (determining the $\wmle_i$s),
-%we're actually finding a subspace
-%The estimator $\wreopt$ lives in a subs
+Now consider the estimator
+\begin{equation}
+\wreoptfull = \hat W \afull
+\end{equation}
+where
+\begin{equation}
+\afull = \argmax_\alpha \sum _{(\x,y)\in Z} f\left(y,\trans\x \hat W \alpha \right)
+\end{equation}
+%This is the maximum likelihood estimator when we restrict our
+Notice that $\wreoptfull$ is the maximum likelihood estimator when the parameter space $\Theta$ is restricted $\W = \vecspan \{\wmle_i\}_{i=1}^m$.
+Figure \ref{fig:contour} shows graphically that no other estimator in $\W$ can have higher empirical likelihood than $\wreoptfull$.
+
+Calculating the weights $\afull$ directly is infeasible because it requires access to the full dataset.
+Fortunately, we don't need to consider all the data points for an accurate estimator.
+The parameter space $\W$ is $m$-dimensional.
+So intuitively, we only need $O(m)$ data points to solve such a problem to our desired accuracy.
+%(This intuition is formalized in Section \ref{sec:anal}.)
+This intuition motivates the following estimator.
+Let $\Zreopt_i\subset Z_i$ be a set of $\nreopt$ data points uniformly sampled from $Z_i$ without replacement,
+and let $\Zreopt$ be the union of the $\Zreopt_i$s.
+Then the OWA estimator is defined as
+\begin{equation}
+\wreopt = \hat W \ahat
+\end{equation}
+where
+\begin{equation}
+\ahat = \argmax_\alpha \sum _{(\x,y)\in \Zreopt} f\left(y,\trans\x \hat W \alpha \right)
+\end{equation}
+
+Algorithm \ref{fig:alg2} shows the steps for efficiently computing $\wreopt$.
+In the first round, each machine calculates $\wmle_i$ independently and broadcasts the results to every other machine.
+Since the parameter vector has $d$ dimensions and there are $m$ machines, a total of $O(md)$ bits of get transmitted.
+In the second round, each machine projects its local dataset $\Zreopt$ onto the space $\W$.
+These projected data points are then transmitted to a predesignated master machine.
+The projected data points each have dimension $m$, so a total of $O(m^2\nreopt)$ bits are transmitted.
+The master machine now has all of the information to complete the optimization.
 
@@
     ]
-\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.5,-0.5) (-1.05,1) (-0.1,-0.1) (0.75,-0.5) (0.45,-0.7) };
-\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.85,-0.65) (-1.3,1.55) (0.25,0) (2.5,-0.5) (1,-0.95)};
-\draw[gray] plot [smooth cycle,tension=1] coordinates {(-1.15,-1) (-1.5,2) (1,0) (5,0) (3,-0.95)};
-
-\draw[thick] (-2.2,2.15) -- (5.5,1);
-
-\node[dot] (wstar) at (-0.25,-0.25) {};
-\node at (0,-0.45) {$\wstar$};
+\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.5,-0.75) (-1.05,1) (-0.1,-0.1) (0.75,-0.5) (0.45,-0.7) };
+\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.85,-0.85) (-1.3,1.55) (0.25,0) (2.5,-0.5) (1,-0.95)};
+\draw[gray] plot [smooth cycle,tension=1] coordinates {(-1.15,-1.2) (-1.5,2) (1,0) (5,0) (3,-0.95)};
+
+\draw[thick] (-2.2,2.15) -- (5.2,1.03);
+
+\node[dot] (wstar) at (-0.28,-0.25) {};
+\node at (0.05,-0.55) {$\wmle$};
 
@@
 \node at (4.0,1.6) {$\wave$};
-\node[dot] (wave) at (2.9,1.40) {};
+\node[dot] (wave) at (2.80,1.40) {};
 \node at (3.1,1.75) {$\wmle_1$};
-\node[dot] (wave) at (4.7,1.10) {};
+\node[dot] (wave) at (4.75,1.10) {};
 \node at (4.9,1.45) {$\wmle_2$};
@@
 \node[dot] at (-1.35,2.03) {};
-\node at (-1.1,2.4) {$\wreopt$};
-
-\node at (4,-1.25) {$f(y,\trans\x\w)$};
-\node at (5.75,1) {$\W$};
+\node at (-0.9,2.4) {$\wreoptfull$};
+
+\node at (4,-1.1) {$f(y,\trans\x\w)$};
+\node at (5.65,1.0) {$\W$};
 
@@
     Our method performs a second round of optimization to find the best parameter vector in $\W$.
-    Since $\wave\in\W$, our estimator $\wreopt$ is guaranteed to have higher true likelihood with high probability.
+    Since $\wave\in\W$, the estimator $\wreoptfull$ has higher empirical likelihood than $\wave$ almost surely.
 }
@@
 \State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
-\State ~~~~~calculates $\trans\x\hat W$
-\State All machines broadcast $Z^{reopt}_i$ to a master machine
+\State ~~~~~calculates $\Zproj_i=\{(\trans\x\hat W,y) : (\x,y)\in\Zreopt\}$
+\State All machines broadcast $\Zproj_i$ to a master machine
 \State The master calculates $\wreopt$
 \end{algorithmic}
-%\label{fig:alg}
+\label{fig:alg2}
 \end{algorithm}
 
-\begin{algorithm}[t]
-\caption{Estimation with common data}
-\label{alg:distributed}
-\begin{algorithmic}
-\State Each machine $i$ independently
-\State ~~~~~loads dataset $Z_i$
-\State ~~~~~calculates $\wmle_i$
-\State All machines broadcast $\wmle_i$ to all other machines
-\State Each machine $i$ independently
-\State ~~~~~constructs $\hat W$
-\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
-\State ~~~~~calculates $\trans\x\hat W$
-\State All machines broadcast $Z^{reopt}_i$ to a master machine
-\State The master calculates $\wreopt$
-\end{algorithmic}
-%\label{fig:alg}
-\end{algorithm}
+%\begin{algorithm}[t]
+%\caption{Estimation with common data}
+%\label{alg:distributed}
+%\begin{algorithmic}
+%\State Each machine $i$ independently
+%\State ~~~~~loads dataset $Z_i$
+%\State ~~~~~calculates $\wmle_i$
+%\State All machines broadcast $\wmle_i$ to all other machines
+%\State Each machine $i$ independently
+%\State ~~~~~constructs $\hat W$
+%\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
+%\State ~~~~~calculates $\trans\x\hat W$
+%\State All machines broadcast $Z^{reopt}_i$ to a master machine
+%\State The master calculates $\wreopt$
+%\end{algorithmic}
+%%\label{fig:alg}
+%\end{algorithm}
 
@@
 
-\subsection{Notation}
+In this section, we analyze OWA's statistical performance.
+The main tool we use is the Central Limit Theorem (CLT).
+Informally, this theorem states that the $\wmle_i$ parameter vectors are approximately normally distributed.
+%We then use results from the random projection literature to bound the distance
+We then combine this result with standard results in random projections.
+Classical versions of the CLT apply only in the limit as $n\to\infty$.
+But we use a more advanced version of the CLT due to \cite{spokoiny2012parametricestimation} that applies in the finite sample case.
+This approach makes our analysis simpler and more general than the analysis of algorithms discussed in Section \ref{sec:anal}.
+All the technical details are handled by the CLT.
+In particular, we prove novel bounds for the $\wave$ estimator.
+
+\subsection {Subgaussian}
+
+%We say that a random vector $X$ is \emph{subgaussian} with mean $\mu$ and covariance $\Sigma$
+%if there exists a constant $c>0$ such that
+%\begin{equation}
+%\prob{\ltwo{\Sigma(X-\mu)} > t} \le c \exp(-t^2)
+%\end{equation}
+%The simplest example of a subgaussian random vector is a gaussian random vector.
+%We denote this relationship in symbols as $X \lesssim \normal{\mu}{\Sigma}$.
+
+We say that a random vector $\x\in\mathbb{R}^d$ is \emph{subgaussian} with variance proxy $\sigma>0$ if for all $\alpha\in\mathbb{R}^d$,
+\begin{equation}
+\E\exp(\trans\alpha\x) \le \exp(\ltwo{\alpha}^2\sigma^2/2)
+.
+\end{equation}
+The simplest example of a subgaussian random variable is a multivariate normal vector with mean $\zero$ and covariance $I$.
+In this case, $\sigma=1$.
+
+\subsection{The Central Limit Theorem}
+
+For completeness,
+we state in this section a simplified version of Theorem 5.2 in \cite{spokoiny2012parametricestimation}.
+We begin by stating the theorem's assumptions.
+These assumptions are general.
+We note in particular that logistic regression satisfies these assumptions,
+and that these assumptions do not require the likelihood function be convex.
+
+%\begin{enumerate}
+%\item
+\begin{assumption}
+The likelihood function $f(y,\trans\x\w)$ is two times differentiable with respect to $\w$.
+\end{assumption}
+%\item
+\begin{assumption}
+For all $u>0$, there exists a function $g_1(u)>0$ such that for all $|\lambda| \le g_1(u)$
+and for all $\x\in\X$,
+\begin{equation}
+\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} \le u}
+\log\E\exp\left(\lambda\frac{\trans\x\nabla_\w f(y_i,\trans\x\w)}{\ltwo{V\x}}\right)
+\le
+\frac{v_0^2\lambda^2}{2}
+\end{equation}
+where $\trans V V = \I_\wstar$ and $v_0$ is a constant.
+\end{assumption}
+%\item
+\begin{assumption}
+For all $u>0$, there exists a function $b(u)>0$ such that
+%\begin{equation}
+%\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} = u}
+%\log\E\exp\left(\frac{\E\left(f(y,\trans\x\w) - f(y,\trans\x\wstar)\right)}{\ltwo{\I_\wstar(\w-\wstar)}}\right)
+%\le
+%b(u)
+%\end{equation}
+\begin{equation}
+\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} = u}
+\log\E\exp\left(\frac{k(\w,\wstar)}{\ltwo{V(\w-\wstar)}}\right)
+\le
+b(u)
+\end{equation}
+where
+%\begin{equation}
+$
+k(\w,\wstar) = \E\left(f(y,\trans\x\w) - f(y,\trans\x\wstar)\right)
+$
+.
+%\end{equation}
+\end{assumption}
+%\end{enumerate}
+
+\begin{theorem}
+Assume Conditions 1-3 hold.
+Then with probability at least $1-\delta$,
+\begin{equation}
+\ltwo{V(\w-\wstar)} \le \sqrt{\frac{d+\log{1/\delta}}{n}}
+\end{equation}
+\end{theorem}
 
@@
 %
-%We say that a random vector $X$ is \emph{sub-Gaussian} with mean $\mu$ and covariance $\Sigma$
-%if there exists a constant $c>0$ such that
-%\begin{equation}
-%\prob{\ltwo{\Sigma(X-\mu)} > t} \le c \exp(-t^2)
-%\end{equation}
-%We denote this relationship in symbols as $X \lesssim \normal{\mu}{\Sigma}$.
-%
 %\begin{theorem}
@@
 \item the matrix $\I_{\wtave}$ is non-singular;
-\item the support set $A = \{(y,\x) : f(y,\x;\theta) > 0\}$ is independent of $\theta$;
-and \item for all $(y,\x)\in A$, $f(y,\x;\theta)$ is three times differentiable with respect to $\theta$,
+\item the support set $A = \{(y,\x) : f(y,\trans\x\theta) > 0\}$ is independent of $\theta$;
+and \item for all $(y,\x)\in A$, $f(y,\trans\x\theta)$ is three times differentiable with respect to $\theta$,
 and the third derivative is continuous.
@@
 There are many generalizations of Theorem \ref{thm:clt}.
-For example, Theorem 4.2 in \cite{spokoiny2012parametric} provides finite sample bounds with explicit dependence on the degree of model misspecification, optimization error, and the lack of independence in the data variables.
-Because Spokoiny's theorem is non-asymptotic, the theorem that the estimator is a sub-Gaussian random variable.
+For example, Theorem 4.2 in \cite{spokoiny2012parametricestimation} provides finite sample bounds with explicit dependence on the degree of model misspecification, optimization error, and the lack of independence in the data variables.
+Because Spokoiny's theorem is non-asymptotic,
+the distribution of the estimator is a sub-Gaussian random variable.
+This more general setting requires more extensive notation,
+but all of our results hold.
 
@@
 By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
-we can show that the variance of the MLE trained on all of the data is
+we can show that the variance of the maximum likelihood estimator trained on all of the data is
 \begin{equation}
 \wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
-\end{equation}
-The only difference is the location of the Fisher information matrix.
+\label{eq:var3}
+\end{equation}
+The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.
 
@@
 \\
-\Waff_{\E}&=\affspan \{\wmle_i - \wave + \E\wave\}_{i=1}^m
+\WaffE&=\affspan \{\wmle_i - \wave + \E\wave\}_{i=1}^m
 \end{align}
@@
 &\le
-\ltwo{\wstar-\proj{\Waff_\E}\wstar}
+\ltwo{\wstar-\proj\WaffE\wstar}
 %\\&
 +
-\ltwo{\proj{\Waff_\E}\wstar-\proj\Waff\wstar}
+\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
 %\\&+
@@
 and the second by the triangle inequality.
-
-%Notice that $\Waff\subseteq\W$, so $\ltwo{\wstar-\proj\W\wstar} \le \ltwo{\wstar-\proj\Waff\wstar}$.
-%\begin{equation}
-%\Wtave=\{\x - \wave + \wtave : \x\in\Wave\}
-%\end{equation}
-%We have by the triangle inequality that
-%\begin{equation}
-%\begin{split}
-%\ltwo{\wstar-\proj\Wave\wstar}
-%&\le
-%\ltwo{\wstar-\proj\Wtave\wstar}
-%\\&+
-%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
-%\end{split}
-%\end{equation}
-The left term above can be thought of as akin to the bias of the estimator $\proj\Wtave\wstar$,
+The left term of the inequality above can be thought of as akin to the bias of the estimator $\proj\Waff\wstar$,
 and the right term can be thought of as akin to the variance.
@@
 \begin{align}
-&\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
+&\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
 \\
 &=
-\ltwo{(\proj\Wave\wstar+\wave-\wtave)-\proj\Wave\wstar}
+\ltwo{(\proj\Waff\wstar+\wave-\wtave)-\proj\Waff\wstar}
 \\&=
@@
 \end{align}
-The first equality follows because the space $\Wtave$ is a translation of $\Wave$ by the vector $(\wave-\wtave)$.
+The first equality follows because the space $\WaffE$ is a translation of $\Waff$ by the vector $(\wave-\wtave)$.
 The probabilistic inequality is due to Equation \ref{eq:ct1} in the proof of Theorem \ref{thm:wave}.
@@
 We now bound the left term.
-We have that
+Define the vector space
+\begin{equation}
+\V=\vecspan \{\wmle_i - \wtave\}_{i=1}^m
+\end{equation}
+which is the translation of the affine space $\WaffE$ to include the origin.
+Then we have that
 %Note that $\proj\Wtave$ is not a linear projection but an affine projection.
@@
 \begin{align}
-\ltwo{\wstar-\proj\Wtave\wstar}
+\ltwo{\wstar-\proj\WaffE\wstar}
 &=
-\ltwo{(\wstar-\wtave)-\proj\W(\wstar-\wtave)}
+\ltwo{(\wstar-\wtave)-\proj\V(\wstar-\wtave)}
 \\&=
-\ltwo{(I-\proj\W)(\wstar-\wtave)}
+\ltwo{(I-\proj\V)(\wstar-\wtave)}
 \end{align}
@@
 
+We test our parallelization procedure on two logistic regression tasks.
+The first task uses synthetic data.
+The second task uses real world add click data from the tencent search engine.
+
 \subsection{Synthetic Data}
+
+\begin{figure*}[t]
+%\begin{tikzpicture}
+\begin{tabular}{ccc}
+  \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-100-star.pdf}
+& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-1000-star.pdf}
+& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-10000-star.pdf}
+\\
+  \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-uniform-log-1000-100-star.pdf}
+& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-uniform-log-1000-1000-star.pdf}
+& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-uniform-log-1000-10000-star.pdf}
+\\
+  \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-100-star.pdf}
+& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-1000-star.pdf}
+& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-10000-star.pdf}
+\\
+  \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-spike-log-1000-100-star.pdf}
+& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-spike-log-1000-1000-star.pdf}
+& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-spike-log-1000-10000-star.pdf}
+\end{tabular}
+%\end{tikzpicture}
+\caption{}
+\end{figure*}
+
+\begin{figure}
+%\begin{tikzpicture}
+\begin{tabular}{c}
+\\ \includegraphics[width=0.45\textwidth]{img/logl2-star-logl2-auto-normal-log-1000-100-100.pdf}
+%\\ \includegraphics[width=0.45\textwidth]{img/logl2-star-logl2-auto-uniform-log-1000-100-100.pdf}
+\\ \includegraphics[width=0.45\textwidth]{img/logl1-star-logl2-auto-laplace-log-1000-100-100.pdf}
+\\ \includegraphics[width=0.45\textwidth]{img/logl1-star-logl2-auto-spike-log-1000-100-100.pdf}
+%\\ \includegraphics[width=0.45\textwidth]{img/logl2-star-logl2-auto-normal-log-1000-1000-100.pdf}
+%\\ \includegraphics[width=0.45\textwidth]{img/logl2-star-logl2-auto-normal-1000-100-100.pdf}
+\end{tabular}
+%\end{tikzpicture}
+\caption{
+    A similar effect happens when $d=1000$ or $d=10000$,
+    but it is much less pronounced.
+}
+\end{figure}
+
+We consider two logistic regression models.
+In the first,
+each component of the parameter vector $\theta_i$ is sampled i.i.d. from the standard normal distribution.
+This is the Bayesian interpretation of $L2$ regularization.
+In the second, $\theta_i$ is sampled i.i.d. from the Laplace (two-sided exponential) distribution with scale 1.
+This is the Bayesian interpretation of $L1$ regularization.
+In both models, the data points are then sampled as
+\begin{equation}
+\begin{aligned}
+%\theta &\sim \normal{0}{I}
+%\\
+\x_i &\sim \normal{0}{I}
+,
+~~~
+%y_i &= \frac{1}{1+\exp(-\trans\x_i\theta)}
+y_i = \left(1+\exp(-\trans\x_i\theta)\right)^{-1}
+\end{aligned}
+\end{equation}
+
+Our first experiment shows how the parallel algorithms scale as the number of machines $m$ increases.
+We fix $n=1000$ data points per machine, so the size of the dataset $mn$ grows as we add more machines.
+This simulates the typical ``big data'' scenario where data is abundant,
+but processing resources are scarce.
+We further consider three regimes:
+%high dimensional ($d=10000$), medium dimensional ($d=1000$), and low dimensional ($d=100$) data.
+more data than dimensions ($d=100$),
+one data point per dimension ($d=1000$),
+and less data than dimensions ($d=10000$).
+
+%The advantage of artificial data is that we know the true model parameters and can evaluate our estimators' performance using the mean squared error.
+%We test against two GLMs: an ordinary least squares regression and a logistic regression.
+%In both experiments, we pick a constant dimension $d$ and a constant number of data points per machine $n$, then vary the number of machines $m$.
+%In total, there are $mn$ data points for each trial.
+%Each trial is repeated 100 times, and the reported results are the average.
+%
+%To evaluate the performance of our parallelization procedure, we evaluate against 4 baseline estimators.
+%The first (Naive) is the performance of just a single machine on $n$ data points without merging the results.
+%The second (Oracle) is the performance using all $mn$ data points on a single machine.
+%For large values of $n$ or $m$ this would of course be impractical to calculate.
+%The third (Ave) is the average of all the parameter vectors from each machine.
+%The fourth (BootstrapAve) is evaluated according to the following procedure.
+%Each machine $i$ calculates a parameter estimate $\theta_i$ using all the $n$ data points it has available,
+%and a second estimate $\theta_i'$ using only $rn$ data points (where $r$ is a constant between 0 and 1).
+%The final estimator is given by:
+%
+%$$
+%\frac{\theta_i - r\theta_i'}{1-r}
+%$$
+%
+%The idea of this procedure is to estimate the bias of the model using the smaller sample, and return a bias corrected estimate of the final procedure.
+%In practice, the value of the $r$ parameter greatly affects the estimator's performance, and this parameter must be set by cross-validation.
+%In our experiments, we use each of the values in the set $\{0.5,0.4,0.3,0.2,0.1,0.05,0.02,0.01,0.05\}$ and report only the best possible $r$ value.
+%Our method has no similar parameters to tune, and tends to perform better than BootstrapAve even when the latter uses the optimal $r$ parameter.
+%
+%\subsubsection{Data from an ordinary leasts squares model}
+%The true model parameters $\theta$ were generated according to one of three priors:
+%the uniform distribution over $[0,1]$,
+%the standard normal distribution,
+%and the standard laplace distribution.
+%Our data set $X$ was then generated from a standard normal distribution,
+%and the corresponding $Y$ values sampled from $X^T\theta + h(X,\theta) + \epsilon$,
+%where $\epsilon$ is the standard normal distribution.
+%We use three different values of $h$ to evaluate the effect of incorrect modeling assumptions on our estimators.
+%The function $h(X,\theta)=0$ tests the case when our model assumptions are correct.
+%In this case, the OLS estimator is unbiased.
+%%The function $h(X,\theta)=
 

### 2016-09-23 paper.tex
non-asymptotic

@@
 \DeclareMathOperator*{\affspan}{aff}
+\DeclareMathOperator*{\subG}{subG}
 \DeclareMathOperator*{\tr}{tr}
@@
 \newcommand{\W}{{\hat \Theta^{\textit{owa}}}}
+\newcommand{\Wowa}{{\hat \Theta^{\textit{owa}}}}
 \newcommand{\Waff}{\mathcal{\hat A}}
@@
 \newcommand{\E}{\mathbb{E}}
+\newcommand{\vv}{\mathbf{v}}
 \newcommand{\x}{\mathbf{x}}
@@
 \newcommand{\afull}{\ahat^{\textit{full}}}
-\newcommand{\wreopt}{\hat\w^{owa}}
-\newcommand{\wreoptfull}{\hat\w^{\textit{owa,full}}}
+\newcommand{\wowa}{\hat\w^{owa}}
+\newcommand{\wowafull}{\hat\w^{\textit{owa,full}}}
+\newcommand{\wowastar}{\hat\w^{\textit{owa,*}}}
 \newcommand{\wave}{\hat\w^{ave}}
@@
 \newcommand{\normal}[2]{\ensuremath{\mathcal{N}\left({{#1}},{{#2}}\right)}}
+\newcommand{\subnormal}[1]{\ensuremath{\subG\left({{#1}}\right)}}
 \newcommand{\trans}[1]{\ensuremath{{#1}^{\mathsf{T}}}}
-\newcommand{\ltwo}[1]{{\left\lVert {#1} \right\rVert}}
+\newcommand{\ltwo}[1]{{\lVert {#1} \rVert}}
+\newcommand{\ltwobig}[1]{{\left\lVert {#1} \right\rVert}}
 \newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
@@
   year={2010}
+}
+
+@incollection{vershynin2010introduction,
+  title={Introduction to the non-asymptotic analysis of random matrices},
+  author={Vershynin, Roman},
+  journal={arXiv preprint arXiv:1011.3027},
+  chapter=5,
+  booktitle={Compressed Sensing, Theory and Applications},
+  editor={Y. Eldar and G. Kutyniok},
+  year={2012}
 }
@@
 Our analysis is more general than the analysis of similar communication efficient algorithms that also reduce the bias.
-Notably, we do not require the likelihood function to be convex.
+Notably, we do not assume (as do all previous analyses) that the likelihood function is convex or that any random quantities are bounded.
 In practice,
-this bias reduction makes our algorithm significantly less sensitive to the amount of regularization.
-This has the practical benefit of greatly speeding up the model selection process.
+this bias reduction makes our algorithm more robust to the choice of regularization parameter,
+speeding up the model selection process.
 \end{abstract}
@@
 Recent work (discussed in Section \ref{sec:bounds}) has shown that no non-interactive algorithm can achieve regret bounds comparable to an interactive one.
-The popular large scale learning system Vowpal Wabbit cleverly uses both types of algorithms to achieve good real world performance.
+The popular large scale learning system Vowpal Wabbit cleverly combines both types of algorithms to achieve good real world performance.
 A non-interactive algorithm is run first to provide a good warm start to a subsequent interactive algorithm \citep{vw,agarwal2014reliable}.
@@
 
-In this paper, we propose a \emph{semi-interactive} distributed algorithm called XXX.
+In this paper, we propose a \emph{semi-interactive} distributed algorithm called \emph{optimal weighted averaging} (OWA).
 Our algorithm performs two rounds of communication,
@@
 \begin{equation}
-\wreoptfull = \hat W \afull
+\wowafull = \hat W \afull
 \end{equation}
@@
 %This is the maximum likelihood estimator when we restrict our
-Notice that $\wreoptfull$ is the maximum likelihood estimator when the parameter space $\Theta$ is restricted $\W = \vecspan \{\wmle_i\}_{i=1}^m$.
-Figure \ref{fig:contour} shows graphically that no other estimator in $\W$ can have higher empirical likelihood than $\wreoptfull$.
+Notice that $\wowafull$ is the maximum likelihood estimator when the parameter space $\Theta$ is restricted $\W = \vecspan \{\wmle_i\}_{i=1}^m$.
+Figure \ref{fig:contour} shows graphically that no other estimator in $\W$ can have higher empirical likelihood than $\wowafull$.
 
@@
 \begin{equation}
-\wreopt = \hat W \ahat
+\wowa = \hat W \ahat
 \end{equation}
@@
 
-Algorithm \ref{fig:alg2} shows the steps for efficiently computing $\wreopt$.
+Algorithm \ref{fig:alg2} shows the steps for efficiently computing $\wowa$.
 In the first round, each machine calculates $\wmle_i$ independently and broadcasts the results to every other machine.
@@
 \node[dot] at (-1.35,2.03) {};
-\node at (-0.9,2.4) {$\wreoptfull$};
+\node at (-0.9,2.4) {$\wowafull$};
 
@@
     Our method performs a second round of optimization to find the best parameter vector in $\W$.
-    Since $\wave\in\W$, the estimator $\wreoptfull$ has higher empirical likelihood than $\wave$ almost surely.
+    Since $\wave\in\W$, the estimator $\wowafull$ has higher empirical likelihood than $\wave$ almost surely.
 }
@@
 \State All machines broadcast $\Zproj_i$ to a master machine
-\State The master calculates $\wreopt$
+\State The master calculates $\wowa$
 \end{algorithmic}
@@
 %\State All machines broadcast $Z^{reopt}_i$ to a master machine
-%\State The master calculates $\wreopt$
+%\State The master calculates $\wowa$
 %\end{algorithmic}
@@
 Second, it is statistically optimal for the class of non-interactive optimization methods.
-(We show in the next section that their bound does not apply to our $\wreopt$ estimator due to our semi-interactive setting.)
+(We show in the next section that their bound does not apply to our $\wowa$ estimator due to our semi-interactive setting.)
 The main downside of the KL-average is that the minimization has a prohibitively high computational cost.
@@
 
-\subsection {Subgaussian}
-
-%We say that a random vector $X$ is \emph{subgaussian} with mean $\mu$ and covariance $\Sigma$
-%if there exists a constant $c>0$ such that
-%\begin{equation}
-%\prob{\ltwo{\Sigma(X-\mu)} > t} \le c \exp(-t^2)
-%\end{equation}
-%The simplest example of a subgaussian random vector is a gaussian random vector.
-%We denote this relationship in symbols as $X \lesssim \normal{\mu}{\Sigma}$.
-
-We say that a random vector $\x\in\mathbb{R}^d$ is \emph{subgaussian} with variance proxy $\sigma>0$ if for all $\alpha\in\mathbb{R}^d$,
-\begin{equation}
-\E\exp(\trans\alpha\x) \le \exp(\ltwo{\alpha}^2\sigma^2/2)
-.
-\end{equation}
-The simplest example of a subgaussian random variable is a multivariate normal vector with mean $\zero$ and covariance $I$.
-In this case, $\sigma=1$.
-
-\subsection{The Central Limit Theorem}
-
-For completeness,
-we state in this section a simplified version of Theorem 5.2 in \cite{spokoiny2012parametricestimation}.
-We begin by stating the theorem's assumptions.
-These assumptions are general.
-We note in particular that logistic regression satisfies these assumptions,
-and that these assumptions do not require the likelihood function be convex.
-
-%\begin{enumerate}
-%\item
-\begin{assumption}
-The likelihood function $f(y,\trans\x\w)$ is two times differentiable with respect to $\w$.
-\end{assumption}
-%\item
-\begin{assumption}
-For all $u>0$, there exists a function $g_1(u)>0$ such that for all $|\lambda| \le g_1(u)$
-and for all $\x\in\X$,
-\begin{equation}
-\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} \le u}
-\log\E\exp\left(\lambda\frac{\trans\x\nabla_\w f(y_i,\trans\x\w)}{\ltwo{V\x}}\right)
-\le
-\frac{v_0^2\lambda^2}{2}
-\end{equation}
-where $\trans V V = \I_\wstar$ and $v_0$ is a constant.
-\end{assumption}
-%\item
-\begin{assumption}
-For all $u>0$, there exists a function $b(u)>0$ such that
-%\begin{equation}
-%\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} = u}
-%\log\E\exp\left(\frac{\E\left(f(y,\trans\x\w) - f(y,\trans\x\wstar)\right)}{\ltwo{\I_\wstar(\w-\wstar)}}\right)
-%\le
-%b(u)
-%\end{equation}
-\begin{equation}
-\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} = u}
-\log\E\exp\left(\frac{k(\w,\wstar)}{\ltwo{V(\w-\wstar)}}\right)
-\le
-b(u)
-\end{equation}
-where
-%\begin{equation}
-$
-k(\w,\wstar) = \E\left(f(y,\trans\x\w) - f(y,\trans\x\wstar)\right)
-$
-.
-%\end{equation}
-\end{assumption}
-%\end{enumerate}
-
-\begin{theorem}
-Assume Conditions 1-3 hold.
-Then with probability at least $1-\delta$,
-\begin{equation}
-\ltwo{V(\w-\wstar)} \le \sqrt{\frac{d+\log{1/\delta}}{n}}
-\end{equation}
-\end{theorem}
-
-\subsection{Few data points are needed}
-\label{sec:vcdim}
-
-Define the space
-$\Wave=\affspan\{\wave_i\}_{i=1}^m\cup\{\wave\}$.
-Define $L : \mathbb{R} \to \mathbb{R}$ to be the 0-1 loss over the true data distribution.
-
-\begin{theorem}
-Let $\w^\Wave$ be an arbitrary point in $\Wave$ and $\delta\in(0,1)$.
-Then with probability at least $1-\delta$,
-\begin{equation}
-L(\wreopt) \le L(\w^\Wave) + O\left(\sqrt{\frac{m + \log(1/\delta)}{n^{reopt}}}\right)
-\end{equation}
-\end{theorem}
-\begin{proof}
-For each machine $i$, fix the set of data points $Z_i$.
-This in turn fixes the $\wmle_i$ and the space $\Wave$.
-The hypothesis class defined by $\Wave$ has VC dimension $m$ because it is linear with $m$ parameters.
-Since each of the data points in $Z^{reopt}$ is independent of $\Wave$,
-the result follows immediately from standard VC theory.
-See for example Chapter 6 of \cite{shalev2014understanding}.
-%
-%\fixme{
-    %Need to rework the definitions in Section 1 so that the data points in the second optimization don't depend on the first optimization.
-%}
-\end{proof}
-
-\subsection{In the Normal Regime}
-%\label{sec:normal}
-%
-%\begin{theorem}
-%We have that with probability at least
-%\begin{equation}
-%\wmle-\wstar \lesssim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
-%\end{equation}
-%In the limit as $n\to\infty$, we have the stronger statement that
-%\begin{equation}
-%\wmle-\wstar \sim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
-%\end{equation}
-%\end{theorem}
-
-\subsubsection{Assumptions}
-
-We assume that for every estimator $\hat\w$ trained on $n$ data points,
-\begin{equation}
-\label{eq:assumption}
-\hat\w - \E\hat\w \sim \frac{1}{\sqrt n}\normal{\zero}{\I^{-1}_{\E\hat\w}}
-%\hat\w \sim \normal{\E\hat\w}{\frac{1}{n}I^{-1}_{\E\hat\w}}
-\end{equation}
-%
-%Throughout Section \ref{sec:normal} we assume that $\wave_i$ is normally distributed with mean $\wtave$ and covariance $\frac{1}{n}I^{-1}_{\wtave}$.
-%\begin{equation}
-%\label{eq:assumption}
-%\wmle_i \sim \normal{\E\wmle_i}{\frac{1}{n}I^{-1}_{\E\wmle_i}}
-%\end{equation}
-%Here, $I_{\wtave}$ is the Fisher information matrix at point $\wtave$.
-Here, $\I_{\E\hat\w}$ is the Fisher information matrix for parameter $\E\hat\w$.
-This assumption is motivated by the central limit theorem for maximum likelihood estimation.
-%There are many versions of this theorem.
-Theorem 7.5.2 of \cite{lehmann1999elements} is particularly simple version and reproduced below.
-\begin{theorem}
-\label{thm:clt}
-Assume the following mild regularity conditions:
-\begin{enumerate}[noitemsep,topsep=0pt]%,label=(\itshape\roman*)]
-\item all data points in $Z_i$ are i.i.d.;
-\item the parameter space $\Theta$ is open;
-\item the model is identifiable and consistent;
-\item the matrix $\I_{\wtave}$ is non-singular;
-\item the support set $A = \{(y,\x) : f(y,\trans\x\theta) > 0\}$ is independent of $\theta$;
-and \item for all $(y,\x)\in A$, $f(y,\trans\x\theta)$ is three times differentiable with respect to $\theta$,
-and the third derivative is continuous.
-\end{enumerate}
-Then in the limit as $n\to\infty$,
-\begin{equation*}
-\wave_i \law \wtave + \frac{1}{\sqrt n} \normal{0}{\I^{-1}_\wstar}
-\end{equation*}
-Here the notation $\law$ denotes convergence in law,
-which is the standard notion of convergence for a sequence of probability distributions.
-\end{theorem}
-
-Note that Theorem \ref{thm:clt} does not require the likelihood $f$ to be convex.
-
-There are many generalizations of Theorem \ref{thm:clt}.
-For example, Theorem 4.2 in \cite{spokoiny2012parametricestimation} provides finite sample bounds with explicit dependence on the degree of model misspecification, optimization error, and the lack of independence in the data variables.
-Because Spokoiny's theorem is non-asymptotic,
-the distribution of the estimator is a sub-Gaussian random variable.
-This more general setting requires more extensive notation,
-but all of our results hold.
-
-\subsubsection{Analyzing $\wave$}
-
 We provide a simple bound that shows that averaging does not improve the bias of our estimator,
@@
 The main idea of our technique is to show that the variance is normally distributed.
-The ideas in this proof will be used in Section \ref{sec:analreopt} where we show that the variance of $\wreopt$ is the same as $\wave$, but that the bias is reduced.
-
-\ignore{
-\begin{lemma}
-We have that
-\begin{equation}
-\E\ltwo{\wstar-\wave}^2
-\le
-\E\ltwo{\wstar-\E\wmle_i}^2
-+
-\frac{1}{m}\E\ltwo{\E\wmle_i-\wmle_i}^2
-\end{equation}
-\end{lemma}
-\begin{proof}
-It is well known that the MSE of an estimator is the sum of its bias and variance.
-That is,
-\begin{equation}
-\E\ltwo{\wstar-\wave}^2
-=
-\E\ltwo{\wstar-\E\wave}^2
-+
-\E\ltwo{\E\wave-\wave}^2
-\end{equation}
-First we consider the bias.
-By the linearity of expectation, we have that
-\begin{align}
-\E\wave
-&=
-\E\frac{1}{m}\sum_{i=1}^m\wmle_i
-%\\&=
-=
-\frac{1}{m}\sum_{i=1}^m\E\wmle_i
-=
-\E\wmle_i
-\label{eq:expwave}
-\end{align}
-and so the bias
-$
-\E\ltwo{\wstar-\E\wave}^2
-=
-\E\ltwo{\wstar-\E\wmle_i}^2
-$.
-In words, the bias of the averaging estimator is the same as the bias of the estimator on an individual machine.
-
-Next we consider the variance.
-We use the definition of $\wave$ and Equation \ref{eq:expwave} above to conclude that
-\begin{align}
-\wave-\E\wave
-&=
-\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave
-\\&=
-\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wave\right)
-\\&=
-\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)
-%\\&\sim
-%\frac{1}{m}\sum_{i=1}^m\left(\frac{1}{\sqrt n}\normal\zero{I^{-1}_{\E\wave}}\right)
-%\\&=
-%\frac{1}{\sqrt {nm}}\normal\zero{I^{-1}_{\E\wave}}
-\end{align}
-Taking the norm and squaring both sides gives
-\begin{equation}
-\ltwo{\wave-\E\wave}
-=
-\ltwo{
-\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)
-}
-\end{equation}
-\end{proof}
-}
-
+The ideas in this proof will be used in Section \ref{sec:analreopt} where we show that the variance of $\wowa$ is the same as $\wave$, but that the bias is reduced.
 \begin{theorem}
@@
 \end{theorem}
-\begin{proof}
+
+\subsection {Non-Asymptotic Central Limit Theorem}
+
+The standard central limit theorem for maximum likelihood estimation states that as the number of samples $n \to \infty$,
+%the distribution of the parameter vector approaches a Gaussian at the rate $O(n^{-1/2}$).
+\begin{equation}
+\I^{-1/2}_\wstar(\w-\wstar) \to \frac{1}{\sqrt n}\normal{\zero}{I}
+.
+\end{equation}
+See chapter 7 of \cite{lehmann1999elements} for an elementary presentation of this result.
+In the non-asymptotic case, the distribution of the parameter vector is no longer Gaussian;
+instead, it is subgaussian.
+We say a random vector $\x$ with zero mean is \emph{subgaussian} with variance proxy $\sigma^2$ if
+\begin{equation}
+\prob{\ltwo{\x} > t} \le \exp(-\sigma^2 t^2 /2)
+.
+\end{equation}
+In notation, we say
+\begin{equation}
+\x \sim \subnormal{\sigma^2}
+.
+\end{equation}
+%We say that a random vector $\x\in\mathbb{R}^d$ is \emph{subgaussian} with variance proxy $\sigma>0$ if for all $\alpha\in\mathbb{R}^d$,
+%\begin{equation}
+%\E\exp(\trans\alpha\x) \le \exp(\ltwo{\alpha}^2\sigma^2/2)
+%.
+%\end{equation}
+The simplest example of a subgaussian random variable is a multivariate normal vector with mean $\zero$ and covariance $I$.
+In this case, $\sigma=1$.
+See \cite{vershynin2010introduction} for a tutorial on subgaussian random variables,
+including equivalent definitions and important results.
+
+Before stating the theorem, we first state the theorem's three conditions.
+The first condition is straightforward,
+but next two conditions are more technical.
+Despite the technical appearance of these assumptions,
+they are highly general.
+We note in particular that any generalized linear model (e.g. least squares, Poisson, and logistic regression),
+and many models with nonconvex likelihood functions satisfy these conditions.
+We include the conditions here for completeness,
+but defer to \cite{spokoiny2012parametricestimation} for a more detailed discussion.
+
+\begin{assumption}
+The likelihood function $f(y,\trans\x\w)$ is two times differentiable with respect to $\w$.
+\end{assumption}
+\begin{assumption}
+For all $u>0$, there exists a function $g_1(u)>0$ such that for all $|\lambda| \le g_1(u)$
+and for all $\x\in\X$,
+\begin{equation}
+\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} \le u}
+\log\E\exp\left(\lambda\frac{\trans\x\nabla_\w f(y_i,\trans\x\w)}{\ltwo{V\x}}\right)
+\le
+\frac{v_0^2\lambda^2}{2}
+\end{equation}
+where $\trans V V = \I_\wstar$ and $v_0$ is a constant.
+\end{assumption}
+\begin{assumption}
+For all $u>0$, there exists a function $b(u)>0$ such that
+\begin{equation}
+\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} = u}
+\log\E\exp\left(\frac{k(\w,\wstar)}{\ltwo{V(\w-\wstar)}}\right)
+\le
+b(u)
+\end{equation}
+where
+$
+k(\w,\wstar) = \E\left(f(y,\trans\x\w) - f(y,\trans\x\wstar)\right)
+$
+.
+\end{assumption}
+
+The following is a restatement of Theorem 5.2 in \cite{spokoiny2012parametricestimation}.
+\begin{theorem}
+Assume Conditions 1-3 hold.
+Then whenever $t \le c\sqrt{p/n}$,
+we have with probability at least
+%Then with probability at least $1-\delta$,
+%\begin{equation}
+%\ltwo{V(\w-\wstar)} \le c\sqrt{\frac{d+\log{1/\delta}}{n}}
+%\end{equation}
+%where the constant $c$ depends on $v_o$, $g_1(u)$, and $b(u)$ defined above.
+%Equivalently, we have that
+%\begin{equation}
+%V(\w-\wstar) \sim \subnormal{\sigma}
+%\end{equation}
+\end{theorem}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection {Proof of Theorem 1}
+
 We have by the triangle inequality that
@@
 Substituting Equations \ref{eq:expwave} and \ref{eq:ct1} into Equation \ref{eq:biasvar} gives the stated result.
-\end{proof}
 
@@
 The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.
-
-%\begin{theorem}
-%Let $t>0$.
-%Then with probability at least $1-\exp(-t)$,
-%\begin{equation}
-%\ltwo{\wave-\wtave} \le \sqrt{\frac{v_t}{mn}}
-%\end{equation}
-%where
-%\begin{equation}
-%v_t =
-%\tr{\I^{-1}_{\wtave}}
-%+ 2\sqrt{\tr ({\I^{-2}_{\wtave}})t}
-%+ 2\ltwo{\I^{-1}_{\wtave}}t
-%\end{equation}
-%\end{theorem}
-
-%\begin{proof}
-%By our assumption above, we have that
-%\begin{equation}
-%\wave \sim \wtave+\normal{0}{\frac{1}{nm}I^{-1}_{\wtave}}
-%\end{equation}
-%which implies
-%\begin{equation}
-%\wave - \wtave \sim \normal{0}{\frac{1}{nm}I^{-1}_{\wtave}}
-%\end{equation}
-%So our problem reduces to bounding the norm of a Gaussian random variable.
-%Applying the bound in Proposition 1.1 from \cite{hsu2012tail} gives the stated result.
-%\end{proof}
-
-%\ignore{
-%We introduce the following lemma due to \cite{hsu2012tail} for bounding the norm of a Gaussian random variable.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection {Proof of Theorem 2}
+
+Using the triangle inequality, we decompose $\ltwo{\wowa-\wstar}$ into three components:
+
+\begin{equation}
+\begin{aligned}
+\ltwo{\wowa-\wstar}
+&\le
+\ltwo{\wowa-\wowastar}
+\\&
++
+\ltwo{\wowastar - \proj{\Wowa}\wstar}
+\\&
++
+\ltwo{\proj{\Wowa}\wstar - \wstar}
+\end{aligned}
+\end{equation}
+The first term $\ltwo{\wowa-\wowastar}$ captures the approximation error due to the limited number of points in the second round of optimization.
+The result follows by first bounding each of these terms individually,
+then applying the union bound.
+
+\subsubsection{Bounding the first term}
+
+Our first lemma is a direct consequence of the central limit theorem.
 %\begin{lemma}
-%\label{lemma:hsu}
-%Let $\x\in\mathbb{R}^d$ be an isotropic multivariate Gaussian random vector with zero mean.
-%Let $\trans A A = \Sigma$ be the covariance matrix of random vector $A\x$.
-%For all $t>0$,
-%\begin{equation}
-%\prob{\ltwo{A\x}^2 > \tr{\Sigma} + 2\sqrt{\tr (\Sigma^2)t} + 2\ltwo{\Sigma}t}
-%\le
-%\exp(-t)
+%With probability at least $\delta$,
+%\begin{equation}
+%\ltwo{\wowa-\wowastar} \le
 %\end{equation}
 %\end{lemma}
-%A straightforward application of this lemma gives us that with probability at least $1 - \exp(-\tvar)$,
-%\begin{equation}
-%\begin{split}
-%%\prob{
-%&\ltwo{\wave-\wtave}
-%\\
-%&\le
-%\sqrt{
-    %\frac{1}{m}
-    %\left(\tr{\I^{-1}_{\wtave}}
-        %+ 2\sqrt{\tr ({\I^{-1}_{\wtave}}^2)\tvar}
-        %+ 2\ltwo{\I^{-1}_{\wtave}}\tvar
-    %\right)
-%}
-%%}
-%\end{split}
-%\end{equation}
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Analyzing $\wreopt$}
-\label{sec:analreopt}
-
-%We now extend the analysis of the previous section to the $\wreopt$ estimator.
+
+\subsubsection{Bounding the second term}
+The next lemma is a direct consequence of assumptions.
+
+\subsubsection{Bounding the third term}
+Then we have that with probability at least $(1-\exp(-t))^2$,
+\begin{equation}
+\ltwo{\wstar-\proj\W\wstar}
+\le
+%b_t\sqrt{\left(1-\frac{m}{d}\right)}
+b_t\sqrt{\left(1-\frac{m}{d}\right)}
++
+\sqrt\frac{v_{t}}{nm}
+\end{equation}
+where
+%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
+\begin{equation}
+b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
+\end{equation}
+%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
+$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
+and $v_t$ is defined as in Theorem \ref{thm:wave}.
+
+\subsection {Analyzing $\wowa$}
+
+%We now extend the analysis of the previous section to the $\wowa$ estimator.
 %We show that its variance is bounded by the variance of $\wave$,
@@
 %
-%Recall from Section \ref{sec:vcdim} that the estimator $\wreopt$ has low error with high probability compared to any parameters in the space $\Wave$.
+%Recall from Section \ref{sec:vcdim} that the estimator $\wowa$ has low error with high probability compared to any parameters in the space $\Wave$.
 %Denote by $\proj\Wave$ the projection operator onto the space $\Wave$.
@@
 %\begin{equation}
-%\ltwo{\wstar-\wreopt} \le
+%\ltwo{\wstar-\wowa} \le
 %\end{equation}
@@
 
-\ignore{
 \newpage
-\section*{Appendix: Proof of Lemma \ref{lem:normaff}}
-
-%\cite{hsu2012tail}
-%\begin{lemma}
-%\label{lemma:hsu}
-%Let $\x\in\mathbb{R}^d$ be an isotropic multivariate Gaussian random vector with zero mean.
-%For all $t>0$,
-%\begin{equation}
-%\prob{\ltwo{A\z}^2 > \tr{\Sigma} + 2\sqrt{\tr (\Sigma^2)t} + 2\ltwo{\Sigma}t}
-%\le
-%\exp(-t)
-%\end{equation}
-%\end{lemma}
-
-Define $\V$ to be the vector space spanned by the first $m-1$ columns of $\Wave$.
-That is, $\V=\vecspan\{\wmle_i\}_{i=1}^{m-1}$.
-We have that
+\section*{Appendix: Detailed Central Limit Theorem}
+
+\begin{theorem}
+Assume () and ().
+If for all $u_0 > 0$,
 \begin{align}
-\ltwo{\w-\proj{\Wave}\w}
-&\le \ltwo{\w - \proj{\Wave}\proj{\V}\w}
+n^{1/2}u_0b & \ge 6 v_0 \sqrt{x+cp}
+\label{eq:cond1}
 \\
-%&\le \ltwo{}
-& \le \ltwo{\w - \proj{\V}\w} + \ltwo{\proj{\V}\w - \proj{\Wave}\proj{\V}\w}
+1 + \sqrt{x+cp} &\le 3b^{-1}v_0^2g_1(u_0) n^{1/2}
+\label{eq:cond2}
 \end{align}
-The first inequality follows by the definition of $\proj\Wave\w$ as the point in $\Wave$ with minimum distance to $\w$,
-and the second follows by the triangle inequality.
-%\begin{align}
-%\ltwo{\w - \proj{H_m}\w}
-%& \le \ltwo{\w - \proj{H_m}\proj{V_{m-1}}\w} & \text{by definition of $\proj{}$}\\
-%& \le \ltwo{\w - \proj{V_{m-1}}\w} + \ltwo{\proj{V_{m-1}}\w - \proj{H_m}\proj{V_{m-1}}\w} &\text{by triangle ineq.}
-%\end{align}
-We will bound each of these terms separately.
-
-The left term above can be bound directly using standard random projection techniques.
-%The following lemma is a restatement of Lemma 2.2 from \cite{dasgupta2003elementary}.
-%It has been simplified to ignore cases irrelevant for our purpose.
-%\begin{lemma}
-%\label{lem:dasgupta}
-%Let $\x\in\mathbb{R}^d$ be an arbitrary vector and $W\in\mathbb{R}^{d\times m}$ be a random matrix with entries drawn i.i.d. from the standard normal distribution.
-%Then for all $t>1$,
-%\begin{equation*}
-%\prob{\ltwo{W\x} \ge \sqrt\frac{tm}{d}} \le \exp\left(\frac{m}{2}(1-t+\ln t)\right)
-%\end{equation*}
-%%Let $\delta\in(0,1)$ and $c$ be a suitable constant.
-%%Then for all $\x\in\mathbb{R}^d$, we have
-%%\begin{equation}
-%%\prob{\ltwo{W\x} \le (1+\epsilon)\ltwo{\x}} \ge 1 - \delta
-%%\end{equation}
-%\end{lemma}
-%To apply Lemma \ref{lem:dasgupta},
-%notice that $\ltwo{\w-\proj{\V}\w} = \ltwo{(\I-\proj\V)\w}$.
-%\begin{equation}
-%\prob{\ltwo{\w-\proj{V}\w} \ge \sqrt{t\left(1-\frac{m}{d}\right)}} \le \exp\left(\frac{d-m}{2}(1-t+\ln t)\right)
-%\end{equation}
-%
-The following lemma is adapted from Theorem 3.1 in \cite{matouvsek2008variants}.
-It has been simplified to ignore cases irrelevant for our purpose.
-\begin{lemma}
-Let $a$ and $b$ be integers with $a>b$.
-Let $T\in\mathbb{R}^{a\times b}$ be a random matrix with entries drawn i.i.d. from the standard normal distribution.
-Let $\x\in\mathbb{R}^a$ be an arbitrary vector.
-Then there exists a constant $c$ such that for all $\delta\in(0,1)$,
-\begin{equation*}
-\prob{\ltwo{T\x} \le \left(\sqrt{b}+c\sqrt{\log\frac{\delta}{2}}\right)\ltwo\x} \le 1-\delta
-\end{equation*}
-\end{lemma}
-}
+then
+\begin{equation}
+\prob{\ltwo{\vv_0(\w-\wstar)} > u_0} \le \exp(-x)
+.
+\end{equation}
+\end{theorem}
+
+Manipulating Equation \ref{eq:cond1} gives
+\begin{align}
+\end{align}
 

### 2016-09-26 paper.tex
changes

@@
 \newcommand{\X}{\mathcal{X}}
+\newcommand{\matW}{\hat W}
+\newcommand{\matV}{\hat V}
 \newcommand{\W}{{\hat \Theta^{\textit{owa}}}}
@@
 
+@inproceedings{negahban2009unified,
+  title={A unified framework for high-dimensional analysis of M-estimators with decomposable regularizers},
+  author={Negahban, Sahand and Yu, Bin and Wainwright, Martin J and Ravikumar, Pradeep K},
+  booktitle={Advances in Neural Information Processing Systems},
+  pages={1348--1356},
+  year={2009}
+}
+
 @inproceedings{mcdonald2009efficient,
@@
 
+@inproceedings{sivakumar2015beyond,
+  title={Beyond Sub-Gaussian Measurements: High-Dimensional Structured Estimation with Sub-Exponential Designs},
+  author={Sivakumar, Vidyashankar and Banerjee, Arindam and Ravikumar, Pradeep K},
+  booktitle={Advances in Neural Information Processing Systems},
+  pages={2206--2214},
+  year={2015}
+}
+
 @article{han2016bootstrap,
@@
 
-\aistatstitle{Instructions for paper submissions to AISTATS 2015}
+%\aistatstitle{Instructions for paper submissions to AISTATS 2015}
+\aistatstitle{Communication-Efficient Distributed Maximum Likelihood Estimation with the Optimal Weigted Average}
 
@@
 \begin{abstract}
-We present a distributed learning algorithm with low communication complexity.
+We present a distributed algorithm for maximum likelihood estimation (MLE) with low communication complexity.
 Our algorithm is an easy to implement modification of the standard averaging method.
-In the averaging method, each machine in a cluster first solves the learning problem on a subset of the data;
+In the averaging method, each machine in a cluster first solves the MLE problem on a subset of the data;
 the resulting parameter vectors are then averaged together.
@@
 Our analysis is more general than the analysis of similar communication efficient algorithms that also reduce the bias.
-Notably, we do not assume (as do all previous analyses) that the likelihood function is convex or that any random quantities are bounded.
+Notably, we do not assume (as do all previous analyses) that the likelihood function is concave or that any quantities are bounded.
 In practice,
@@
 Recent work (discussed in Section \ref{sec:bounds}) has shown that no non-interactive algorithm can achieve regret bounds comparable to an interactive one.
-The popular large scale learning system Vowpal Wabbit cleverly combines both types of algorithms to achieve good real world performance.
+The popular large scale learning system Vowpal Wabbit combines both types of algorithms to achieve good real world performance.
 A non-interactive algorithm is run first to provide a good warm start to a subsequent interactive algorithm \citep{vw,agarwal2014reliable}.
@@
 so our algorithm retains the speed advantages of non-interactive algorithms.
-
-%In the remainder of this paper, we first formally describe our problem and proposed solution.
-%We then provide a detailed comparison between our method and existing methods.
-%We then provide a theoretical analysis.
-%We conclude with an empirical analysis.
-
-%\section{PROBLEM AND SOLUTION}
-
 We now formally describe our problem and proposed solution.
@@
 \subsection{Problem Statement}
+%\section{Problem Setting and Our Solution}
 
@@
 Our analysis in Section \ref{sec:anal} places very mild restrictions on $f$.
-In particular, $f$ need not be convex.
+In particular, $f$ need not be concave with respect to $\w$.
 %For our experiments in Section \ref{sec:exp}, however, we restrict ourselves to the likelihood for logistic regression
@@
 Let $Z\subset\X\times\Y$ be a dataset of $mn$ i.i.d. observations.
+Finally, let $R : \Theta \to \mathbb{R}$ be a regularization function (typically the L1 or L2 norm)
+and $\lambda\in\mathbb{R}$ be the regularization strength.
 Then the maximum likelihood estimator is
@@
 \wmle=\argmax_\w \sum_{(\x,y)\in Z} f(y,\trans\x\w)
++ \lambda R(\theta)
+.
 \end{equation}
@@
 \wmle_i = \argmax_\w \sum_{(\x,y) \in Z_i} f(y,\trans\x\w)
++ \lambda R(\theta)
+.
 \end{equation}
@@
 %The resulting merged estimate should have error as close to $\wmle$ as possible.
-A non-interactive distributed estimator merges these $\wmle_i$ into a single improved estimate.
-The simplest example is the averaging estimator
+%A non-interactive distributed estimator merges these $\wmle_i$ into a single improved estimate.
+Our goal is to merge the $\wmle_i$ into a single improved estimate.
+The baseline example is the averaging estimator
 \begin{equation}
 \wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
-\end{equation}
-This estimator is well studied, and in Section \ref{sec:anal} we comprae this previous work to our own.
-Here we briefly note that $\wave$ is known to reduce the variance of an estimator but not its bias.
+.
+\end{equation}
+This estimator is well studied, and in Section \ref{sec:anal} we compare this previous work to our own.
+Here we briefly recall that the analysis of an estimator's error $\ltwo{\wstar-\wave}$ is typically decomposed into bias $\ltwo{\wstar - \E\wave}$ and variance $\ltwo{\E\wave-\wave}$ components.
+The $\wave$ estimator is known to reduce the variance but not the bias.
 Our goal is to design an estimator that reduces both variance and bias.
@@
 
-In this section we present a modification to the averaging estimator called the \emph{Optimal Weighted Average} (OWA).
+We propose a modification to the averaging estimator called the \emph{optimal weighted average} (OWA) that reduces both variance and bias.
 OWA uses a second round of optimization to calculate the optimal linear combination of the $\wmle_i$s.
@@
 This second optimization occurs over a small fraction of the dataset,
-so its computational and communication cost is negligible.
-
-To motivate our estimator, we first describe a simpler estimator that uses the entire dataset for the second round of optimization.
+so its computational and communication overhead is negligible.
+
+To motivate our estimator,
+we first describe a simpler estimator that uses the entire dataset for the second round of optimization.
 %The idea of our proposed solution is to find the best linear combination of the $\wmle_i$s using a second round of optimization.
-Define the matrix $\hat W : \mathbb{R}^{d\times m}$ to have $i$th column equal to $\wmle_i$.
+Define the matrix $\matW : \mathbb{R}^{d\times m}$ to have $i$th column equal to $\wmle_i$.
 Now consider the estimator
 \begin{equation}
-\wowafull = \hat W \afull
+\wowafull = \matW \afull
+,
 \end{equation}
@@
 \begin{equation}
-\afull = \argmax_\alpha \sum _{(\x,y)\in Z} f\left(y,\trans\x \hat W \alpha \right)
+\label{eq:afull}
+\afull = \argmax_\alpha \sum _{(\x,y)\in Z} f\left(y,\trans\x \matW \alpha \right)
++
+\lambda R(\matW\alpha)
+.
 \end{equation}
 %This is the maximum likelihood estimator when we restrict our
-Notice that $\wowafull$ is the maximum likelihood estimator when the parameter space $\Theta$ is restricted $\W = \vecspan \{\wmle_i\}_{i=1}^m$.
-Figure \ref{fig:contour} shows graphically that no other estimator in $\W$ can have higher empirical likelihood than $\wowafull$.
-
-Calculating the weights $\afull$ directly is infeasible because it requires access to the full dataset.
-Fortunately, we don't need to consider all the data points for an accurate estimator.
-The parameter space $\W$ is $m$-dimensional.
-So intuitively, we only need $O(m)$ data points to solve such a problem to our desired accuracy.
-%(This intuition is formalized in Section \ref{sec:anal}.)
-This intuition motivates the following estimator.
-Let $\Zreopt_i\subset Z_i$ be a set of $\nreopt$ data points uniformly sampled from $Z_i$ without replacement,
-and let $\Zreopt$ be the union of the $\Zreopt_i$s.
-Then the OWA estimator is defined as
-\begin{equation}
-\wowa = \hat W \ahat
-\end{equation}
-where
-\begin{equation}
-\ahat = \argmax_\alpha \sum _{(\x,y)\in \Zreopt} f\left(y,\trans\x \hat W \alpha \right)
-\end{equation}
-
-Algorithm \ref{fig:alg2} shows the steps for efficiently computing $\wowa$.
-In the first round, each machine calculates $\wmle_i$ independently and broadcasts the results to every other machine.
-Since the parameter vector has $d$ dimensions and there are $m$ machines, a total of $O(md)$ bits of get transmitted.
-In the second round, each machine projects its local dataset $\Zreopt$ onto the space $\W$.
-These projected data points are then transmitted to a predesignated master machine.
-The projected data points each have dimension $m$, so a total of $O(m^2\nreopt)$ bits are transmitted.
-The master machine now has all of the information to complete the optimization.
+Notice also that $\wowafull$ is the maximum likelihood estimator when the parameter space $\Theta$ is restricted to $\W = \vecspan \{\wmle_i\}_{i=1}^m$.
+In other words, the $\afull$ vector contains the optimal weights to apply to each $\wmle_i$ when averaging.
+Figure \ref{fig:contour} shows graphically that no other estimator in $\W$ can have higher regularized empirical likelihood than $\wowafull$.
 
@@
 
-\node at (4,-1.1) {$f(y,\trans\x\w)$};
+\node[dot] at (-0.55,1.9) {};
+\node at (-0.1,2.1) {$\wowa$};
+
+\node at (4,-1.3) {$f(y,\trans\x\w)+\lambda R(\theta)$};
 \node at (5.65,1.0) {$\W$};
@@
     Our method performs a second round of optimization to find the best parameter vector in $\W$.
-    Since $\wave\in\W$, the estimator $\wowafull$ has higher empirical likelihood than $\wave$ almost surely.
+    Since $\wave\in\W$, the estimator $\wowafull$ has higher empirical likelihood than $\wave$.
+    Since $\W$ has low dimension, we can use relatively few data points in the second round of optimization to ensure that with high probability $\wowa$ has higher empirical likelihood than $\wave$.
 }
@@
 
+Calculating the weights $\afull$ directly is infeasible because it requires access to the full dataset.
+Fortunately, we don't need to consider all the data points for an accurate estimator.
+The parameter space $\W$ is $m$-dimensional.
+So intuitively, we only need $O(m)$ data points to solve the second optimization to our desired accuracy.
+%(This intuition is formalized in Section \ref{sec:anal}.)
+This intuition motivates the following estimator.
+Let $\Zreopt_i\subset Z_i$ be a set of $\nreopt$ data points uniformly sampled from $Z_i$ without replacement,
+and let $\Zreopt$ be the union of the $\Zreopt_i$s.
+Then the OWA estimator is defined as
+\begin{equation}
+\wowa = \matW \ahat,
+\end{equation}
+where
+\begin{equation}
+\label{eq:ahat}
+\ahat = \argmax_\alpha \sum _{(\x,y)\in \Zreopt} f\left(y,\trans\x \matW \alpha \right)
++ \lambda R (\matW\alpha)
+.
+\end{equation}
+
+Algorithm \ref{fig:alg2} shows the steps for efficiently computing $\wowa$.
+In the first round, each machine calculates $\wmle_i$ independently and broadcasts the result to every other machine.
+Since the parameter vector has $d$ dimensions and there are $m$ machines, a total of $O(md)$ bits are transmitted.
+In the second round, each machine projects its local dataset $\Zreopt_i$ onto the space $\W$.
+These projected data points are then transmitted to a predesignated master machine.
+The projected data points each have dimension $m$, so a total of $O(m^2\nreopt)$ bits are transmitted.
+The master machine now has all of the information to complete the optimization.
+
 \begin{algorithm}[t]
-\caption{Estimation in two rounds}
+\caption{Distributed calculation of $\wowa$}
 \label{alg:distributed}
@@
 \State Each machine $i$ independently
-\State ~~~~~constructs $\hat W$
+\State ~~~~~constructs $\matW$
 \State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
-\State ~~~~~calculates $\Zproj_i=\{(\trans\x\hat W,y) : (\x,y)\in\Zreopt\}$
+\State ~~~~~calculates $\Zproj_i=\{(\trans\x\matW,y) : (\x,y)\in\Zreopt\}$
 \State All machines broadcast $\Zproj_i$ to a master machine
@@
 
-%\begin{algorithm}[t]
-%\caption{Estimation with common data}
-%\label{alg:distributed}
-%\begin{algorithmic}
-%\State Each machine $i$ independently
-%\State ~~~~~loads dataset $Z_i$
-%\State ~~~~~calculates $\wmle_i$
-%\State All machines broadcast $\wmle_i$ to all other machines
-%\State Each machine $i$ independently
-%\State ~~~~~constructs $\hat W$
-%\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
-%\State ~~~~~calculates $\trans\x\hat W$
-%\State All machines broadcast $Z^{reopt}_i$ to a master machine
-%\State The master calculates $\wowa$
-%\end{algorithmic}
-%%\label{fig:alg}
-%\end{algorithm}
+\subsection{Implementation Tips}
+
+Equations \ref{eq:afull} and \ref{eq:ahat} cannot be solved directly using off the shelf optimizers because existing optimizers do not support the non-standard regularization term $R(\matW\alpha)$.
+In practice, it is sufficient to approximate this regularization by L2 regularization directly on the $\alpha$ vector:
+\begin{equation}
+\lambda R(\matW\alpha) \approx \lambda_2 \ltwo{\alpha}
+.
+\end{equation}
+Intuitively, this is because even when we want the parameter vector $\w$ to be sparse (and so are regularizing by $R = $ the L1 norm), we have no reason to believe that the $\alpha$ vector should be sparse.
+The desired sparsity is induced by the regularization when solving for $\wmle_i$s and maintained in any linear combination of the $\wmle_i$s.
+The new $\lambda_2$ regularization parameter should be set by cross validation.
+This will be a fast procedure, however, because there are few data points to optimize over,
+and the L2 regularized problem is much easier to solve than the L1 problem.
+The new $\lambda_2$ regularization parameter should be set by cross validation.
+This will be a fast procedure, however, because there are few data points to optimize,
+and the L2 regularized problem is much easier to solve than the L1 problem.
+With this minor modification, our distributed estimator can be implemented using any existing optimizer.
+
+It is also important when performing the second round of optimization to not extend the parameter vector with an intercept term as is typically done.
+It is fine to do this for the first round as long as the intercepts also get scaled.
 
@@
 Related work can be divided into two categories:
-alternative $q$ estimators,
+alternative estimators,
 and bounds on the communication complexity of distributed learning.
@@
 \wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
+.
 \end{equation}
@@
 \cite{mcdonald2009efficient} gave the first analysis of this estimator in the special case of L2 regularized maximum entropy models.
-They provide tail bounds on $\ltwo{\wstar-\wave}$, showing that the variance of the estimator reduces as $O((nm)^{-1/2})$ with high probability,
+They provide tail bounds on $\ltwo{\wstar-\wave}$, showing that the variance $\ltwo{\E\wave-\wave}$ reduces as $O((nm)^{-1/2})$,
 but they could not show a reduction in bias.
@@
 This is a particularly limiting assumption as even the simple case of Gaussian distributed data does not satisfy it.
-\cite{zhang2012communication} provide a more general analysis showing that the Mean Squared Error (MSE) $\E\ltwo{\wstar-\wave}^2$ decays as $O((nm)^{-1} + n^{-2})$.
+\cite{zhang2012communication} provide a more general analysis showing that the mean squared error (MSE) $\E\ltwo{\wstar-\wave}{}^2$ decays as $O((nm)^{-1} + n^{-2})$.
 This matches the optimal MSE of $\wmle$ whenever $m<n$.
@@
 They further make strong convexity and 8\emph{th} order smoothness assumptions which guarantee that $\wmle_i$ is a ``nearly unbiased estimator'' of $\wstar$.
-(As we will see in the next section,
-recently developed information theoretic bounds show these assumptions are required to show a reduction in MSE to the single machine rate.)
-Surprisingly, \cite{zhang2013divide} showed that in the special case of kernel ridge regression, a reduction in bias was not needed to have the MSE decay at the optimal sequential rate.
-By a careful choice of regularization parameter, they caused $\wmle_i$ to have lower bias but higher variance, so that the final estimate of $\wave$ had both reduced bias and variance.
+%(As we will see in the next subsection,
+%recently developed information theoretic bounds show these assumptions are required to show a reduction in MSE to the single machine rate.)
+%All previous analysis of $\wave$ assumes that the likelihood $f$ is concave.
+In Section \ref{sec:anal}, we provide a simpler analysis of $\wave$ that relaxes these assumptions.
+In particular, our analysis does not require any quantities to be bounded or the likelihood to be concave.
 
@@
 which works as follows.
-Let $r\in(0,1)$, and $X_i^r$ be a bootstrap sample of $X_i$ with $rn$ data points.
+Let $r\in(0,1)$, and $Z_i^r$ be a bootstrap sample of $Z_i$ with $rn$ data points.
 Then the bootstrap average estimator is
@@
 \begin{aligned}
-\wmler_i &= \argmax_\w \sum_{\x\in X_i^r} f(x;\w)
+\wmler_i &= \argmax_\w \sum_{(\x,y)\in Z_i^r} f(y,\trans\x\w) + \lambda R(\w)
+,
 \\
 \waver &= \frac{1}{m}\sum_{i=1}^m \wmler_i
+,
 \\
 \wboot & = \frac{\wave-r\waver}{1-r}
+.
 \end{aligned}
@@
 The intuition behind this estimator is to use the bootstrap sample to directly estimate and correct for the bias.
-This estimator enjoys a MSE decaying as $O((nm)^{-1}+n^{-3})$.
+This estimator enjoys a MSE decaying as $O((nm)^{-1}+n^{-3})$ under similar assumptions as their analysis of $\wave$.
+There are two main limitations to $\wboot$.
+First, the optimal value of $r$ is not obvious and setting the parameter requires cross validation on the entire data set.
+Our proposed $\wowa$ estimator has a similar parameter $\lambda_2$ that also needs tuning,
+but this tuning happens on a small fraction of the data and always with the L2 regularizer.
+So properly tuning $\lambda_2$ is more efficient than $r$.
+Second, performing a bootstrap on an unbiased estimator increases the variance.
+This means that $\wboot$ could perform worse than $\wave$ on unbiased estimators.
+Our $\wowa$ estimator, in contrast, will perform at least as well as $\wave$ with high probability as seen in Figure \ref{fig:contour}.
+In Section \ref{sec:exp}, we show that our estimator has better empirical performance.
 
@@
 Instead of averaging the model's parameters,
-they directly ``average the models'' with the following KL-average estimator.
+they directly ``average the models'' with the following KL-average estimator:
 \begin{equation}
 \wkl = \argmin_\w \sum_{i=1}^m \kl[\bigg]{p(\cdot;\wmle_i)}{p(\cdot;\w)}
+.
 \end{equation}
@@
 Second, it is statistically optimal for the class of non-interactive optimization methods.
-(We show in the next section that their bound does not apply to our $\wowa$ estimator due to our semi-interactive setting.)
+(We show in the next section that this optimality bound does not apply to our $\wowa$ estimator due to our semi-interactive setting.)
 The main downside of the KL-average is that the minimization has a prohibitively high computational cost.
@@
 Their experiments show the procedure scaling only to datasets of size $nm\approx10^4$,
-whereas our experiments involve a dataset of size $nm\approx10^8$
+whereas our experiments involve a dataset of size $nm\approx10^8$.
+
+Surprisingly, \cite{zhang2013divide} showed that in the special case of kernel ridge regression,
+a reduction in bias was not needed to have the MSE of $\wave$ decay at the optimal sequential rate.
+By a careful choice of regularization parameter,
+they caused $\wmle_i$ to have lower bias but higher variance,
+so that the final estimate of $\wave$ had both reduced bias and variance.
+%This suggests that once the proper regularization parameter is known,
+%there is no need for a bias reduction at all.
+This suggests that a merging procedure that reduces bias is not crucial to good performance if we set the regularization parameter correctly.
+Typically there is a narrow range of good regularization parameters, and finding a $\lambda$ in this range is expensive computationally.
+We show experimentally in Section \ref{sec:exp} that our method has significantly reduced sensitivity to the regularization parameter.
+We consider this to be the main practical strength of our method.
 
@@
 Performance bounds come in two flavors: statistical and information theoretic.
-On the statistical side, \cite{liu2014distributed} showed that for an arbitrary function $q$ \emph{which does not depend on the data},
-$\ltwo{\wq-\wmle}^2$ decays as $\Omega(\gamma^2_\wstar \I^{-1}_\wstar/n^2)$.
+On the statistical side, \cite{liu2014distributed} showed that for any non-interactive distributed estimator $\wq$,
+the quantity $\ltwo{\wq-\wmle}{}^2$ decays as $\Omega(\gamma^2_\wstar \I^{-1}_\wstar/n^2)$.
 Here $\gamma_\wstar$ is the statistical curvature of the model and $\I_\wstar$ is the Fisher information.
 Furthermore, they show that $\wkl$ matches this bound.
-This bound is not relevant for our estimator because our merge function depends on the data.
-Furthermore, we directly optimize $\ltwo{\wq-\wstar}^2$ which is a more useful quantity in practice.
+This bound is not relevant for our estimator because of our semi-interactive setting.
+A crucial assumption of Liu and Ihler's analysis is that the merge function not depend on the data as our merge function does.
+%Furthermore, we directly optimize $\ltwo{\wq-\wstar}^2$ which is a more useful quantity in practice.
 
@@
 
-In this section, we analyze OWA's statistical performance.
-The main tool we use is the Central Limit Theorem (CLT).
-Informally, this theorem states that the $\wmle_i$ parameter vectors are approximately normally distributed.
-%We then use results from the random projection literature to bound the distance
-We then combine this result with standard results in random projections.
-Classical versions of the CLT apply only in the limit as $n\to\infty$.
-But we use a more advanced version of the CLT due to \cite{spokoiny2012parametricestimation} that applies in the finite sample case.
-This approach makes our analysis simpler and more general than the analysis of algorithms discussed in Section \ref{sec:anal}.
-All the technical details are handled by the CLT.
-In particular, we prove novel bounds for the $\wave$ estimator.
-
-We provide a simple bound that shows that averaging does not improve the bias of our estimator,
-but reduces the variance at the rate $O((nm)^{-1})$.
-Similar bounds are well known (see Section \ref{sec:alt}).
-Our analysis has the following advantages:
+In this section, we analyze the statistical performance of our $\wowa$ estimator.
+As part of the analysis, we provide a novel proof of the statistical performance of $\wave$ with an easier to interpret constant factor.
+Our analysis is simpler and more general than previous analyses of distributed estimators.
+In particular, we do not assume that any random variables are bounded
+or that our likelihood function is concave.
+
+\subsection{Conditions}
+
+The only condition of our analysis is that the estimation error has sub-Gaussian tails.
+We first state this condition formally,
+then explain why this condition is mild.
+
+%That is, we assume there exists a constant $c$ such that
+%\begin{equation}
+%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{\frac{d}{n}}} \le \exp\left(-ct^2\right)
+%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{d/n}} \le \exp\left(-ct^2\right)
+%%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(-\sigma^2t^2/2\right)
+%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(\frac{-\sigma^2t^2}{2}\right)
+%.
+%\end{equation}
+\begin{assumption}
+Let $t>0$ and $\sigma^2$ be a constant called the \emph{variance proxy}.
+Then with probability at least $1-\exp(-\sigma^2t^2/2)$,
+\begin{equation}
+%\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} < \frac{t}{\sqrt n}
+\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < \frac{t}{\sqrt n}
+,
+\end{equation}
+where $\I_\wstar$ is the Fisher information matrix.
+\end{assumption}
+
+This is a mild condition that is known to hold in many situations of interest.
+In the asymptotic regime when $n\to\infty$,
+%this condition is known as root-$n$ consistency.
+%In this case,
+the variance proxy $\sigma^2=1$,
+and the distribution of the estimation error is truly Gaussian.
+Theorem 7.5.2 of \cite{lehmann1999elements} provides a particularly simple exposition.
+Lehman's theorem requires only that $f$ be three times differentiable,
+the data points be i.i.d.,
+and some mild identifiability conditions.
+Very strong results are known in the asymptotic regime.
+For example, Condition 1 is satisfied even without the i.i.d. assumption.
+See for example \cite{spokoiny2012parametricestimation}.
+
+Similar results hold in the non-asymptotic case $n<\infty$,
+but known results are less strong.
+Most non-asymptotic results of this form require that the data points be sub-Gaussian.
+For example, \cite{negahban2009unified} considers the case when the data points are sub-Gaussian, the likelihood satisfies a ``restricted strong convexity condition,'' and the regularizer is decomposable.
+More recently, \cite{sivakumar2015beyond} showed Condition 1 holds when the data are only sub-exponential.
+The strongest non-asymptotic results on the estimation error known to the authors is due to \cite{spokoiny2012parametricestimation}.
+Spokoiny requires only a mild exponential moment condition (and does not even require the data be i.i.d.);
+but can show that Condition 1 is satisfied only for $t < O(n)$.
+
+Our condition is more general than the conditions used in previous work.
+\cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded,
+and every bounded random variable is automatically sub-Gaussian by definition.
+
+In conclusion,
+there are many ways to verify that Condition 1 holds.
+Pick whichever one suits your problem.
+With this condition assumed, our results are elementary.
+
+\subsection{Analysis of $\wave$}
+
+%The main tool we use is the central limit theorem (CLT) for maximum likelihood estimators.
+%%We then combine this result with standard results in random projections.
+%There are many versions of this theorem,
+%but they all conclude that the variance of an estimator is normally distributed.
+%In particular,
+%\begin{equation}
+%\label{eq:assumption}
+%\E\wmle_i - \wmle_i \sim \normal{\zero}{\I^{-1}_{\E\wmle_i}}
+%\end{equation}
+%where $\I^{-1}_{\E\wmle_i}$ is the Fisher information matrix.
+%The classical version of the CLT applies only as the number of samples $n\to\infty$.
+%Non-classical versions handle the non-asymptotic case.
+%Section \ref{sec:clt} below provides an overview.
+
+%We begin by presenting our theorem on the statistical performance of $\wave$.
+We provide a simple bound that shows that averaging improves the variance,
+but not the bias of an estimator.
+Similar bounds are well known (see Section \ref{sec:alt}),
+but our analysis has the following advantages:
 it requires fewer assumptions, has a simpler proof, and has an easy to interpret constant factor.
-The main idea of our technique is to show that the variance is normally distributed.
-The ideas in this proof will be used in Section \ref{sec:analreopt} where we show that the variance of $\wowa$ is the same as $\wave$, but that the bias is reduced.
+%The main idea of our technique is to show that the variance is normally distributed.
+%The ideas in this proof will be used in Section \ref{sec:analreopt} where we show that the variance of $\wowa$ is the same as $\wave$, but that the bias is reduced.
+
 \begin{theorem}
+\label{thm:wave}
+Assume Condition 1 holds.
 Let $t>0$.
@@
 v_t =
-\tr{\I^{-1}_{\wtave}}
-+ 2\sqrt{\tr \left({\I^{-2}_{\wtave}}\right)t}
-+ 2\ltwo{\I^{-1}_{\wtave}}t
+\sigma^2
+\left(
+\tr{\I^{-1}_{\wstar}}
++ 2\sqrt{\tr \left({\I^{-2}_{\wstar}}\right)t}
++ 2\ltwo{\I^{-1}_{\wstar}}t
+\right)
 \end{equation}
@@
 
-\subsection {Non-Asymptotic Central Limit Theorem}
-
-The standard central limit theorem for maximum likelihood estimation states that as the number of samples $n \to \infty$,
-%the distribution of the parameter vector approaches a Gaussian at the rate $O(n^{-1/2}$).
-\begin{equation}
-\I^{-1/2}_\wstar(\w-\wstar) \to \frac{1}{\sqrt n}\normal{\zero}{I}
+\begin{proof}
+We have by the triangle inequality that
+\begin{equation}
+\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wave} + \ltwo{\E\wave-\wave}
 .
-\end{equation}
-See chapter 7 of \cite{lehmann1999elements} for an elementary presentation of this result.
-In the non-asymptotic case, the distribution of the parameter vector is no longer Gaussian;
-instead, it is subgaussian.
-We say a random vector $\x$ with zero mean is \emph{subgaussian} with variance proxy $\sigma^2$ if
-\begin{equation}
-\prob{\ltwo{\x} > t} \le \exp(-\sigma^2 t^2 /2)
-.
-\end{equation}
-In notation, we say
-\begin{equation}
-\x \sim \subnormal{\sigma^2}
-.
-\end{equation}
-%We say that a random vector $\x\in\mathbb{R}^d$ is \emph{subgaussian} with variance proxy $\sigma>0$ if for all $\alpha\in\mathbb{R}^d$,
-%\begin{equation}
-%\E\exp(\trans\alpha\x) \le \exp(\ltwo{\alpha}^2\sigma^2/2)
-%.
-%\end{equation}
-The simplest example of a subgaussian random variable is a multivariate normal vector with mean $\zero$ and covariance $I$.
-In this case, $\sigma=1$.
-See \cite{vershynin2010introduction} for a tutorial on subgaussian random variables,
-including equivalent definitions and important results.
-
-Before stating the theorem, we first state the theorem's three conditions.
-The first condition is straightforward,
-but next two conditions are more technical.
-Despite the technical appearance of these assumptions,
-they are highly general.
-We note in particular that any generalized linear model (e.g. least squares, Poisson, and logistic regression),
-and many models with nonconvex likelihood functions satisfy these conditions.
-We include the conditions here for completeness,
-but defer to \cite{spokoiny2012parametricestimation} for a more detailed discussion.
-
-\begin{assumption}
-The likelihood function $f(y,\trans\x\w)$ is two times differentiable with respect to $\w$.
-\end{assumption}
-\begin{assumption}
-For all $u>0$, there exists a function $g_1(u)>0$ such that for all $|\lambda| \le g_1(u)$
-and for all $\x\in\X$,
-\begin{equation}
-\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} \le u}
-\log\E\exp\left(\lambda\frac{\trans\x\nabla_\w f(y_i,\trans\x\w)}{\ltwo{V\x}}\right)
+\label{eq:biasvar}
+\end{equation}
+The left term above captures the estimator's bias and the right term the variance.
+First we consider the bias.
+By the linearity of expectation, we have that
+\begin{align}
+\E\wave
+&=
+\E\frac{1}{m}\sum_{i=1}^m\wmle_i
+%\\&=
+=
+\frac{1}{m}\sum_{i=1}^m\E\wmle_i
+=
+\E\wmle_i
+,
+\label{eq:expwave}
+\end{align}
+and so the bias term
+$\ltwo{\wstar-\E\wave}
+=
+\ltwo{\wstar-\E\wmle_i}
+$.
+In words, the bias of the combined averaging estimator is the same as the bias of the estimator on an individual machine.
+
+Now we consider the variance.
+We have that
+\begin{align}
+\ltwobig{\wave-\E\wave}
+&=
+%\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave
+%\\&=
+%\ltwobig{\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave}
+%\label{eq:var1}
+%\\&=
+\ltwobig{\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)}
+\\&=
+\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
+\\&\sim
+\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\subnormal{\sigma^2}}
+\\&\sim
+\frac{1}{\sqrt m}\ltwobig{\I^{-1/2}_\wstar\subnormal{\sigma^2}}
+%\\&=
+%\frac{1}{\sqrt {nm}}\normal\zero{\I^{-1}_{\E\wave}}
+\label{eq:var2}
+%\\&\sim
+%\frac{1}{m}\normal{m\E\wmle_i}{\frac{m}{n}I^{-1}}
+%\\&\sim
+%\normal{\E\wmle_i}{\frac{1}{nm}I^{-1}}
+%\\
+%\wave-\E\wmle_i
+%&\sim
+%\normal{\zero}{\frac{1}{nm}I^{-1}}
+%\\
+%\wave-\E\wave
+%&\sim
+\end{align}
+%Centering the distribution and applying Equation \ref{eq:expwave} gives
+%\begin{equation}
+%\E\wave-\wave
+%\sim
+%\normal{\zero}{\frac{1}{nm}I^{-1}}
+%\end{equation}
+In the last step we used the fact that adding a centered Gaussian to itself $m$ times is equivalent to multiplying the Gaussian by $\sqrt{m}$.
+%So bounding the variance reduces to bounding the norm of a Gaussian.
+Our last step is to bound the norm of the Gaussian in Equation \ref{eq:var2}.
+Applying the bound in Proposition 1.1 from \cite{hsu2012tail} gives that with probability at least $1-\exp(-t)$,
+\begin{equation}
+\ltwo{\E\wave-\wave} \le \sqrt\frac{v_t}{nm}
+\label{eq:ct1}
+\end{equation}
+%where
+%\begin{equation}
+%\begin{aligned}
+%v_t'
+   %= \tr\left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)
+%+ 2&\sqrt{\tr \left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)^2t}&
+%\\+ 2\ltwo{\frac{1}{nm}\I^{-1}_{\wtave}}t&
+%\end{aligned}
+%\end{equation}
+%Then by the linearity of trace and scalability of norms,
+%\begin{equation}
+%v_t' =
+%\frac{1}{mn}\left(
+%\tr\I^{-1}_{\wtave}
+%+ 2\sqrt{\tr \I^{-2}_{\wtave}t}
+%+ 2\ltwo{\I^{-1}_{\wtave}}t
+%\right)
+%\label{eq:ct2}
+%\end{equation}
+Substituting Equations \ref{eq:expwave} and \ref{eq:ct1} into Equation \ref{eq:biasvar} gives the stated result.
+\end{proof}
+
+Note that this reduction in variance is essentially optimal.
+By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
+we can show that the variance of the maximum likelihood estimator trained on all of the data is
+\begin{equation}
+\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
+\label{eq:var3}
+\end{equation}
+The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.
+
+\subsection {Analysis of $\wowa$}
+
+%Now we describe our main result.
+%First we introduce some notation.
+%Let $\proj{\Wowa}$ be the projection operator onto the space $\Wowa$.
+For our next results, we need to introduce the following notation.
+If $\A$ is an arbitrary space,
+then the projection of a point $\x$ onto $\A$ is denoted by
+\begin{equation}
+\proj{\A}\x = \argmin_{\mathbf{a}\in\A} \ltwo{\mathbf{a}-\x}
+\end{equation}
+Recall that $\W$ is the vector space spanned by the $\wmle_i$s,
+and that $\wowa$ is the maximum likelihood estimator when the parameter space is restricted to $\W$.
+Our next result shows that as the number of machines increases,
+the distance between $\W$ and $\wstar$ decreases.
+
+\begin{theorem}
+Assume Equation \ref{eq:assumption} holds.
+Let $t>0$.
+Then we have that with probability at least $(1-\exp(-t))^2$,
+\begin{equation}
+\ltwo{\wstar-\proj\W\wstar}
 \le
-\frac{v_0^2\lambda^2}{2}
-\end{equation}
-where $\trans V V = \I_\wstar$ and $v_0$ is a constant.
-\end{assumption}
-\begin{assumption}
-For all $u>0$, there exists a function $b(u)>0$ such that
-\begin{equation}
-\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} = u}
-\log\E\exp\left(\frac{k(\w,\wstar)}{\ltwo{V(\w-\wstar)}}\right)
-\le
-b(u)
+%b_t\sqrt{\left(1-\frac{m}{d}\right)}
+b_t\sqrt{\left(1-\frac{m}{d}\right)}
++
+\sqrt\frac{v_{t}}{nm}
 \end{equation}
 where
-$
-k(\w,\wstar) = \E\left(f(y,\trans\x\w) - f(y,\trans\x\wstar)\right)
-$
-.
-\end{assumption}
-
-The following is a restatement of Theorem 5.2 in \cite{spokoiny2012parametricestimation}.
+%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
+\begin{equation}
+b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
+\end{equation}
+%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
+$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
+and $v_t$ is defined as in Theorem \ref{thm:wave}.
+\end{theorem}
+
+In the remainder of this section,
+we first state some simple conditions wherein the central limit theorem holds;
+then we prove Theorems 1 and 2.
+
+\subsection { The Central Limit Theorem}
+\label{sec:clt}
+
+The assumption in Equation \ref{eq:assumption} is motivated by the central limit theorem for maximum likelihood estimation.
+There are many versions of this theorem.
+Theorem 7.5.2 of \cite{lehmann1999elements} is particularly simple version and reproduced below.
+The notation has been changed to match ours.
 \begin{theorem}
-Assume Conditions 1-3 hold.
-Then whenever $t \le c\sqrt{p/n}$,
-we have with probability at least
-%Then with probability at least $1-\delta$,
-%\begin{equation}
-%\ltwo{V(\w-\wstar)} \le c\sqrt{\frac{d+\log{1/\delta}}{n}}
-%\end{equation}
-%where the constant $c$ depends on $v_o$, $g_1(u)$, and $b(u)$ defined above.
-%Equivalently, we have that
-%\begin{equation}
-%V(\w-\wstar) \sim \subnormal{\sigma}
-%\end{equation}
+\label{thm:clt}
+Assume the following mild regularity conditions:
+\begin{enumerate}[noitemsep,topsep=0pt]%,label=(\itshape\roman*)]
+\item the model is correctly specified, identifiable, and consistent;
+\item all data points in $Z_i$ are i.i.d.;
+\item the parameter space $\Theta$ is open;
+\item the matrix $\I_{\wtave}$ is non-singular;
+\item the support set $A = \{(y,\x) : f(y,\trans\x\theta) > 0\}$ is independent of $\theta$;
+and \item for all $(y,\x)\in A$, $f(y,\trans\x\theta)$ is three times differentiable with respect to $\theta$,
+and the third derivative is continuous.
+\end{enumerate}
+Then in the limit as $n\to\infty$,
+\begin{equation*}
+\wtave - \wave_i \law \frac{1}{\sqrt n} \normal{0}{\I^{-1}_\wstar}
+\end{equation*}
+Here the notation $\law$ denotes convergence in law,
+which is the standard notion of convergence for a sequence of probability distributions.
 \end{theorem}
+
+Note that Theorem \ref{thm:clt} does not require the likelihood $f$ to be convex.
+
+There are many generalizations of Theorem \ref{thm:clt}.
+For example, Theorem 5.2 in \cite{spokoiny2012parametricestimation} provides finite sample bounds with explicit dependence on the degree of model misspecification, optimization error, and the lack of independence in the data variables.
+Because Spokoiny's theorem is non-asymptotic,
+the distribution of the estimator is a sub-Gaussian random variable.
+This more general setting requires more extensive notation,
+but all of our results hold.
 
@@
 The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Using the triangle inequality, we decompose $\ltwo{\wowa-\wstar}$ into three components:
-
-\begin{equation}
-\begin{aligned}
-\ltwo{\wowa-\wstar}
-&\le
-\ltwo{\wowa-\wowastar}
-\\&
-+
-\ltwo{\wowastar - \proj{\Wowa}\wstar}
-\\&
-+
-\ltwo{\proj{\Wowa}\wstar - \wstar}
-\end{aligned}
-\end{equation}
-The first term $\ltwo{\wowa-\wowastar}$ captures the approximation error due to the limited number of points in the second round of optimization.
-The result follows by first bounding each of these terms individually,
-then applying the union bound.
-
-\subsubsection{Bounding the first term}
-
-Our first lemma is a direct consequence of the central limit theorem.
-%\begin{lemma}
-%With probability at least $\delta$,
-%\begin{equation}
-%\ltwo{\wowa-\wowastar} \le
-%\end{equation}
-%\end{lemma}
-
-\subsubsection{Bounding the second term}
-The next lemma is a direct consequence of assumptions.
-
-\subsubsection{Bounding the third term}
-Then we have that with probability at least $(1-\exp(-t))^2$,
-\begin{equation}
-\ltwo{\wstar-\proj\W\wstar}
-\le
-%b_t\sqrt{\left(1-\frac{m}{d}\right)}
-b_t\sqrt{\left(1-\frac{m}{d}\right)}
-+
-\sqrt\frac{v_{t}}{nm}
-\end{equation}
-where
-%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
-\begin{equation}
-b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
-\end{equation}
-%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
-$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
-and $v_t$ is defined as in Theorem \ref{thm:wave}.
-
-\subsection {Analyzing $\wowa$}
-
+%Using the triangle inequality, we decompose $\ltwo{\wowa-\wstar}$ into three components:
+%
+%\begin{equation}
+%\begin{aligned}
+%\ltwo{\wowa-\wstar}
+%&\le
+%\ltwo{\wowa-\wowastar}
+%\\&
+%+
+%\ltwo{\wowastar - \proj{\Wowa}\wstar}
+%\\&
+%+
+%\ltwo{\proj{\Wowa}\wstar - \wstar}
+%\end{aligned}
+%\end{equation}
+%The first term $\ltwo{\wowa-\wowastar}$ captures the approximation error due to the limited number of points in the second round of optimization.
+%The result follows by first bounding each of these terms individually,
+%then applying the union bound.
+%
+%\subsubsection{Bounding the first term}
+%
+%Our first lemma is a direct consequence of the central limit theorem.
+%%\begin{lemma}
+%%With probability at least $\delta$,
+%%\begin{equation}
+%%\ltwo{\wowa-\wowastar} \le
+%%\end{equation}
+%%\end{lemma}
+%
+%\subsubsection{Bounding the second term}
+%The next lemma is a direct consequence of assumptions.
+%
+%\subsubsection{Bounding the third term}
+%
+%
+%\subsection {Analyzing $\wowa$}
+%
 %We now extend the analysis of the previous section to the $\wowa$ estimator.
@@
 %If $\A$ is an affine space, then denote by $\proj{\A}$ the projection operator onto $\A$.
-We introduce the following notation.
-If $\A$ is an arbitrary space,
-then the projection of a point $\x$ onto $\A$ is denoted by
-\begin{equation}
-\proj{\A}\x = \argmin_{\mathbf{a}\in\A} \ltwo{\mathbf{a}-\x}
-\end{equation}
-Recall that $\W$ is the vector space spanned by the $\wmle_i$s.
-In this section we analyze the estimator $\proj{\W}\wstar$.
-
-\begin{theorem}
-\label{thm:wave}
-Let $t>0$.
-Then we have that with probability at least $(1-\exp(-t))^2$,
-\begin{equation}
-\ltwo{\wstar-\proj\W\wstar}
-\le
+
+%\begin{theorem}
+%Let $t>0$.
+%Then we have that with probability at least $(1-\exp(-t))^2$,
+%\begin{equation}
+%\ltwo{\wstar-\proj\W\wstar}
+%\le
+%%b_t\sqrt{\left(1-\frac{m}{d}\right)}
 %b_t\sqrt{\left(1-\frac{m}{d}\right)}
-b_t\sqrt{\left(1-\frac{m}{d}\right)}
-+
-\sqrt\frac{v_{t}}{nm}
-\end{equation}
-where
-%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
-\begin{equation}
-b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
-\end{equation}
-%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
-$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
-and $v_t$ is defined as in Theorem \ref{thm:wave}.
-\end{theorem}
+%+
+%\sqrt\frac{v_{t}}{nm}
+%\end{equation}
+%where
+%%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
+%\begin{equation}
+%b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
+%\end{equation}
+%%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
+%$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
+%and $v_t$ is defined as in Theorem \ref{thm:wave}.
+%\end{theorem}
 
@@
 
-\begin{proof}
 This proof follows the same form as the proof for Theorem \ref{thm:wave}.
@@
 In the first step, we translate the vectors so that the affine projection onto $\Wtave$ becomes a linear projection onto $\W$.
-The operator $(I-\proj\W)$ is a random projection onto a $(d-m)$-dimensional space.
+Define $\matV$ to be the $d \times m$ matrix whose $i$th column is the vector $(\wmle_i - \E\wave)$.
+Then $\proj{V} = \matV\matV^+$, where the superscript $^+$ indicates the Moore-Penrose pseudoinverse.
+Each entry of $\proj{V}$ is distributed as the sum of products of Gaussian distributions,
+and is therefore subGaussian.
 
@@
 }
-
-\end{proof}
 
@@
 
-\newpage
-\section*{Appendix: Detailed Central Limit Theorem}
-
-\begin{theorem}
-Assume () and ().
-If for all $u_0 > 0$,
-\begin{align}
-n^{1/2}u_0b & \ge 6 v_0 \sqrt{x+cp}
-\label{eq:cond1}
-\\
-1 + \sqrt{x+cp} &\le 3b^{-1}v_0^2g_1(u_0) n^{1/2}
-\label{eq:cond2}
-\end{align}
-then
-\begin{equation}
-\prob{\ltwo{\vv_0(\w-\wstar)} > u_0} \le \exp(-x)
-.
-\end{equation}
-\end{theorem}
-
-Manipulating Equation \ref{eq:cond1} gives
-\begin{align}
-\end{align}
-
 \end{document}

### 2016-09-28 paper.tex
changes

@@
 
+%\newcommand{\smin}[1]{s_\text{min}({#1})}
+\newcommand{\smin}{s_\text{min}}
+\newcommand{\smax}{s_\text{max}}
 \newcommand{\zero}{\text{\textbf{0}}}
@@
 
+@article{scikit-learn,
+ title={Scikit-learn: Machine Learning in {P}ython},
+ author={Pedregosa, F. and Varoquaux, G. and Gramfort, A. and Michel, V.
+         and Thirion, B. and Grisel, O. and Blondel, M. and Prettenhofer, P.
+         and Weiss, R. and Dubourg, V. and Vanderplas, J. and Passos, A. and
+         Cournapeau, D. and Brucher, M. and Perrot, M. and Duchesnay, E.},
+ journal={Journal of Machine Learning Research},
+ volume={12},
+ pages={2825--2830},
+ year={2011}
+}
+
+@article{kddcup2012,
+  author={Yanzhi Niu and Yi Wang and Gordon Sun and Aden Yue and Brian Dalessandro and Claudia Perlich and Ben Hamner},
+  title = {The Tencent Dataset and KDD-Cup'12},
+  howpublished = {URL \url{http://www.kddcup2012.org/c/kddcup2012-track2}},
+  year = {2012}
+}
+
 @book{lehmann1999elements,
@@
 
+%\begin{assumption}
+%The Fisher information matrix $\I_\wstar$ is positive definite.
+%\end{assumption}
+%
+%Condition 1 above implies that the inverse $\I_\wstar^{-1}$ and square root $\I_\wstar^{1/2}$ exist.
+
 %That is, we assume there exists a constant $c$ such that
@@
 %\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} < \frac{t}{\sqrt n}
-\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < \frac{t}{\sqrt n}
+\sqrt n \ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < t
 ,
 \end{equation}
-where $\I_\wstar$ is the Fisher information matrix.
+where $\I_\wstar$ is the positive definite Fisher information matrix.
 \end{assumption}
@@
 In the asymptotic regime when $n\to\infty$,
+very strong results are known.
 %this condition is known as root-$n$ consistency.
 %In this case,
-the variance proxy $\sigma^2=1$,
+The variance proxy $\sigma^2=1$,
 and the distribution of the estimation error is truly Gaussian.
@@
 and some mild identifiability conditions.
-Very strong results are known in the asymptotic regime.
-For example, Condition 1 is satisfied even without the i.i.d. assumption.
+More sophisticated analyses show that Condition 1 holds very generally.
+For example, the i.i.d. assumption is not strictly necessary.
 See for example \cite{spokoiny2012parametricestimation}.
@@
 The strongest non-asymptotic results on the estimation error known to the authors is due to \cite{spokoiny2012parametricestimation}.
-Spokoiny requires only a mild exponential moment condition (and does not even require the data be i.i.d.);
-but can show that Condition 1 is satisfied only for $t < O(n)$.
-
-Our condition is more general than the conditions used in previous work.
-\cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded,
-and every bounded random variable is automatically sub-Gaussian by definition.
+Spokoiny does not require any conditions on the data,
+but can show that Condition 1 is satisfied only up to $t < O(n)$.
+
+Our condition is more general than the conditions required in previous work.
+\cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded.
+This condition implies Condition 1 because every bounded random variable is automatically sub-Gaussian by definition.
 
@@
 Pick whichever one suits your problem.
-With this condition assumed, our results are elementary.
+%With this condition assumed, our results are elementary.
 
@@
 \right)
+\label{eq:vt}
 \end{equation}
@@
 \\&=
-\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
-\\&\sim
-\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\subnormal{\sigma^2}}
-\\&\sim
-\frac{1}{\sqrt m}\ltwobig{\I^{-1/2}_\wstar\subnormal{\sigma^2}}
+%\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
+%\frac{1}{m\sqrt{n}}\ltwobig{\I^{-1/2}_\wstar\sum_{i=1}^m\sqrt{n}\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
+\frac{1}{m\sqrt{n}}\ltwobig{\I^{-1/2}_\wstar\sum_{i=1}^m\Delta_i}
+%\\&\sim
+%\ltwobig{\frac{1}{m}\sum_{i=1}^m\I^{-1/2}_\wstar\subnormal{\sigma^2}}
+%\frac{1}{m\sqrt{n}}\ltwobig{\I^{-1/2}_\wstar\sum_{i=1}^m\subnormal{\sigma^2}}
+%\\&\sim
+%\frac{1}{\sqrt {mn}}\ltwobig{\I^{-1/2}_\wstar\subnormal{\sigma^2}}
+%\frac{1}{\sqrt{mn}}\ltwobig{\I^{-1/2}_\wstar\sqrt{n}\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
 %\\&=
 %\frac{1}{\sqrt {nm}}\normal\zero{\I^{-1}_{\E\wave}}
-\label{eq:var2}
+\label{eq:var1}
 %\\&\sim
@@
 \end{align}
+where
+\begin{equation}
+\Delta_i = \sqrt n \I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)
+.
+\end{equation}
+Notice that each of the $\Delta_i$ are i.i.d. centered sub-Gaussian random vectors.
+Centered sub-Gaussian random vectors inherit the following basic property from centered Gaussian random vectors:
+The sum of $m$ i.i.d. sub-Gaussians has the same distribution as $\sqrt{m}$ times the sub-Gaussian.
+That is, $\sum_{i=1}^m \Delta_i \sim \sqrt{m}\Delta_1$.
+Substituting into Equation \ref{eq:var1} gives
+\begin{equation}
+\ltwobig{\wave-\E\wave}
+\sim
+\frac{1}{\sqrt{mn}}\ltwobig{\I^{-1/2}_\wstar\Delta_1}
+.
+\label{eq:var2}
+\end{equation}
+
 %Centering the distribution and applying Equation \ref{eq:expwave} gives
@@
 %\end{equation}
-In the last step we used the fact that adding a centered Gaussian to itself $m$ times is equivalent to multiplying the Gaussian by $\sqrt{m}$.
+%In the last step we used the fact that adding a sub-Gaussian random variable to itself $m$ times is equivalent lto multiplying the Gaussian by $\sqrt{m}$.
 %So bounding the variance reduces to bounding the norm of a Gaussian.
-Our last step is to bound the norm of the Gaussian in Equation \ref{eq:var2}.
-Applying the bound in Proposition 1.1 from \cite{hsu2012tail} gives that with probability at least $1-\exp(-t)$,
-\begin{equation}
-\ltwo{\E\wave-\wave} \le \sqrt\frac{v_t}{nm}
+
+Our last step is to bound the norm in the right hand side of Equation \ref{eq:var2}.
+Theorem 2.1 from \cite{hsu2012tail} gives a bound on the norm of a positive semidefinite matrix times a sub-Gaussian random variable.
+Applying the theorem gives that %with probability at least $1-\exp(-t)$,
+\begin{equation}
+\prob{
+    \ltwobig{\I^{-1/2}_\wstar\Delta_1} < \sqrt{v_t}
+}
+\ge 1-\exp(-t)
+,
+\end{equation}
+where $v_t$ is defined as in Equation \ref{eq:vt}.
+And so the variance of $\wave$ satisfies
+\begin{equation}
+\prob{
+\ltwobig{\wave-\E\wave}
+<
+\sqrt\frac{v_t}{{mn}}
+}
+\ge 1-\exp(-t)
+.
 \label{eq:ct1}
 \end{equation}
+%and so
+%\begin{equation}
+%\ltwo{\E\wave-\wave} \le \sqrt\frac{v_t}{nm}
+%\label{eq:ct1}
+%\end{equation}
 %where
@@
 Note that this reduction in variance is essentially optimal.
-By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
-we can show that the variance of the maximum likelihood estimator trained on all of the data is
-\begin{equation}
-\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
-\label{eq:var3}
-\end{equation}
-The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.
+If we apply Condition 2 to the estimator $\wmle$ instead of $\wmle_i$,
+then we are assuming that the variance of $\wmle$ reduces at this same rate.
+%By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
+%we can show that the variance of the maximum likelihood estimator trained on all of the data is
+%\begin{equation}
+%\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
+%\label{eq:var3}
+%\end{equation}
+%The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.
 
@@
 
+\begin{lemma}
+Assume Condition 1 holds.
+Let $t>0$.
+Then we have that with probability at least $1-\exp((d-m)(-t+\ln(t+1)))$,
+\begin{equation}
+\ltwo{\wstar-\proj\W\wstar}
+=
+(t+1)\sqrt{\smin\smax\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
+\end{equation}
+where $\smin$ and $\smax$ are the minimum and maximum eigenvalues of $\I_\wstar$.
+\end{lemma}
+
+\begin{proof}
+We have that
+\begin{align}
+%&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
+\ltwo{\wstar-\proj\W\wstar}
+%\nonumber
+%\\
+&=
+\ltwo{(\wstar-\wave) - (\proj\W\wstar - \wave)}
+\nonumber
+\\
+&=
+\ltwo{(\wstar-\wave) - \proj\W(\wstar - \wave)}
+\nonumber
+\\
+&=
+\ltwo{(I-\proj\W)(\wstar - \wave)}
+.
+\label{eq:i-projW}
+\end{align}
+Recall that we defined $\matW$ to be the $d\times m$ matrix with $i$th column equal $\wmle_i$.
+The column space of $\matW$ is $\W$.
+So the projection $\proj\W = \matW\matW^+$, where the superscript $^+$ is the Moore-Penrose pseudoinverse.
+Define the matrix $\matV=\sqrt{n}I^{1/2}\matW$.
+Then the columns of $\matV$ are sub-Gaussian by Condition 1,
+and we can write
+\begin{equation}
+\proj\W
+= \I^{-1/2} \matV \matV ^+ \I^{1/2}
+.
+\label{eq:vvp}
+\end{equation}
+Let $U\Sigma \trans V$ be the singular value decomposition of $\matV$,
+where $U$ and $V$ are orthogonal matrices and $\Sigma$ is the diagonal matrix of singular values.
+By the rotational invariance of the sub-Gaussian distribution,
+$U$ and $V$ are distributed uniformly over the orthogonal group of matrices.
+Substituting into Equation \ref{eq:vvp} gives
+\begin{align}
+\proj\W
+&= \I^{-1/2} (U\Sigma V^T)(U \Sigma V^T) ^+ \I^{1/2}
+\\
+&= \I^{-1/2} (U\Sigma V^T)(V \Sigma^+ U^T) \I^{1/2}
+\\
+&= \I^{-1/2} U\Sigma \Sigma^+ U^T \I^{1/2}
+.
+\label{eq:vvp}
+\end{align}
+Substituting into Equation \ref{eq:i-projW} gives
+\begin{align}
+~~~~~&\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\W\wstar}
+\nonumber
+\\
+&=
+\ltwo{(I-\I^{-1/2} U\Sigma \Sigma^+ U^T \I^{1/2})(\wstar - \wave)}
+\\
+&=
+\ltwo{\I^{-1/2} U(I-\Sigma \Sigma^+ )U^T \I^{1/2}(\wstar - \wave)}
+\\
+&=
+\sqrt{\smin}\ltwo{U(I-\Sigma \Sigma^+ )U^T \I^{1/2}(\wstar - \wave)}
+\\
+&=
+\sqrt{\smin}\ltwo{(I-\Sigma \Sigma^+ )U^T \I^{1/2}(\wstar - \wave)}
+\end{align}
+%The last line above follows from the rotational invariance of the $L2$ norm.
+The operator $(I-\Sigma\Sigma^+)\trans U$ projects a vector onto a randomly chosen subspace of dimension $d-m$.
+Lemma 2.2 from \cite{dasgupta2003elementary} provides a bound for this situation.
+For all $t>0$, with probability at least $1-\exp((d-m)(-t+\ln (t+1)))$,
+\begin{align}
+\ltwo{(I-\Sigma\Sigma^+)\trans U \x} \le (t+1) \sqrt{1-\frac{m}{d}}
+.
+\end{align}
+Substituting gives
+\begin{align}
+~~~~~&\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\W\wstar}
+\nonumber
+\\
+&=
+(t+1)\sqrt{\smin\left(1-\frac{m}{d}\right)}\ltwo{\I^{1/2}(\wstar - \wave)}
+\\
+&=
+(t+1)\sqrt{\smin\smax\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
+\end{align}
+
+\ignore{
 \begin{theorem}
-Assume Equation \ref{eq:assumption} holds.
+Assume Condition 1 holds.
 Let $t>0$.
@@
 
-In the remainder of this section,
-we first state some simple conditions wherein the central limit theorem holds;
-then we prove Theorems 1 and 2.
-
-\subsection { The Central Limit Theorem}
-\label{sec:clt}
-
-The assumption in Equation \ref{eq:assumption} is motivated by the central limit theorem for maximum likelihood estimation.
-There are many versions of this theorem.
-Theorem 7.5.2 of \cite{lehmann1999elements} is particularly simple version and reproduced below.
-The notation has been changed to match ours.
-\begin{theorem}
-\label{thm:clt}
-Assume the following mild regularity conditions:
-\begin{enumerate}[noitemsep,topsep=0pt]%,label=(\itshape\roman*)]
-\item the model is correctly specified, identifiable, and consistent;
-\item all data points in $Z_i$ are i.i.d.;
-\item the parameter space $\Theta$ is open;
-\item the matrix $\I_{\wtave}$ is non-singular;
-\item the support set $A = \{(y,\x) : f(y,\trans\x\theta) > 0\}$ is independent of $\theta$;
-and \item for all $(y,\x)\in A$, $f(y,\trans\x\theta)$ is three times differentiable with respect to $\theta$,
-and the third derivative is continuous.
-\end{enumerate}
-Then in the limit as $n\to\infty$,
-\begin{equation*}
-\wtave - \wave_i \law \frac{1}{\sqrt n} \normal{0}{\I^{-1}_\wstar}
-\end{equation*}
-Here the notation $\law$ denotes convergence in law,
-which is the standard notion of convergence for a sequence of probability distributions.
-\end{theorem}
-
-Note that Theorem \ref{thm:clt} does not require the likelihood $f$ to be convex.
-
-There are many generalizations of Theorem \ref{thm:clt}.
-For example, Theorem 5.2 in \cite{spokoiny2012parametricestimation} provides finite sample bounds with explicit dependence on the degree of model misspecification, optimization error, and the lack of independence in the data variables.
-Because Spokoiny's theorem is non-asymptotic,
-the distribution of the estimator is a sub-Gaussian random variable.
-This more general setting requires more extensive notation,
-but all of our results hold.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection {Proof of Theorem 1}
-
-We have by the triangle inequality that
-\begin{equation}
-\ltwo{\wstar-\wave} \le \ltwo{\wstar-\E\wave} + \ltwo{\E\wave-\wave}
-\label{eq:biasvar}
-\end{equation}
-The left term above captures the estimator's bias and the right term the variance.
-First we consider the bias.
-By the linearity of expectation, we have that
-\begin{align}
-\E\wave
-&=
-\E\frac{1}{m}\sum_{i=1}^m\wmle_i
-%\\&=
-=
-\frac{1}{m}\sum_{i=1}^m\E\wmle_i
-=
-\E\wmle_i
-\label{eq:expwave}
-\end{align}
-and so the bias term
-$\ltwo{\wstar-\E\wave}
-=
-\ltwo{\wstar-\E\wmle_i}
-$.
-In words, the bias of the combined averaging estimator is the same as the bias of the estimator on an individual machine.
-
-Now we consider the variance.
-We have by Equations \ref{eq:expwave} and \ref{eq:assumption} that
-\begin{align}
-\wave-\E\wave
-&=
-%\frac{1}{m}\sum_{i=1}^m\wmle_i-\E\wave
-%\\&=
-\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wave\right)
-\label{eq:var1}
-\\&=
-\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)
-\\&\sim
-\frac{1}{m}\sum_{i=1}^m\left(\frac{1}{\sqrt n}\normal\zero{\I^{-1}_{\E\wave}}\right)
-\\&=
-\frac{1}{\sqrt {nm}}\normal\zero{\I^{-1}_{\E\wave}}
-\label{eq:var2}
-%\\&\sim
-%\frac{1}{m}\normal{m\E\wmle_i}{\frac{m}{n}I^{-1}}
-%\\&\sim
-%\normal{\E\wmle_i}{\frac{1}{nm}I^{-1}}
-%\\
-%\wave-\E\wmle_i
-%&\sim
-%\normal{\zero}{\frac{1}{nm}I^{-1}}
-%\\
-%\wave-\E\wave
-%&\sim
-\end{align}
-%Centering the distribution and applying Equation \ref{eq:expwave} gives
-%\begin{equation}
-%\E\wave-\wave
-%\sim
-%\normal{\zero}{\frac{1}{nm}I^{-1}}
-%\end{equation}
-In the last step we used the fact that adding a centered Gaussian to itself $m$ times is equivalent to multiplying the Gaussian by $\sqrt{m}$.
-%So bounding the variance reduces to bounding the norm of a Gaussian.
-Our last step is to bound the norm of the Gaussian in Equation \ref{eq:var2}.
-Applying the bound in Proposition 1.1 from \cite{hsu2012tail} gives that with probability at least $1-\exp(-t)$,
-\begin{equation}
-\ltwo{\E\wave-\wave} \le \sqrt\frac{v_t}{nm}
-\label{eq:ct1}
-\end{equation}
-%where
-%\begin{equation}
-%\begin{aligned}
-%v_t'
-   %= \tr\left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)
-%+ 2&\sqrt{\tr \left({\frac{1}{mn}\I^{-1}_{\wtave}}\right)^2t}&
-%\\+ 2\ltwo{\frac{1}{nm}\I^{-1}_{\wtave}}t&
-%\end{aligned}
-%\end{equation}
-%Then by the linearity of trace and scalability of norms,
-%\begin{equation}
-%v_t' =
-%\frac{1}{mn}\left(
-%\tr\I^{-1}_{\wtave}
-%+ 2\sqrt{\tr \I^{-2}_{\wtave}t}
-%+ 2\ltwo{\I^{-1}_{\wtave}}t
-%\right)
-%\label{eq:ct2}
-%\end{equation}
-Substituting Equations \ref{eq:expwave} and \ref{eq:ct1} into Equation \ref{eq:biasvar} gives the stated result.
-
-Note that this reduction in variance is essentially optimal.
-By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
-we can show that the variance of the maximum likelihood estimator trained on all of the data is
-\begin{equation}
-\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
-\label{eq:var3}
-\end{equation}
-The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection {Proof of Theorem 2}
-
-%Using the triangle inequality, we decompose $\ltwo{\wowa-\wstar}$ into three components:
-%
-%\begin{equation}
-%\begin{aligned}
-%\ltwo{\wowa-\wstar}
-%&\le
-%\ltwo{\wowa-\wowastar}
-%\\&
-%+
-%\ltwo{\wowastar - \proj{\Wowa}\wstar}
-%\\&
-%+
-%\ltwo{\proj{\Wowa}\wstar - \wstar}
-%\end{aligned}
-%\end{equation}
-%The first term $\ltwo{\wowa-\wowastar}$ captures the approximation error due to the limited number of points in the second round of optimization.
-%The result follows by first bounding each of these terms individually,
-%then applying the union bound.
-%
-%\subsubsection{Bounding the first term}
-%
-%Our first lemma is a direct consequence of the central limit theorem.
-%%\begin{lemma}
-%%With probability at least $\delta$,
-%%\begin{equation}
-%%\ltwo{\wowa-\wowastar} \le
-%%\end{equation}
-%%\end{lemma}
-%
-%\subsubsection{Bounding the second term}
-%The next lemma is a direct consequence of assumptions.
-%
-%\subsubsection{Bounding the third term}
-%
-%
-%\subsection {Analyzing $\wowa$}
-%
-%We now extend the analysis of the previous section to the $\wowa$ estimator.
-%We show that its variance is bounded by the variance of $\wave$,
-%but that the bias shrinks as $O(1-m/d)$.
-%
-%Recall from Section \ref{sec:vcdim} that the estimator $\wowa$ has low error with high probability compared to any parameters in the space $\Wave$.
-%Denote by $\proj\Wave$ the projection operator onto the space $\Wave$.
-%In this section, we analyze the bias and variance of $\proj\Wave\wstar$.
-
-%If $\A$ is an affine space, then denote by $\proj{\A}$ the projection operator onto $\A$.
-
-%\begin{theorem}
-%Let $t>0$.
-%Then we have that with probability at least $(1-\exp(-t))^2$,
-%\begin{equation}
-%\ltwo{\wstar-\proj\W\wstar}
-%\le
-%%b_t\sqrt{\left(1-\frac{m}{d}\right)}
-%b_t\sqrt{\left(1-\frac{m}{d}\right)}
-%+
-%\sqrt\frac{v_{t}}{nm}
-%\end{equation}
-%where
-%%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
-%\begin{equation}
-%b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
-%\end{equation}
-%%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
-%$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
-%and $v_t$ is defined as in Theorem \ref{thm:wave}.
-%\end{theorem}
-
-\ignore{
-\begin{theorem}
-%Assume that $n$ is large enough that $\wmle_i$ is distributed according to $\normal{\wtave}{\I^{-1}_{\wtave}}$.
-%Let $\delta\in(0,1)$.
-%Then with probability at least $1-\delta$,
-We have that
-\begin{equation}
-\ltwo{\wstar-\proj\Wave\wstar}
-\le \epsilon_{bias} + \epsilon_{var}
-\end{equation}
-where for all $\tbias>0$,
-\begin{equation*}
-%\begin{split}
-%\prob{
-    \epsilon_{bias}
-    \le
-    \ltwo{\wstar-\wtave}\sqrt{(\tbias + 1)\left(1-\frac{m}{d}\right)}
-%}
-%\\
-%\ge
-%1-\exp \bigg(-\frac{b}{2} (\beta-\ln(\beta+1))\bigg)~~&
-%\end{split}
-\end{equation*}
-with probability at least
-$1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$
-and for all $\tvar>0$,
-\begin{equation*}
-\begin{split}
-&\epsilon_{var}
-\\
-&\le
-\sqrt{
-    \frac{1}{m}
-    \left(\tr{\I^{-1}_{\wtave}}
-        + 2\sqrt{\tr ({\I^{-2}_{\wtave}})\tvar}
-        + 2\ltwo{\I^{-1}_{\wtave}}\tvar
-    \right)
-}
-\end{split}
-\end{equation*}
-with probability at least $1-\exp(-\tvar)$.
-
-%\begin{equation}
-%\begin{split}
-%&\ltwo{\wstar-\proj\Wave\wstar}
-%\\
-%&\le
-%\ltwo{\wstar-\wtave}\sqrt{(\beta + 1)\left(1-\frac{m}{d}\right)}
-%\\
-%&+
-%\sqrt{
-    %\frac{1}{m}
-    %\left(\tr{\I^{-1}_{\wtave}}
-        %+ 2\sqrt{\tr ({\I^{-1}_{\wtave}}^2)t}
-        %+ 2\ltwo{\I^{-1}_{\wtave}}t
-    %\right)
-%}
-%\le
-%\exp \bigg(-\frac{b}{2} (\beta-\ln(\beta+1))\bigg)~~&
-%\end{split}
-%\end{equation}
-\end{theorem}
-}
-
 This proof follows the same form as the proof for Theorem \ref{thm:wave}.
@@
 then we bound each component separately.
-To decompose the error, we need to introduce the following affine space.
+
+Before we begin, we need to introduce the following vector space:
 \begin{align}
-\Waff&=\affspan \{\wmle_i\}_{i=1}^m
-\\
-\WaffE&=\affspan \{\wmle_i - \wave + \E\wave\}_{i=1}^m
+\V &= \vecspan \{\wmle_i - \wtave\}_{i=1}^m
+%,\\
+%A &= \{ \vv + \E\wave : \vv \in \V \}
+%,\\
+%A &= \{ \vv + \wave : \vv \in \V \}
+%.
 \end{align}
-We now have that
+%We also introduce notation for constructing affine spaces:
+%\begin{equation}
+%\V + \x = \{ \vv + \x : \vv \in \V \}
+%.
+%\end{equation}
+%We now have that
+We now have that:
+\begin{align}
+~~~~&\!\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\W\wstar}
+%&\ltwo{\wstar-\proj\W\wstar}
+\nonumber
+\\
+&\le
+\ltwo{(\wstar-\wtave) - \proj\W(\wstar-\wtave)}
+\\
+&\begin{aligned}
+&\le
+%\ltwo{
+\lVert(\wstar-\wtave)
+\\
+&~~~
+- (\proj\V(\wstar-\wtave) + \wtave - \wave)
+%}
+\rVert
+\end{aligned}
+\\
+&\begin{aligned}
+&\le
+\ltwo{(\wstar-\wtave) - \proj\V(\wstar-\wtave)}
+\\
+&~~~
++ \ltwo{\wtave - \wave}
+\end{aligned}
+\\
+&=
+\ltwo{(I-\proj\V)(\wstar-\wtave)} + \ltwo{\wtave - \wave}
+%
+%
+%
+%\\
+%&\le
+%\ltwo{\wstar-(\proj{\V}\wstar + \wave)}
+%\label{eq:step1}
+%\\
+%&\!\begin{aligned}
+%&\le
+%\ltwo{\wstar-(\proj{\V}\wstar+\wtave)}
+%\\&~~~
+%+
+%\ltwo{(\proj{\V}\wstar+\wtave)-(\proj{\V}\wstar+\wave)}
+%\end{aligned}
+%\label{eq:step2}
+%\\
+%&=
+%\ltwo{\wstar-(\proj{\V}\wstar+\wtave)}
+%+
+%\ltwo{\wtave-\wave}
+%\label{eq:step3}
+\end{align}
+Inequality \ref{eq:step1} follows because $(\proj\V\wstar + \wave) \in \W$.
+Inequality \ref{eq:step2} follows from the triangle inequality.
+The rightmost term in Equation \ref{eq:step3} is the variance of $\wave$,
+which we previously bounded with high probability in Equation \ref{eq:ct1}.
+
+The affine space $\V+\wave$ is a subset of $\W$.
+So,
 \begin{align}
@@
 &\le
-\ltwo{\wstar-\proj\Waff\wstar}
-\\
+\ltwo{\wstar-\proj{(\V+\wave)}\wstar}
+\end{align}
+By the triangle inequality, we have
+\begin{align}
+%\\
+&\begin{aligned}
+\ltwo{\wstar-\proj\W\wstar}
 &\le
-\ltwo{\wstar-\proj\WaffE\wstar}
-%\\&
+\ltwo{\wstar-\proj{(\V+\E\wave)}\wstar}
+\\&~~~~
 +
-\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
+\ltwo{\proj{(\V+\E\wave)}\wstar-\proj{(\V+\wave)}\wstar}
+\end{aligned}
 %\\&+
@@
 \end{align}
-The first line above follows because $\Waff\subseteq\W$,
-and the second by the triangle inequality.
-The left term of the inequality above can be thought of as akin to the bias of the estimator $\proj\Waff\wstar$,
+The left term of the inequality above can be thought of as akin to the bias of the ``estimator'' $\proj\Waff\wstar$,
 and the right term can be thought of as akin to the variance.
@@
 \begin{align}
-&\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
+&\ltwo{\proj{(\V+\wtave)}\wstar-\proj{(\V+\wave)}\wstar}
 \\
 &=
-\ltwo{(\proj\Waff\wstar+\wave-\wtave)-\proj\Waff\wstar}
+\ltwo{(\proj{(\V+\wave)}\wstar + \wave - \wtave) -\proj{(\V+\wave)}\wstar}
+%\ltwo{(\proj\Waff\wstar+\wave-\wtave)-\proj\Waff\wstar}
 \\&=
@@
 \end{align}
-The first equality follows because the space $\WaffE$ is a translation of $\Waff$ by the vector $(\wave-\wtave)$.
+%The first equality follows because the space $\WaffE$ is a translation of $\Waff$ by the vector $(\wave-\wtave)$.
 The probabilistic inequality is due to Equation \ref{eq:ct1} in the proof of Theorem \ref{thm:wave}.
+This shows that the variance of our estimator is the same as the variance of the averaging estimator,
+which we already argued is essentially optimal.
+
+%To decompose the error, we need to introduce the following two affine spaces:
+%\begin{align}
+%\Waff&=\affspan \{\wmle_i\}_{i=1}^m
+%,
+%\\
+%\WaffE&=\affspan \{\wmle_i + (\wave - \E\wave)\}_{i=1}^m
+%.
+%\end{align}
+%We now have that
+%\begin{align}
+%\ltwo{\wstar-\proj\W\wstar}
+%&\le
+%\ltwo{\wstar-\proj\Waff\wstar}
+%\\
+%&\le
+%\ltwo{\wstar-\proj\WaffE\wstar}
+%%\\&
+%+
+%\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
+%%\\&+
+%%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
+%\end{align}
+%The first line above follows because $\Waff\subseteq\W$,
+%and the second by the triangle inequality.
+%The left term of the inequality above can be thought of as akin to the bias of the ``estimator'' $\proj\Waff\wstar$,
+%and the right term can be thought of as akin to the variance.
+
+%We begin by bounding the right term.
+%We have that with probability at least $1-\exp(-t)$,
+%\begin{align}
+%&\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
+%\\
+%&=
+%\ltwo{(\proj\Waff\wstar+\wave-\wtave)-\proj\Waff\wstar}
+%\\&=
+%\ltwo{\wave-\wtave}
+%\\&\le
+%\sqrt\frac{v_t}{mn}
+%\end{align}
+%The first equality follows because the space $\WaffE$ is a translation of $\Waff$ by the vector $(\wave-\wtave)$.
+%The probabilistic inequality is due to Equation \ref{eq:ct1} in the proof of Theorem \ref{thm:wave}.
 
@@
 and is therefore subGaussian.
-
-%Now, $(I-\proj\W)$ is a random projection onto a subspace of dimension $d-m$.
-%(\fixme{This projection is not chosen uniformly at random because the covariance of $\wmle_i$ is not the identity.})
-%This gives us that for all $\tbias>0$,
-%\begin{equation}
-%\begin{split}
-%\prob{\ltwo{\wstar-\proj\Wtave\wstar} \ge \ltwo{\wstar-\wtave}\sqrt{(\tbias + 1)\left(1-\frac{m}{d}\right)}}&
-%\\
-%\le
-%\exp \bigg(-\frac{b}{2} (\tbias-\ln(\tbias+1))\bigg)~~&
-%\end{split}
-%\end{equation}
-\ignore{
-Note that the space $\Wave$ depends on the data.
-Therefore, we can think of $\proj\Wave\wstar$ as an estimator of $\wstar$.
-As in the proof of Theorem \ref{thm:wave},
-we begin by decomposing the MSE of $\proj\Wave\wstar$ into its bias and variance components.
-\begin{equation}
-\begin{split}
-\ltwo{\wstar-\proj\Wave\wstar}^2
-&=
-\ltwo{\wstar-\E\proj\Wave\wstar}^2
-\\&+
-\ltwo{\E\proj\Wave\wstar-\proj\Wave\wstar}^2
-\end{split}
-\end{equation}
-We begin with the variance.
-
-%Define $\Wtave = \vecspan\{\wmle_i-\wtave\}_{i=1}^m$.
-%Then,
-We have that
-\begin{align}
-\ltwo{\wstar-\proj\Wave\wstar}
-&\le
-\ltwo{\wstar-\proj\Wtave\wstar}
-+
-\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
-\\
-&=
-\ltwo{\wstar-\proj\Wtave\wstar}
-+
-\ltwo{\wave-\wtave}
-\label{eq:errdecomp}
-\end{align}
-The first line follows from the triangle inequality.
-The second line follows because every point in $\Wtave$ is a translation of a point in $\Wave$ by the vector $\wave-\wtave$.
-
-Define the spaces
-\begin{align}
-\W&=\vecspan\{\wmle_i - \wave\}_{i=1}^m\\
-\Wave&=\affspan\{\wave_i\}_{i=1}^m\cup\{\wave\}\\
-\Wtave&=\affspan\{\wave_i\}_{=1}^m\cup\{\wtave\}
-%\Wave&=\vecspan\{\wave_i-\wave\}_{i=1}^m\\
-%\Wtave&=\vecspan\{\wave_i-\wtave\}_{i=1}^m
-\end{align}
-Denote by $\proj\A$ to be the projection operator onto the space $\A$.
-Denote by $\I_\w$ the Fischer information matrix at the point $\w$.
-
-%\begin{lemma}
-%Let $a$ and $b$ be integers with $b<a$.
-%Let $\x$ be a random vector on the $a$ dimensional unit sphere.
-%Let $T : \mathbb{R}^a \to \mathbb{R}^b$ be the projection operator that preserves the first $b$ coordinates of a vector and drops the rest.
-%Then for all $\beta>0$,
-%\begin{equation}
-%\prob{\ltwo{T\x} \ge \sqrt{(\beta + 1)\frac{b}{a}}}
-%\le
-%\exp \left(\frac{b}{2}(-\beta+\ln(\beta+1))\right)
-%\end{equation}
-%\end{lemma}
-
-We will now bound the two terms in Equation \ref{eq:errdecomp} separately.
-To bound the left term, we introduce the following lemma.
-It is adapted from Lemma 2.2 in \cite{dasgupta2003elementary}.
-It has been simplified to ignore cases irrelevant for our purpose.
-\begin{lemma}
-Let $a$ and $b$ be integers with $b<a$.
-Let $\x\in\mathbb{R}^a$ be an arbitrary vector.
-Let $T : \mathbb{R}^a \to \mathbb{R}^b$ be projection onto a subspace of size $b$ distributed uniformly at random.
-Then for all $\beta>0$,
-\begin{equation}
-\begin{split}
-\prob{\ltwo{T\x} \ge \ltwo\x\sqrt{(\beta + 1)\frac{b}{a}}}&
-\\
-\le
-\exp \bigg(-\frac{b}{2} & (\beta-\ln(\beta+1))\bigg)
-\end{split}
-\end{equation}
-\end{lemma}
-To apply this lemma, note that $\proj\Wtave$ is not a linear projection but an affine projection.
-We first apply a translation so that we are working with the linear projection $\proj\W$.
-\begin{align}
-\ltwo{\wstar-\proj\Wtave\wstar}
-&=
-\ltwo{(\wstar-\wtave)-\proj\W(\wstar-\wtave)}
-\\
-&=
-\ltwo{(I-\proj\W)(\wstar-\wtave)}
-\end{align}
-Now, $(I-\proj\W)$ is a random projection onto a subspace of dimension $d-m$.
-(\fixme{This projection is not chosen uniformly at random because the covariance of $\wmle_i$ is not the identity.})
-This gives us that for all $\tbias>0$,
-\begin{equation}
-\begin{split}
-\prob{\ltwo{\wstar-\proj\Wtave\wstar} \ge \ltwo{\wstar-\wtave}\sqrt{(\tbias + 1)\left(1-\frac{m}{d}\right)}}&
-\\
-\le
-\exp \bigg(-\frac{b}{2} (\tbias-\ln(\tbias+1))\bigg)~~&
-\end{split}
-\end{equation}
-}
-
-%The following lemma is adapted from Theorem 3.1 in \cite{matouvsek2008variants}.
-%It has been simplified to ignore cases irrelevant for our purpose.
-%\begin{lemma}
-%Let $a$ and $b$ be integers with $a>b$.
-%Let $T\in\mathbb{R}^{a\times b}$ be a random matrix with entries drawn i.i.d. from the standard normal distribution.
-%Let $\x\in\mathbb{R}^a$ be an arbitrary vector.
-%Then there exists a constant $c$ such that for all $\delta\in(0,1)$,
-%\begin{equation*}
-%\prob{\ltwo{T\x} \le \left(\sqrt{b}+c\sqrt{\log\frac{\delta}{2}}\right)\ltwo\x} \le 1-\delta
-%\end{equation*}
-%\end{lemma}
-
-%\begin{cor}
-%%Assume that Conditions (1-5) hold.
-%%Then as $n\to\infty$,
-%%\begin{enumerate}
-%%\item
-%If $f$ is locally $\sigma$-strongly convex,
-%\begin{equation}
-%\ltwo{\wstar-\wowa} \le
-%\end{equation}
-%%\end{enumerate}
-%\end{cor}
-%
-%\begin{proof}
-%\end{proof}
-
-%\subsection{Proposed Solution}
-%
-%We propose the following solution.
-%Define the matrix $W : \mathbb{R}^{d\times m}$ to have $i$th column equal $\wmle_i$.
-%%That is,
-%%\begin{equation}
-%%W = (\wmle_1, \wmle_2, ..., \wmle_m)
-%%\end{equation}
-%Then our estimator is defined as
-%\begin{equation}
-%\begin{aligned}
-%\alpha &= \argmax_\w \sum _{(\x,y)\in Z_i} f(y,\trans\x W\w)
-%\\
-%\w &= \trans W \alpha
-%\end{aligned}
-%\end{equation}
-%Note that the inner $\w$ above is a vector of dimension $m$, not $d$ (as elsewhere).
-%In words
-%%In words, this estimator takes the dataset $Z'$ and projects it onto the subspace spanned by the
-
-\newpage
+}
+\end{proof}
 
@@
 
+\begin{figure*}[ht!]
+%\begin{tikzpicture}
+%\node at (0,2) {};
+%\node at (0,-2) {};
+\begin{tabular}{ccc}
+  \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-autoreg-laplace-log-1000-100-20}
+& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-laplace-log-1000-100-100}
+& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-laplace-log-1000-1000-100}
+\\
+  \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-autoreg-spike-log-1000-100-20}
+& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-spike-log-1000-100-100}
+& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-spike-log-1000-1000-100}
+\end{tabular}
+%\end{tikzpicture}
+\caption{test}
+\label{fig:lambda}
+\end{figure*}
+
 \section{EXPERIMENTS}
@@
 
-We test our parallelization procedure on two logistic regression tasks.
+We test our distributed estimator on two logistic regression tasks.
 The first task uses synthetic data.
-The second task uses real world add click data from the tencent search engine.
+The second task uses real world add click data from the Tencent search engine.
+
+In each experiment, we compare our $\wowa$ estimator with four baseline estimators:
+the naive estimator using the data from only a single machine $\wmle_i$;
+the averaging estimator $\wave$;
+the bootstrap estimator $\wboot$;
+and the oracle estimator of all data trained on a single machine $\wmle$.
+The $\wboot$ estimator has a parameter $r$ that needs to be tuned.
+We make an effort to present the $\wboot$ estimator in the best light possible.
+In all experiments we evaluate $\wboot$ with $r \in \{0.005,0.01,0.02,0.04,0.1,0.2,0.3\}$,
+which is a set recommended in the original paper \citep{zhang2012communication},
+and then report only the value of $r$ with highest true likelihood.
+Thus we are reporting an overly optimistic estimate of the performance of $\wboot$,
+and as we shall see our estimator $\wowa$ still tends to perform better.
 
@@
 \begin{tabular}{ccc}
-  \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-100-star.pdf}
-& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-1000-star.pdf}
-& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-10000-star.pdf}
-\\
-  \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-uniform-log-1000-100-star.pdf}
-& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-uniform-log-1000-1000-star.pdf}
-& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-uniform-log-1000-10000-star.pdf}
-\\
+  %\includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-100-star.pdf}
+%& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-1000-star.pdf}
+%& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-10000-star.pdf}
+%\\
+  %\includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-uniform-log-1000-100-star.pdf}
+%& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-uniform-log-1000-1000-star.pdf}
+%& \includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-uniform-log-1000-10000-star.pdf}
+%\\
   \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-100-star.pdf}
@@
 \caption{}
+\label{fig:synscale}
 \end{figure*}
 
-\begin{figure}
-%\begin{tikzpicture}
-\begin{tabular}{c}
-\\ \includegraphics[width=0.45\textwidth]{img/logl2-star-logl2-auto-normal-log-1000-100-100.pdf}
-%\\ \includegraphics[width=0.45\textwidth]{img/logl2-star-logl2-auto-uniform-log-1000-100-100.pdf}
-\\ \includegraphics[width=0.45\textwidth]{img/logl1-star-logl2-auto-laplace-log-1000-100-100.pdf}
-\\ \includegraphics[width=0.45\textwidth]{img/logl1-star-logl2-auto-spike-log-1000-100-100.pdf}
-%\\ \includegraphics[width=0.45\textwidth]{img/logl2-star-logl2-auto-normal-log-1000-1000-100.pdf}
-%\\ \includegraphics[width=0.45\textwidth]{img/logl2-star-logl2-auto-normal-1000-100-100.pdf}
-\end{tabular}
-%\end{tikzpicture}
-\caption{
-    A similar effect happens when $d=1000$ or $d=10000$,
-    but it is much less pronounced.
-}
-\end{figure}
-
-We consider two logistic regression models.
-In the first,
-each component of the parameter vector $\theta_i$ is sampled i.i.d. from the standard normal distribution.
-This is the Bayesian interpretation of $L2$ regularization.
-In the second, $\theta_i$ is sampled i.i.d. from the Laplace (two-sided exponential) distribution with scale 1.
-This is the Bayesian interpretation of $L1$ regularization.
-In both models, the data points are then sampled as
+We generate the data according to the following sparse logistic regression model.
+Each component of the true parameter vector $\wstar$ is sampled i.i.d. from a spike and slab distribution.
+With probability 0.9, the component is 0;
+with probability 0.1, the component is sampled from a standard normal distribution.
+The data points are then sampled as
 \begin{equation}
 \begin{aligned}
-%\theta &\sim \normal{0}{I}
-%\\
 \x_i &\sim \normal{0}{I}
@@
 ~~~
-%y_i &= \frac{1}{1+\exp(-\trans\x_i\theta)}
-y_i = \left(1+\exp(-\trans\x_i\theta)\right)^{-1}
+y_i = \left(1+\exp(-\trans\x_i\wstar)\right)^{-1}
+.
 \end{aligned}
 \end{equation}
-
-Our first experiment shows how the parallel algorithms scale as the number of machines $m$ increases.
+In all experiments, we use the $L1$ regularizer to induce sparsity in our estimates of $\wstar$.
+Our estimators will be biased because the model is misspecified.
+$L1$ regularization actually corresponds to a Laplace prior on the parameter vector.
+Results are qualitatively similar when using a uniform, normal, and Laplace prior on $\wstar$.
+These results are not shown due to space limitations.
+
+Our first experiment shows the sensitivity of the estimators to the strength of regularization $\lambda$.
+In this experiment, $\lambda$ was allowed to vary from $10^{-4}$ to $10^4$.
+For each value of $\lambda$, we randomly generated 50 datasets and calculated the corresponding estimators.
+Figure \ref{fig:lambda} shows the average of the results.
+Our $\wowa$ estimator exhibits significantly less sensitivity to the selection of $\lambda$.
+
+Our second experiment shows how the parallel algorithms scale as the number of machines $m$ increases.
 We fix $n=1000$ data points per machine, so the size of the dataset $mn$ grows as we add more machines.
@@
 but processing resources are scarce.
-We further consider three regimes:
-%high dimensional ($d=10000$), medium dimensional ($d=1000$), and low dimensional ($d=100$) data.
-more data than dimensions ($d=100$),
-one data point per dimension ($d=1000$),
-and less data than dimensions ($d=10000$).
+%We further consider three regimes:
+%more data than dimensions ($d=100$),
+%one data point per dimension ($d=1000$),
+%and less data than dimensions ($d=10000$).
+Each machine independently uses cross validation to select the $\lambda$ that best fits the data locally.
+There are three advantages to this method of parameter selection.
+First, there is no additional communication overhead for model selection because model selection is a completely local task.
+Second, existing optimizers have built-in model selection routines which make the process easy to implement.
+In our experiments, we use the default model selection procedure from Python's SciKit-Learn \citep{scikit-learn}.
+Third, the data may be best fit using different regularization strengths for each machine,
+and this method allows that to happen.
+The results are shown in Figure \ref{fig:synscale}.
+The performance of our method scales well with the number of machines,
+and in some cases even outperforms the oracle on all of the data.
 
@@
 
+In this section,
+we evaluate our estimator on real world data from the KDD 2012 Cup \citep{kddcup2012}.
+The goal is to predict whether a user will click on an add from the Tencent internet search engine.
+This dataset was previously used to evaluate the performance of $\wboot$ \citep{zhang2012communication}.
+
+This dataset is too large to fit on a single machine.
+There are 235,582,879 distinct data points,
+each of dimension 741,725.
+For each test, we split the data into 80 percent training data and 20 percent test data.
+The training data is further subdivided into 128 partitions,
+one for each of the machines used.
+
 \section{CONCLUSION}
@@
 
+\clearpage
 \bibliographystyle{plainnat}

### 2016-09-28 paper.tex
changes

@@
 \newcommand{\smax}{s_\text{max}}
+\newcommand{\qhi}{q_\text{hi}}
+\newcommand{\qlo}{q_\text{lo}}
 \newcommand{\zero}{\text{\textbf{0}}}
@@
 so our algorithm retains the speed advantages of non-interactive algorithms.
-We now formally describe our problem and proposed solution.
-
-\subsection{Problem Statement}
+
+In the next section, we formally describe our problem setting and proposed solution.
+In Section 3, we compare our solution to related solutions.
+We highlight how the analysis of existing algorithms requires more limiting assumptions than our own,
+and show in detail why most existing non-interactive regret bounds do not apply to our problem.
+Section 4.1 presents a novel, very general analysis of the averaging estimator.
+Section 4.2 shows that two parameters to our algorithm $m$ and $\nreopt$ act as ``knobs'' that let the user easily trade off statistical performance for communication cost.
+Section 5 shows experimentally that our algorithm performs well.
+
+\section{Problem Statement and Solution}
 %\section{Problem Setting and Our Solution}
@@
 %\end{equation}
-\begin{assumption}
+
+\newtheorem*{sgt}{The Sub-Gaussian Tail (SGT) Condition}
+\begin{sgt}
+Let $\theta$ be an estimator trained on $n$ data points.
 Let $t>0$ and $\sigma^2$ be a constant called the \emph{variance proxy}.
@@
 %\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} < \frac{t}{\sqrt n}
-\sqrt n \ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < t
+%\sqrt n \ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < t
+\sqrt n \ltwobig{\I^{-1/2}_\wstar(\theta-\E\theta)} < t
 ,
 \end{equation}
-where $\I_\wstar$ is the positive definite Fisher information matrix.
-\end{assumption}
+where $\I_\wstar$ is the positive definite Fisher information matrix at the parameter vector's true value $\wstar$.
+\end{sgt}
 
@@
 
-Our condition is more general than the conditions required in previous work.
-\cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded.
-This condition implies Condition 1 because every bounded random variable is automatically sub-Gaussian by definition.
+%Our condition is more general than the conditions required in previous work.
+We require that $\wmle_i$ satisfy the SGT Condition.
+This is strictly more general than the conditions in previous work.
+\cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded (in addition to a number of conditions on the moments of the likelihood).
+A bounded parameter space implied that $\wmle_i$ satisfies the SGT Condition because every bounded random variable is automatically sub-Gaussian by definition.
 
@@
 \label{thm:wave}
-Assume Condition 1 holds.
+Assume $\wmle_i$ satisfies the SGT Condition.
 Let $t>0$.
@@
 Note that this reduction in variance is essentially optimal.
-If we apply Condition 2 to the estimator $\wmle$ instead of $\wmle_i$,
-then we are assuming that the variance of $\wmle$ reduces at this same rate.
+If we assume that $\wmle$ satisfies the SGT Condition,
+then we are assuming that the variance of $\wmle$ reduces at the same $O(1/\sqrt{mn})$ rate.
 %By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
@@
 
-%Now we describe our main result.
-%First we introduce some notation.
-%Let $\proj{\Wowa}$ be the projection operator onto the space $\Wowa$.
-For our next results, we need to introduce the following notation.
-If $\A$ is an arbitrary space,
-then the projection of a point $\x$ onto $\A$ is denoted by
-\begin{equation}
-\proj{\A}\x = \argmin_{\mathbf{a}\in\A} \ltwo{\mathbf{a}-\x}
-\end{equation}
-Recall that $\W$ is the vector space spanned by the $\wmle_i$s,
-and that $\wowa$ is the maximum likelihood estimator when the parameter space is restricted to $\W$.
-Our next result shows that as the number of machines increases,
-the distance between $\W$ and $\wstar$ decreases.
+Unlike the $\wave$ estimator,
+$\wowa$ reduces both bias and variance.
+The $m$ and $\nreopt$ parameters act as ``knobs'' that let us trade off lower statistical error for higher communication cost.
+%In particular, the number of machines $m$ acts as a ``knob'' that lets us
+%In this section, we present one main theorem and two supporting lemmas.
+%We begin our discussion with the following informal theorem statement.
+Before we present the formal analysis,
+we present a motivating informal analysis.
+
+\subsubsection{Informal Analysis}
+%The following theorem statement ignores constant factors.
+%Our main result can be stated with the following informal theorem.
+We first state our main result,
+then describe its implications.
+\newtheorem*{theoreminf}{Theorem 2 (Informal)}
+\begin{theoreminf}
+With high probability, we have that
+\begin{equation}
+\ltwo{\wstar-\wowa}
+\le
+O\left(
+    \sqrt{\frac{1}{m\nreopt}}
+    +
+    \sqrt{\left(1-\frac{m}{d}\right)\ltwo{\wstar-\wave}}
+\right)
+\label{eq:informal}
+\end{equation}
+\label{thm:informal}
+\end{theoreminf}
+The rightmost term captures the error of the optimal parameter vector constrained to $\W$.
+That is, if
+\begin{equation}
+\wowastar = \argmin_{\w\in\W} \int_{\X\times\Y} \left(f(y,\trans\x\w) + \lambda R(\w)\right) d(\x,y)
+,
+\end{equation}
+then the rightmost term is the error of $\ltwo{\wstar-\wowastar}$.
+Clearly, as $m \to d$, the space $\W \to \Theta$, so $\wowastar\to\wstar$, and $\ltwo{\wstar-\wowastar}\to 0$.
+
+The leftmost term above captures the error due to using only $\nreopt$ data points in the second round of optimization.
+That is, the left most term captures the error $\ltwo{\wowa-\wowastar}$.
+As $\nreopt\to n$, this error approaches the error of the oracle estimator trained on all the data $\ltwo{\wstar-\wmle}$.
+By increasing $m$ and $\nreopt$ we get an estimator that is closer to the sequential oracle
+but at the expense of a higher communication cost.
+
+\subsubsection{Formal Analysis}
+
+In this section we prove a version of Theorem 2 with explicit constant factors.
+We begin by introducing the following notation.
+The orthogonal projection of a point $\x$ onto the space $\W$ is denoted by
+\begin{equation}
+\proj{\W}\x = \argmin_{\w\in\W} \ltwo{\w-\x}
+.
+\end{equation}
+Our proof is divided into two lemmas.
+The most interesting work happens in the first lemma,
+wherein we bound $\ltwo{\wstar-\proj\W\wstar}$.
 
 \begin{lemma}
-Assume Condition 1 holds.
-Let $t>0$.
-Then we have that with probability at least $1-\exp((d-m)(-t+\ln(t+1)))$,
+\label{lem:proj}
+Assume $\wmle_i$ satisfies the SGT Condition.
+Let $t>0$ and
+\begin{equation}
+\delta = 1-\exp((d-m)(-t+\ln(t+1)))
+.
+\label{eq:delta}
+\end{equation}
+Then we have that with probability at least $\delta$,
 \begin{equation}
@@
 \end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Our next lemma bounds the error of $\wowa$ in terms of the error of other estimators whose error we know.
+%We introduce the following notation.
+
+\begin{lemma}
+%Assume that $\wmle_i$ satisfies Condition 1.
+%Let $\delta$ be defined as in Equation \ref{eq:delta} from Lemma \ref{lem:proj}.
+%Then with probability at least $1-\delta$,
+%\begin{equation}
+%\ltwo{\wstar-\wowa}
+%\le
+%\ltwo{\wowastar-\wowa}
+%+
+%\qlo^{-1} \qhi (
+%)
+%.
+%\end{equation}
+Let $\qhi,\qlo : \mathbb{R}^+ \to \mathbb{R}^+$ be monotonically increasing functions such that for all points $\w\in\Theta$,
+\begin{align}
+F(\wstar) - F(\w) &\ge \qlo \left( \ltwo {\wstar - \w} \right)
+,
+\\
+F(\wstar) - F(\w) &\le \qhi \left( \ltwo {\wstar - \w} \right)
+.
+\end{align}
+Then we have that
+\begin{equation}
+\ltwo {\wstar-\wowastar}
+\le
+\qlo^{-1} \left(
+    \qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
+\right)
+.
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+By the triangle inequality, we have that
+\begin{align}
+\ltwo{\wstar-\wowa}
+\le
+\ltwo{\wstar-\wowastar} + \ltwo{\wowastar-\wowa}
+.
+\end{align}
+We bound the left hand term as follows.
+By the definition of $\qhi$ and $\qlo$, we have
+\begin{align}
+\qlo \left( \ltwo {\wstar-\wowastar} \right)
+&\le F(\wstar) - F(\wowastar)
+\\
+&\le F(\wstar) - F(\proj\W\wstar)
+\\
+&\le \qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
+.
+\end{align}
+%and so
+%\begin{equation}
+%\ltwo {\wstar-\wowastar}
+%\le
+%\qlo^{-1} \left(
+    %\qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
+%\right)
+%.
+%\end{equation}
+%Applying Lemma \ref{lem:proj} gives the stated result.
+Applying $\qlo^{-1}$ to both sides gives the stated result.
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We are now ready to state a formal version of the informal Theorem \ref{thm:informal}.
+
+\newtheorem*{theoremformal}{Theorem 2 (Formal)}
+\begin{theoremformal}
+Assume that both $\wmle_i$ and $\alpha$ satisfy the SGT Condition.
+Let $t>0$ and
+\begin{equation}
+\delta = (1-\exp((d-m)(-t+\ln(t+1))))(1-\exp(-\sigma^2t^2/2))
+.
+\end{equation}
+Then, with probability at least $\delta$,
+\begin{equation}
+\begin{aligned}
+%\ltwo{\wstar-\wowa}
+%&\le
+%{\wstar-\E\w}
+\end{aligned}
+\end{equation}
+\end{theoremformal}
 

### 2016-10-12 paper.tex
lost of changes

@@
 \pgfplotsset{width=7cm,compat=1.8}
+\definecolor{darkgreen}{RGB}{0,127,0}
 
@@
 \newtheorem{assumption}{Condition}
+\newtheorem{defn}{Definition}
 
@@
 \newcommand{\Zproj}{Z^{\textit{proj}}}
-\newcommand{\Zreopt}{Z^{\textit{owa}}}
-\newcommand{\nreopt}{n^{\textit{owa}}}
+\newcommand{\Zowa}{Z^{\textit{owa}}}
+\newcommand{\nowa}{n^{\textit{owa}}}
 \newcommand{\Q}{\mathcal{Q}}
@@
 \newcommand{\ahat}{\hat\alpha}
+\newcommand{\astar}{\alpha^*}
 \newcommand{\afull}{\ahat^{\textit{full}}}
@@
 \newcommand{\trans}[1]{\ensuremath{{#1}^{\mathsf{T}}}}
+\newcommand{\pinv}[1]{\ensuremath{{#1}^{\mathsf{\dagger}}}}
 \newcommand{\ltwo}[1]{{\lVert {#1} \rVert}}
@@
 
+%\newcommand{\plots}[1]{}
+\newcommand{\plots}[1]{#1}
 \newcommand{\ignore}[1]{}
@@
   author={Vershynin, Roman},
-  journal={arXiv preprint arXiv:1011.3027},
   chapter=5,
@@
 %\aistatstitle{Instructions for paper submissions to AISTATS 2015}
-\aistatstitle{Communication-Efficient Distributed Maximum Likelihood Estimation with the Optimal Weigted Average}
+\aistatstitle{Communication-Efficient Distributed Maximum Likelihood Estimation with the Optimal Weighted Average}
 
@@
 \begin{abstract}
-We present a distributed algorithm for maximum likelihood estimation (MLE) with low communication complexity.
-Our algorithm is an easy to implement modification of the standard averaging method.
-In the averaging method, each machine in a cluster first solves the MLE problem on a subset of the data;
-the resulting parameter vectors are then averaged together.
-In our method, we perform a second round of optimization to find the \emph{optimal weighted average} of the parameter vectors.
-This second optimization happens in a low dimensional space,
-and so it adds negligible computational and communication costs.
-%Given $m$ machines (each with $n$ data points of dimension $d$),
-%we show that our algorithm reduces bias at a rate of $O(\frac{1-m/d}{n})$.
-Whereas the averaging method reduces only the variance of the estimator,
-our method reduces both the bias and the variance.
-Our analysis is more general than the analysis of similar communication efficient algorithms that also reduce the bias.
+We present a distributed algorithm for maximum likelihood estimation (MLE) with low communication cost. % called optimal weighted averaging (OWA).
+In our algorithm, each machine in a cluster first independently solves the MLE problem on a subset of the data;
+a second round of optimization then finds the optimal weighted average of the parameter vectors.
+%In the first step, each machine independently solves the MLE problem on a local subset of the data.
+%Then a second round of optimization finds the optimal weighted average of the parameter vectors from step 1.
+%The resulting parameter vectors are then merged together
+%OWA proceeds in two rounds.
+%In the first round, each machine solves the MLE problem on a local subset of the data.
+%In the second round, each machine solves a modified MLE problem.
+%The number of machines $m$
+%We then perform a second round of optimization to find the optimal weighted average of the results.
+%Our algorithm is an easy to implement improvement to the standard averaging method.
+%In the averaging method, each machine in a cluster first solves the MLE problem on a subset of the data;
+%the resulting parameter vectors are then averaged together.
+%In our method, we perform a second round of optimization to find the \emph{optimal weighted average} of the parameter vectors.
+%This second optimization happens in a low dimensional space,
+%and so it adds negligible computational and communication costs.
+%Whereas the averaging method reduces only the variance of the estimator,
+%our method reduces both the bias and the variance.
+%A major advantage of OWA is that two tunable parameters let the user trade better statistical p
+We show analytically that this method reduces both the bias and variance of the estimator when compared to the estimator trained by a single machine.
+(The straight average is known to reduce only the variance.)
+Our analysis relies on simple properties of sub-Gaussian random vectors.
+It is therefore simpler and more general than the analysis of similar distributed algorithms.
 Notably, we do not assume (as do all previous analyses) that the likelihood function is concave or that any quantities are bounded.
-In practice,
-this bias reduction makes our algorithm more robust to the choice of regularization parameter,
-speeding up the model selection process.
+A major practical advantage of our method is that it is robust to the amount of regularization,
+which speeds up model selection.
 \end{abstract}
@@
 To analyze these datasets, we need distributed algorithms.
-Existing distributed algorithms can be classified as either interactive or non-interactive.
+Existing distributed algorithms can be classified as either interactive or non-interactive depending on their communication complexity.
 In this paper we propose an algorithm that exhibits the benefits of both types.
 
-\emph{Interactive algorithms} require many rounds of communication between machines.
+\emph{Interactive} algorithms require many rounds of communication between machines.
 These algorithms often resemble standard iterative algorithms where each iteration is followed by a communication step.
 The appeal of interactive algorithms is that they enjoy the same statistical regret bounds as standard sequential algorithms.
-Unfortunately, these algorithms can be too slow in practice because communication is the main bottleneck in modern distributed architectures,
-and these algorithms require many rounds of communication.
+But there are two downsides.
+First, these algorithms can be too slow in practice because communication is the main bottleneck in modern distributed architectures.
+Second, these algorithms require special implementations and do not work with off-the-shelf statistics libraries provided by (for example) Python, R, and Matlab.
 
 \emph{Non-interactive} algorithms require only a single round of communication.
-They are significantly faster than interactive algorithms
-but have worse regret bounds.
+They are significantly faster than interactive algorithms and  easily implemented with standard libraries.
+The downside is worse regret bounds.
 Recent work (discussed in Section \ref{sec:bounds}) has shown that no non-interactive algorithm can achieve regret bounds comparable to an interactive one.
-The popular large scale learning system Vowpal Wabbit combines both types of algorithms to achieve good real world performance.
-A non-interactive algorithm is run first to provide a good warm start to a subsequent interactive algorithm \citep{vw,agarwal2014reliable}.
-%Synonyms for non-interactive algorithms used in the literature include \emph{one-shot distributed learning} and \emph{communication efficient learning}.
 
@@
 so it is not subject to the existing regret bounds of non-interactive algorithms.
-The amount of information communicated in the second round is small, however,
-so our algorithm retains the speed advantages of non-interactive algorithms.
-
-In the next section, we formally describe our problem setting and proposed solution.
-In Section 3, we compare our solution to related solutions.
+The algorithm has two tunable parameters that let the user trade better statistical performance for worse communication complexity.
+%The amount of information communicated in the second round is small, however,
+%so our algorithm retains the speed advantages of non-interactive algorithms.
+The OWA algorithm is easily implemented in a MapReduce architecture with standard packages.
+
+%The popular large scale learning system Vowpal Wabbit combines both types of algorithms to achieve good real world performance.
+%A non-interactive algorithm is run first to provide a good warm start to a subsequent interactive algorithm \citep{vw,agarwal2014reliable}.
+
+In the next section, we formally describe our OWA algorithm.
+In Section 3, we compare OWA to existing distributed algorithms.
 We highlight how the analysis of existing algorithms requires more limiting assumptions than our own,
-and show in detail why most existing non-interactive regret bounds do not apply to our problem.
-Section 4.1 presents a novel, very general analysis of the averaging estimator.
-Section 4.2 shows that two parameters to our algorithm $m$ and $\nreopt$ act as ``knobs'' that let the user easily trade off statistical performance for communication cost.
+and show in detail why existing non-interactive regret bounds do not apply to OWA.
+%Section 4.1 presents a novel, more general analysis of the widely used averaging estimator.
+%Section 4.2 shows that OWA's regret bounds interpolates between the averaging estimator's regret and the optimal regret.
+Section 4 shows that OWA's regret bounds interpolate between the averaging estimator's regret and the optimal regret.
+As part of the analysis, we provide novel, more general regret bounds for the averaging estimator.
 Section 5 shows experimentally that our algorithm performs well.
-
-\section{Problem Statement and Solution}
+We emphasize that our algorithm is robust to the strength of regularization,
+which is one of the reasons it performs well in practice.
+
+\section{THE OWA ALGORITHM}
 %\section{Problem Setting and Our Solution}
@@
 
+This section first formally introduces the problem of communication efficient distributed estimation,
+then describes our proposed OWA distributed estimator.
+
+\subsection{Problem Setting}
+
 Let $\Y\subseteq\mathbb{R}$ be the space of response variables,
-$X\subseteq\mathbb{R}^d$ be the space of the covariates,
+$X\subseteq\mathbb{R}^d$ be the space of covariates,
 and $\Theta\subseteq\mathbb{R}^d$ be the parameter space.
@@
 and $\lambda\in\mathbb{R}$ be the regularization strength.
-Then the maximum likelihood estimator is
+Then the regularized maximum likelihood estimator (MLE) is
 \begin{equation}
@@
 \end{equation}
+In the remainder of this paper, it should be understood that all MLEs are regularized.
+
 Assume that $Z$ has been partitioned onto $m$ machines so that each machine $i$ has dataset $Z_i$ of size $n$, and all the $Z_i$ are disjoint.
-Then each machine calculates the local estimator
+Then each machine calculates the local MLE
 \begin{equation}
@@
 Our goal is to merge the $\wmle_i$ into a single improved estimate.
-The baseline example is the averaging estimator
+%A non-interactive distributed estimator depends on the $\wmle_i$s and nothing else.
+A baseline merging procedure is the averaging estimator
 \begin{equation}
@@
 \end{equation}
-This estimator is well studied, and in Section \ref{sec:anal} we compare this previous work to our own.
+This estimator is well studied, and in Section \ref{sec:relwork} we compare this previous work to our own.
 Here we briefly recall that the analysis of an estimator's error $\ltwo{\wstar-\wave}$ is typically decomposed into bias $\ltwo{\wstar - \E\wave}$ and variance $\ltwo{\E\wave-\wave}$ components.
@@
 To motivate our estimator,
-we first describe a simpler estimator that uses the entire dataset for the second round of optimization.
+we first present an estimator that uses the entire dataset for the second round of optimization.
 %The idea of our proposed solution is to find the best linear combination of the $\wmle_i$s using a second round of optimization.
@@
 \end{equation}
-%This is the maximum likelihood estimator when we restrict our
-Notice also that $\wowafull$ is the maximum likelihood estimator when the parameter space $\Theta$ is restricted to $\W = \vecspan \{\wmle_i\}_{i=1}^m$.
+Notice that $\wowafull$ is the maximum likelihood estimator when the parameter space $\Theta$ is restricted to $\W = \vecspan \{\wmle_i\}_{i=1}^m$.
 In other words, the $\afull$ vector contains the optimal weights to apply to each $\wmle_i$ when averaging.
@@
     dot/.style = {minimum width=0.15cm,inner sep=0pt,line width=0pt,fill,circle,black,font=\small}
+    , yscale=0.65
     ]
+\small
 \draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.5,-0.75) (-1.05,1) (-0.1,-0.1) (0.75,-0.5) (0.45,-0.7) };
@@
 \node[dot] (wave) at (3.8,1.25) {};
-\node at (4.0,1.6) {$\wave$};
+\node at (4.0,1.7) {$\wave$};
 \node[dot] (wave) at (2.80,1.40) {};
-\node at (3.1,1.75) {$\wmle_1$};
+\node at (3.1,1.85) {$\wmle_1$};
 \node[dot] (wave) at (4.75,1.10) {};
-\node at (4.9,1.45) {$\wmle_2$};
+\node at (4.9,1.55) {$\wmle_2$};
 %\node[dot] (wave) at (3.8,1.25) {};
@@
 \node[dot] at (-1.35,2.03) {};
-\node at (-0.9,2.4) {$\wowafull$};
+\node at (-1.0,2.5) {$\wowafull$};
 
 \node[dot] at (-0.55,1.9) {};
-\node at (-0.1,2.1) {$\wowa$};
+\node at (-0.1,2.3) {$\wowa$};
 
@@
 \node at (5.65,1.0) {$\W$};
-
 \end{tikzpicture}
+\vspace{-0.35in}
 \caption{
     Our method performs a second round of optimization to find the best parameter vector in $\W$.
-    Since $\wave\in\W$, the estimator $\wowafull$ has higher empirical likelihood than $\wave$.
+    %Since $\wave\in\W$, the estimator $\wowafull$ has higher empirical likelihood than $\wave$.
     Since $\W$ has low dimension, we can use relatively few data points in the second round of optimization to ensure that with high probability $\wowa$ has higher empirical likelihood than $\wave$.
+    %Since $\W$ has low dimension, we can use relatively few data points in the second round of optimization to ensure that $\ltwo{\wowafull-\wowa}$ is small.
 }
@@
 %(This intuition is formalized in Section \ref{sec:anal}.)
-This intuition motivates the following estimator.
-Let $\Zreopt_i\subset Z_i$ be a set of $\nreopt$ data points uniformly sampled from $Z_i$ without replacement,
-and let $\Zreopt$ be the union of the $\Zreopt_i$s.
+This intuition motivates the OWA estimator.
+Let $\Zowa_i\subset Z_i$ be a set of $\nowa$ data points uniformly sampled from $Z_i$ without replacement,
+and let $\Zowa$ be the union of the $\Zowa_i$s.
 Then the OWA estimator is defined as
@@
 \label{eq:ahat}
-\ahat = \argmax_\alpha \sum _{(\x,y)\in \Zreopt} f\left(y,\trans\x \matW \alpha \right)
+\ahat = \argmax_\alpha \sum _{(\x,y)\in \Zowa} f\left(y,\trans\x \matW \alpha \right)
 + \lambda R (\matW\alpha)
@@
 Since the parameter vector has $d$ dimensions and there are $m$ machines, a total of $O(md)$ bits are transmitted.
-In the second round, each machine projects its local dataset $\Zreopt_i$ onto the space $\W$.
+In the second round, each machine projects its local dataset $\Zowa_i$ onto the space $\W$.
 These projected data points are then transmitted to a predesignated master machine.
-The projected data points each have dimension $m$, so a total of $O(m^2\nreopt)$ bits are transmitted.
+The projected data points each have dimension $m$, so a total of $O(m^2\nowa)$ bits are transmitted.
 The master machine now has all of the information to complete the optimization.
+In total, $O(md + m^2\nowa)$ bits were transmitted.
+When $d<m\nowa$, this is the same order of bits as the averaging estimator.
 
@@
 \begin{algorithmic}
-\State Each machine $i$ independently
+\State Round 1, each machine $i$ independently:
 \State ~~~~~loads dataset $Z_i$
 \State ~~~~~calculates $\wmle_i$
-\State All machines broadcast $\wmle_i$ to all other machines
-\State Each machine $i$ independently
+\State ~~~~~broadcasts $\wmle_i$ to all other machines
+\State Round 2, each machine $i$ independently:
 \State ~~~~~constructs $\matW$
-\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
-\State ~~~~~calculates $\Zproj_i=\{(\trans\x\matW,y) : (\x,y)\in\Zreopt\}$
-\State All machines broadcast $\Zproj_i$ to a master machine
+\State ~~~~~randomly selects a dataset $\Zowa_i\subset Z_i$
+\State ~~~~~calculates $\Zproj_i=\{(\trans\x\matW,y) : (\x,y)\in\Zowa\}$
+\State ~~~~~broadcasts $\Zproj_i$ to a master machine
 \State The master calculates $\wowa$
@@
 
-\subsection{Implementation Tips}
-
+%\subsection{Implementation Tips}
+%
 Equations \ref{eq:afull} and \ref{eq:ahat} cannot be solved directly using off the shelf optimizers because existing optimizers do not support the non-standard regularization term $R(\matW\alpha)$.
@@
 .
+\label{eq:approxreg}
 \end{equation}
@@
 and the L2 regularized problem is much easier to solve than the L1 problem.
-The new $\lambda_2$ regularization parameter should be set by cross validation.
-This will be a fast procedure, however, because there are few data points to optimize,
-and the L2 regularized problem is much easier to solve than the L1 problem.
 With this minor modification, our distributed estimator can be implemented using any existing optimizer.
-
-It is also important when performing the second round of optimization to not extend the parameter vector with an intercept term as is typically done.
-It is fine to do this for the first round as long as the intercepts also get scaled.
+%Our experimental results in Section \ref{sec:exp} use this modified version of $\hat\alpha$.
+
+%It is also important when performing the second round of optimization to not extend the parameter vector with an intercept term as is typically done.
+%It is fine to do this for the first round as long as the intercepts also get scaled.
 
 \section{RELATED WORK}
+\label{sec:relwork}
 
@@
 \end{equation}
-
 Previous analysis of $\wave$ makes a number of limiting assumptions.
-\cite{mcdonald2009efficient} gave the first analysis of this estimator in the special case of L2 regularized maximum entropy models.
+\cite{mcdonald2009efficient} analyze $\wave$ in the special case of L2 regularized maximum entropy models.
 They provide tail bounds on $\ltwo{\wstar-\wave}$, showing that the variance $\ltwo{\E\wave-\wave}$ reduces as $O((nm)^{-1/2})$,
-but they could not show a reduction in bias.
+but they do not show a reduction in bias.
 Their analysis uses a martingale technique that requires the radius of the dataset be independent of the size of the dataset.
@@
 Other research has focused on modifications to the $\wave$ estimator to reduce bias.
-\cite{zinkevich2010parallelized} showed that if the training sets partially overlap each other (instead of being disjoint), then the resulting estimator will have lower bias.
-The authors admit their analysis is ``highly technical.''
-\cite{zhang2012communication} also provided a bootstrap average estimator,
+\cite{zinkevich2010parallelized} show that if the training sets partially overlap each other (instead of being disjoint), then the resulting estimator will have lower bias.
+%The authors admit their analysis is ``highly technical.''
+\cite{zhang2012communication} also provide a bootstrap average estimator,
 which works as follows.
-Let $r\in(0,1)$, and $Z_i^r$ be a bootstrap sample of $Z_i$ with $rn$ data points.
+Let $r\in(0,1)$, and $Z_i^r$ be a bootstrap sample of $Z_i$ of size $rn$.
 Then the bootstrap average estimator is
 \begin{equation}
+\wboot = \frac{\wave-r\waver}{1-r}
+,
+\end{equation}
+where
+\begin{equation}
 \begin{aligned}
-\wmler_i &= \argmax_\w \sum_{(\x,y)\in Z_i^r} f(y,\trans\x\w) + \lambda R(\w)
-,
-\\
 \waver &= \frac{1}{m}\sum_{i=1}^m \wmler_i
@@
 \\
-\wboot & = \frac{\wave-r\waver}{1-r}
-.
+\wmler_i &= \argmax_\w \sum_{(\x,y)\in Z_i^r} f(y,\trans\x\w) + \lambda R(\w)
+.
+%\\
+%,
+%\\
+%\wboot & = \frac{\wave-r\waver}{1-r}
+%.
 \end{aligned}
@@
 First, the optimal value of $r$ is not obvious and setting the parameter requires cross validation on the entire data set.
-Our proposed $\wowa$ estimator has a similar parameter $\lambda_2$ that also needs tuning,
+Our proposed $\wowa$ estimator has a similar parameter $\lambda_2$ that needs tuning,
 but this tuning happens on a small fraction of the data and always with the L2 regularizer.
@@
 
-Surprisingly, \cite{zhang2013divide} showed that in the special case of kernel ridge regression,
-a reduction in bias was not needed to have the MSE of $\wave$ decay at the optimal sequential rate.
+Surprisingly, \cite{zhang2013divide} show that in the special case of kernel ridge regression,
+a reduction in bias is not needed to have the MSE of $\wave$ decay at the optimal sequential rate.
 By a careful choice of regularization parameter,
@@
 This suggests that a merging procedure that reduces bias is not crucial to good performance if we set the regularization parameter correctly.
-Typically there is a narrow range of good regularization parameters, and finding a $\lambda$ in this range is expensive computationally.
-We show experimentally in Section \ref{sec:exp} that our method has significantly reduced sensitivity to the regularization parameter.
-We consider this to be the main practical strength of our method.
+Typically there is a narrow range of good regularization parameters,
+and finding a $\lambda$ in this range is expensive computationally.
+We show experimentally in Section \ref{sec:exp} that our method has significantly reduced sensitivity to $\lambda$.
+It is computationally cheaper to find a good $\lambda$ for our method than for the other methods discussed in this section.
 
@@
 Performance bounds come in two flavors: statistical and information theoretic.
-On the statistical side, \cite{liu2014distributed} showed that for any non-interactive distributed estimator $\wq$,
+On the statistical side, \cite{liu2014distributed} show that for any non-interactive distributed estimator $\wq$,
 the quantity $\ltwo{\wq-\wmle}{}^2$ decays as $\Omega(\gamma^2_\wstar \I^{-1}_\wstar/n^2)$.
@@
 Furthermore, they show that $\wkl$ matches this bound.
-This bound is not relevant for our estimator because of our semi-interactive setting.
-A crucial assumption of Liu and Ihler's analysis is that the merge function not depend on the data as our merge function does.
+This bound is not relevant for our $\wowa$ estimator because of our semi-interactive setting.
+A crucial assumption of Liu and Ihler's analysis is that the merge function not depend on the data.
 %Furthermore, we directly optimize $\ltwo{\wq-\wstar}^2$ which is a more useful quantity in practice.
@@
 \cite{shamir2014fundamental}, \cite{zhang2013information}, and \cite{garg2014communication} all provide information theoretic lower bounds on the sample complexity of non-interactive learning problems.
-As above, however, there results are not applicable in our semi-interactive setting.
+As above, however, their results are not applicable in our semi-interactive setting.
 %provides general information theoretic bounds for what he terms learning with information constraints.
@@
 %and \cite{garg2014communication} improved this lower bound to $\Omega(mk)$.
-Surprisingly, \cite{braverman2015communication} showed that the minimax optimal error rate for least squares regression requires $\Omega(m\cdot\min\{n,d\})$ bits of communication even in the fully interactive setting.
+Surprisingly, \cite{braverman2015communication} show that the minimax optimal error rate for least squares regression requires $\Omega(m\cdot\min\{n,d\})$ bits of communication (independent of $k$) even in the fully interactive setting.
 This is important because sparsity does not reduce the amount of communication required, and this bound does apply in our setting.
@@
 
-%\begin{assumption}
-%The Fisher information matrix $\I_\wstar$ is positive definite.
-%\end{assumption}
-%
-%Condition 1 above implies that the inverse $\I_\wstar^{-1}$ and square root $\I_\wstar^{1/2}$ exist.
-
-%That is, we assume there exists a constant $c$ such that
+\begin{defn}
+We say that a random vector $\x$ is \emph{sub-Gaussian} with variance proxy $\sigma^2$ if it obeys the following concentration bound.
+Let $t>0$, then with probability at least $1-\exp(-\sigma^2t^2/2)$,
 %\begin{equation}
-%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{\frac{d}{n}}} \le \exp\left(-ct^2\right)
-%%\prob{\ltwo{\wmle_i-\wstar} > t\sqrt{d/n}} \le \exp\left(-ct^2\right)
-%%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(-\sigma^2t^2/2\right)
-%\prob{\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} > t\sqrt{d/n}} \le \exp\left(\frac{-\sigma^2t^2}{2}\right)
-%.
+$
+\ltwo{\x} < t
+$.
 %\end{equation}
+\end{defn}
+
+Note in particular that if $\x$ is a Gaussian random vector with mean $\mu$ and covariance $\Sigma$,
+then $\Sigma^{1/2}(\x-\mu)$ is sub-Gaussian with $\sigma^2=1$.
+Sub-Gaussian random vectors have recently become an important tool in the analysis of high dimensional statistics.
+\cite{vershynin2010introduction} provides an accessible tutorial of these results.
+We now state our condition.
 
@@
 Let $\theta$ be an estimator trained on $n$ data points.
-Let $t>0$ and $\sigma^2$ be a constant called the \emph{variance proxy}.
-Then with probability at least $1-\exp(-\sigma^2t^2/2)$,
-\begin{equation}
-%\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} < \frac{t}{\sqrt n}
-%\sqrt n \ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < t
-\sqrt n \ltwobig{\I^{-1/2}_\wstar(\theta-\E\theta)} < t
-,
-\end{equation}
-where $\I_\wstar$ is the positive definite Fisher information matrix at the parameter vector's true value $\wstar$.
+%Let $t>0$ and $\sigma^2$ be a constant called the \emph{variance proxy}.
+%Then with probability at least $1-\exp(-\sigma^2t^2/2)$,
+%\begin{equation}
+%%\ltwobig{\I^{-1/2}_\wstar(\wmle_i-\wstar)} < \frac{t}{\sqrt n}
+%%\sqrt n \ltwobig{\I^{-1/2}_\wstar(\wmle_i-\E\wmle_i)} < t
+%\sqrt n \ltwobig{\I^{-1/2}_\wstar(\theta-\E\theta)} < t
+%,
+%\end{equation}
+Then the random vector
+\begin{equation}
+%\Delta_\theta = \sqrt{n}\ltwo{\I^{1/2}_\wstar(\theta-\E\theta)}
+\Delta_\theta = \sqrt{n}{\I^{1/2}_\wstar(\theta-\E\theta)}
+\end{equation}
+is sub-Gaussian for some $\sigma^2$.
+Above, $\I_\wstar$ is the positive definite Fisher information matrix at the parameter vector's true value $\wstar$,
+which is sub-Gaussian by assumption.
 \end{sgt}
 
-This is a mild condition that is known to hold in many situations of interest.
+%Sub-Gaussian distributions have recently become an important tool in the analysis of high dimensional statistics.
+%\cite{vershynin2010introduction} provides an accessible tutorial of these results.
+
+The SGT condition is mild and known to hold in many situations of interest.
 In the asymptotic regime when $n\to\infty$,
@@
 %In this case,
-The variance proxy $\sigma^2=1$,
-and the distribution of the estimation error is truly Gaussian.
-Theorem 7.5.2 of \cite{lehmann1999elements} provides a particularly simple exposition.
+%The error $\Delta_\theta$ is exactly Gaussian.
+Theorem 7.5.2 of \cite{lehmann1999elements} is a simple example that shows $\Delta_\theta$ is an isotropic centered Gaussian.
 Lehman's theorem requires only that $f$ be three times differentiable,
@@
 and some mild identifiability conditions.
-More sophisticated analyses show that Condition 1 holds very generally.
-For example, the i.i.d. assumption is not strictly necessary.
-See for example \cite{spokoiny2012parametricestimation}.
+More sophisticated analyses show that the SGT Condition holds very generally.
+For example \cite{spokoiny2012parametricestimation} shows that $\Delta_\w$ is normally distributed even when the data points are correlated.
 
@@
 but known results are less strong.
-Most non-asymptotic results of this form require that the data points be sub-Gaussian.
+Most non-asymptotic results of this form require that the data points also be sub-Gaussian.
 For example, \cite{negahban2009unified} considers the case when the data points are sub-Gaussian, the likelihood satisfies a ``restricted strong convexity condition,'' and the regularizer is decomposable.
-More recently, \cite{sivakumar2015beyond} showed Condition 1 holds when the data are only sub-exponential.
-The strongest non-asymptotic results on the estimation error known to the authors is due to \cite{spokoiny2012parametricestimation}.
+More recently, \cite{sivakumar2015beyond} showed the SGT Condition holds when the data are only sub-exponential.
+The strongest non-asymptotic results on the estimation error known to the authors are due to \cite{spokoiny2012parametricestimation}.
 Spokoiny does not require any conditions on the data,
-but can show that Condition 1 is satisfied only up to $t < O(n)$.
-
-%Our condition is more general than the conditions required in previous work.
-We require that $\wmle_i$ satisfy the SGT Condition.
+but can show that the SGT Condition is satisfied only up to $t < O(n)$.
+
+Our analysis below requires that $\wmle_i$ satisfy the SGT Condition.
 This is strictly more general than the conditions in previous work.
 \cite{zhang2012communication} for example require that the parameter space $\Theta$ be bounded (in addition to a number of conditions on the moments of the likelihood).
-A bounded parameter space implied that $\wmle_i$ satisfies the SGT Condition because every bounded random variable is automatically sub-Gaussian by definition.
-
-In conclusion,
-there are many ways to verify that Condition 1 holds.
-Pick whichever one suits your problem.
-%With this condition assumed, our results are elementary.
+A bounded parameter space automatically implies that $\wmle_i$ satisfies the SGT Condition because every bounded random variable is sub-Gaussian by definition.
+
+%In conclusion,
+%there are many ways to verify that the SGT Condition holds.
+%Pick whichever one suits your problem.
 
@@
 
-%The main tool we use is the central limit theorem (CLT) for maximum likelihood estimators.
-%%We then combine this result with standard results in random projections.
-%There are many versions of this theorem,
-%but they all conclude that the variance of an estimator is normally distributed.
-%In particular,
-%\begin{equation}
-%\label{eq:assumption}
-%\E\wmle_i - \wmle_i \sim \normal{\zero}{\I^{-1}_{\E\wmle_i}}
-%\end{equation}
-%where $\I^{-1}_{\E\wmle_i}$ is the Fisher information matrix.
-%The classical version of the CLT applies only as the number of samples $n\to\infty$.
-%Non-classical versions handle the non-asymptotic case.
-%Section \ref{sec:clt} below provides an overview.
-
-%We begin by presenting our theorem on the statistical performance of $\wave$.
 We provide a simple bound that shows that averaging improves the variance,
@@
 but our analysis has the following advantages:
-it requires fewer assumptions, has a simpler proof, and has an easy to interpret constant factor.
-%The main idea of our technique is to show that the variance is normally distributed.
-%The ideas in this proof will be used in Section \ref{sec:analreopt} where we show that the variance of $\wowa$ is the same as $\wave$, but that the bias is reduced.
+It requires fewer assumptions, has a simpler proof, and has an easy to interpret constant factor.
 
@@
 $.
-In words, the bias of the combined averaging estimator is the same as the bias of the estimator on an individual machine.
+%In words, the bias of the combined averaging estimator is the same as the bias of the estimator on an individual machine.
 
@@
 %\\&=
-\ltwobig{\frac{1}{m}\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)}
+\frac{1}{m}\ltwobig{\sum_{i=1}^m\left(\wmle_i-\E\wmle_i\right)}
 \\&=
@@
 %\frac{1}{m\sqrt{n}}\ltwobig{\I^{-1/2}_\wstar\sum_{i=1}^m\sqrt{n}\I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)}
-\frac{1}{m\sqrt{n}}\ltwobig{\I^{-1/2}_\wstar\sum_{i=1}^m\Delta_i}
+\frac{1}{m\sqrt{n}}\ltwobig{\I^{-1/2}_\wstar\sum_{i=1}^m\Delta_{\wmle_i}}
 %\\&\sim
@@
 %&\sim
+.
 \end{align}
-where
-\begin{equation}
-\Delta_i = \sqrt n \I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)
-.
-\end{equation}
-Notice that each of the $\Delta_i$ are i.i.d. centered sub-Gaussian random vectors.
-Centered sub-Gaussian random vectors inherit the following basic property from centered Gaussian random vectors:
+%where
+%\begin{equation}
+%\Delta_i = \sqrt n \I^{1/2}_\wstar\left(\wmle_i-\E\wmle_i\right)
+%.
+%\end{equation}
+Equation \ref{eq:var1} follows by multiplying inside the norm by $1=\sqrt{n}\I_{\wstar}^{1/2}\I_{\wstar}^{-1/2}/\sqrt{n}$ and applying the definition of $\Delta_{\wmle_i}$.
+Notice that each of the $\Delta_{\wmle_i}$ are i.i.d. sub-Gaussian random vectors.
+Sub-Gaussian random vectors inherit the following basic property from centered Gaussian random vectors:
 The sum of $m$ i.i.d. sub-Gaussians has the same distribution as $\sqrt{m}$ times the sub-Gaussian.
-That is, $\sum_{i=1}^m \Delta_i \sim \sqrt{m}\Delta_1$.
+That is, $\sum_{i=1}^m \Delta_{\wmle_i} \sim \sqrt{m}\Delta_1$.
 Substituting into Equation \ref{eq:var1} gives
@@
 \sim
-\frac{1}{\sqrt{mn}}\ltwobig{\I^{-1/2}_\wstar\Delta_1}
+\frac{1}{\sqrt{mn}}\ltwobig{\I^{1/2}_\wstar\Delta_1}
 .
@@
 Note that this reduction in variance is essentially optimal.
-If we assume that $\wmle$ satisfies the SGT Condition,
+If we assume that $\wmle$ satisfies the SGT condition
+(in most cases this follows from $\wmle_i$ satisfying the SGT condition),
 then we are assuming that the variance of $\wmle$ reduces at the same $O(1/\sqrt{mn})$ rate.
@@
 $\wowa$ reduces both bias and variance.
-The $m$ and $\nreopt$ parameters act as ``knobs'' that let us trade off lower statistical error for higher communication cost.
+The $m$ and $\nowa$ parameters act as ``knobs'' that let us trade off lower statistical error for higher communication cost.
 %In particular, the number of machines $m$ acts as a ``knob'' that lets us
@@
 %Our main result can be stated with the following informal theorem.
-We first state our main result,
-then describe its implications.
+%We first state our main result informally,
+%then describe its implications.
+Our main result is captured in the following theorem.
 \newtheorem*{theoreminf}{Theorem 2 (Informal)}
@@
 \begin{equation}
+\nonumber
 \ltwo{\wstar-\wowa}
 \le
-O\left(
-    \sqrt{\frac{1}{m\nreopt}}
+O\!\left(\!\!
+    \sqrt{\frac{1}{m\nowa}}
     +
-    \sqrt{\left(1-\frac{m}{d}\right)\ltwo{\wstar-\wave}}
+    \sqrt{\left(1-\frac{m}{d}\right)}\ltwo{\wstar-\wave}
 \right)
@@
 \end{theoreminf}
-The rightmost term captures the error of the optimal parameter vector constrained to $\W$.
+\vspace{-0.25in}
+The rightmost term above captures the error of the optimal parameter vector constrained to $\W$.
 That is, if
@@
 
-The leftmost term above captures the error due to using only $\nreopt$ data points in the second round of optimization.
-That is, the left most term captures the error $\ltwo{\wowa-\wowastar}$.
-As $\nreopt\to n$, this error approaches the error of the oracle estimator trained on all the data $\ltwo{\wstar-\wmle}$.
-By increasing $m$ and $\nreopt$ we get an estimator that is closer to the sequential oracle
-but at the expense of a higher communication cost.
+The leftmost term above captures the error due to using only $\nowa$ data points in the second round of optimization.
+That is, the leftmost term captures the error $\ltwo{\wowa-\wowastar}$.
+As $\nowa\to n$, this error approaches the error of the oracle estimator trained on all the data $\ltwo{\wstar-\wmle}=O(1/\sqrt{mn})$.
+By increasing $m$ and $\nowa$ our estimator is closer to the sequential oracle
+at the expense of a higher communication cost.
 
@@
 
-In this section we prove a version of Theorem 2 with explicit constant factors.
-We begin by introducing the following notation.
-The orthogonal projection of a point $\x$ onto the space $\W$ is denoted by
-\begin{equation}
-\proj{\W}\x = \argmin_{\w\in\W} \ltwo{\w-\x}
-.
-\end{equation}
-Our proof is divided into two lemmas.
-The most interesting work happens in the first lemma,
-wherein we bound $\ltwo{\wstar-\proj\W\wstar}$.
+In this section we give a version of Theorem 2 with explicit constant factors.
+The proof is divided into two lemmas,
+which we present first because they introduce important notation.
+In the first lemma we bound the distance $\ltwo{\wstar-\proj{\W}\wstar}$,
+where $\proj{\W}$ denotes the orthogonal projection onto $\W$.
+In the second lemma, we show that $\proj{\W}\wstar \approx \wowa$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \label{lem:proj}
-Assume $\wmle_i$ satisfies the SGT Condition.
+Assume $\wmle_i$ satisfies the SGT condition.
 Let $t>0$ and
@@
 =
-(t+1)\sqrt{\smin\smax\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
-\end{equation}
-where $\smin$ and $\smax$ are the minimum and maximum eigenvalues of $\I_\wstar$.
+(t+1)\sqrt{\frac\smin\smax\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
+\end{equation}
+where $\smin$ and $\smax$ are the minimum and maximum eigenvalues of $\I_\wstar^{-1}$.
 \end{lemma}
 
-\begin{proof}
-We have that
-\begin{align}
-%&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
-\ltwo{\wstar-\proj\W\wstar}
-%\nonumber
-%\\
-&=
-\ltwo{(\wstar-\wave) - (\proj\W\wstar - \wave)}
-\nonumber
-\\
-&=
-\ltwo{(\wstar-\wave) - \proj\W(\wstar - \wave)}
-\nonumber
-\\
-&=
-\ltwo{(I-\proj\W)(\wstar - \wave)}
-.
-\label{eq:i-projW}
-\end{align}
-Recall that we defined $\matW$ to be the $d\times m$ matrix with $i$th column equal $\wmle_i$.
-The column space of $\matW$ is $\W$.
-So the projection $\proj\W = \matW\matW^+$, where the superscript $^+$ is the Moore-Penrose pseudoinverse.
-Define the matrix $\matV=\sqrt{n}I^{1/2}\matW$.
-Then the columns of $\matV$ are sub-Gaussian by Condition 1,
-and we can write
-\begin{equation}
-\proj\W
-= \I^{-1/2} \matV \matV ^+ \I^{1/2}
-.
-\label{eq:vvp}
-\end{equation}
-Let $U\Sigma \trans V$ be the singular value decomposition of $\matV$,
-where $U$ and $V$ are orthogonal matrices and $\Sigma$ is the diagonal matrix of singular values.
-By the rotational invariance of the sub-Gaussian distribution,
-$U$ and $V$ are distributed uniformly over the orthogonal group of matrices.
-Substituting into Equation \ref{eq:vvp} gives
-\begin{align}
-\proj\W
-&= \I^{-1/2} (U\Sigma V^T)(U \Sigma V^T) ^+ \I^{1/2}
-\\
-&= \I^{-1/2} (U\Sigma V^T)(V \Sigma^+ U^T) \I^{1/2}
-\\
-&= \I^{-1/2} U\Sigma \Sigma^+ U^T \I^{1/2}
-.
-\label{eq:vvp}
-\end{align}
-Substituting into Equation \ref{eq:i-projW} gives
-\begin{align}
-~~~~~&\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\W\wstar}
-\nonumber
-\\
-&=
-\ltwo{(I-\I^{-1/2} U\Sigma \Sigma^+ U^T \I^{1/2})(\wstar - \wave)}
-\\
-&=
-\ltwo{\I^{-1/2} U(I-\Sigma \Sigma^+ )U^T \I^{1/2}(\wstar - \wave)}
-\\
-&=
-\sqrt{\smin}\ltwo{U(I-\Sigma \Sigma^+ )U^T \I^{1/2}(\wstar - \wave)}
-\\
-&=
-\sqrt{\smin}\ltwo{(I-\Sigma \Sigma^+ )U^T \I^{1/2}(\wstar - \wave)}
-\end{align}
-%The last line above follows from the rotational invariance of the $L2$ norm.
-The operator $(I-\Sigma\Sigma^+)\trans U$ projects a vector onto a randomly chosen subspace of dimension $d-m$.
-Lemma 2.2 from \cite{dasgupta2003elementary} provides a bound for this situation.
-For all $t>0$, with probability at least $1-\exp((d-m)(-t+\ln (t+1)))$,
-\begin{align}
-\ltwo{(I-\Sigma\Sigma^+)\trans U \x} \le (t+1) \sqrt{1-\frac{m}{d}}
-.
-\end{align}
-Substituting gives
-\begin{align}
-~~~~~&\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\W\wstar}
-\nonumber
-\\
-&=
-(t+1)\sqrt{\smin\left(1-\frac{m}{d}\right)}\ltwo{\I^{1/2}(\wstar - \wave)}
-\\
-&=
-(t+1)\sqrt{\smin\smax\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
-\end{align}
-
-\ignore{
-\begin{theorem}
-Assume Condition 1 holds.
-Let $t>0$.
-Then we have that with probability at least $(1-\exp(-t))^2$,
-\begin{equation}
-\ltwo{\wstar-\proj\W\wstar}
-\le
-%b_t\sqrt{\left(1-\frac{m}{d}\right)}
-b_t\sqrt{\left(1-\frac{m}{d}\right)}
-+
-\sqrt\frac{v_{t}}{nm}
-\end{equation}
-where
-%$\lzero{\wave} < d$ is the number of non-zero entries in the $\wave$ vector,
-\begin{equation}
-b_t = \frac{\lambda_{\min}}{\lambda_{\max}}\ltwo{\wstar-\E\wmle_i}\sqrt{t + 1}
-\end{equation}
-%with probability at least $1-\exp (-\frac{b}{2} (\tbias-\ln(\tbias+1)))$,
-$\lambda_{\min}$ and $\lambda_{\max}$ are the minimum and maximum eigenvalues of $\I^{-1}_{\wave}$,
-and $v_t$ is defined as in Theorem \ref{thm:wave}.
-\end{theorem}
-
-This proof follows the same form as the proof for Theorem \ref{thm:wave}.
-We first decompose the error into a bias and variance component;
-then we bound each component separately.
-
-Before we begin, we need to introduce the following vector space:
-\begin{align}
-\V &= \vecspan \{\wmle_i - \wtave\}_{i=1}^m
-%,\\
-%A &= \{ \vv + \E\wave : \vv \in \V \}
-%,\\
-%A &= \{ \vv + \wave : \vv \in \V \}
-%.
-\end{align}
-%We also introduce notation for constructing affine spaces:
-%\begin{equation}
-%\V + \x = \{ \vv + \x : \vv \in \V \}
-%.
-%\end{equation}
-%We now have that
-We now have that:
-\begin{align}
-~~~~&\!\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\W\wstar}
-%&\ltwo{\wstar-\proj\W\wstar}
-\nonumber
-\\
-&\le
-\ltwo{(\wstar-\wtave) - \proj\W(\wstar-\wtave)}
-\\
-&\begin{aligned}
-&\le
-%\ltwo{
-\lVert(\wstar-\wtave)
-\\
-&~~~
-- (\proj\V(\wstar-\wtave) + \wtave - \wave)
-%}
-\rVert
-\end{aligned}
-\\
-&\begin{aligned}
-&\le
-\ltwo{(\wstar-\wtave) - \proj\V(\wstar-\wtave)}
-\\
-&~~~
-+ \ltwo{\wtave - \wave}
-\end{aligned}
-\\
-&=
-\ltwo{(I-\proj\V)(\wstar-\wtave)} + \ltwo{\wtave - \wave}
+The proof uses standard properties of projection matrices with sub-Gaussian columns.
+Due to space limitations, it is included in Appendix A of the supplement.
 %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%
-%\\
-%&\le
-%\ltwo{\wstar-(\proj{\V}\wstar + \wave)}
-%\label{eq:step1}
-%\\
-%&\!\begin{aligned}
-%&\le
-%\ltwo{\wstar-(\proj{\V}\wstar+\wtave)}
-%\\&~~~
-%+
-%\ltwo{(\proj{\V}\wstar+\wtave)-(\proj{\V}\wstar+\wave)}
-%\end{aligned}
-%\label{eq:step2}
-%\\
-%&=
-%\ltwo{\wstar-(\proj{\V}\wstar+\wtave)}
-%+
-%\ltwo{\wtave-\wave}
-%\label{eq:step3}
-\end{align}
-Inequality \ref{eq:step1} follows because $(\proj\V\wstar + \wave) \in \W$.
-Inequality \ref{eq:step2} follows from the triangle inequality.
-The rightmost term in Equation \ref{eq:step3} is the variance of $\wave$,
-which we previously bounded with high probability in Equation \ref{eq:ct1}.
-
-The affine space $\V+\wave$ is a subset of $\W$.
-So,
-\begin{align}
-\ltwo{\wstar-\proj\W\wstar}
-&\le
-\ltwo{\wstar-\proj{(\V+\wave)}\wstar}
-\end{align}
-By the triangle inequality, we have
-\begin{align}
-%\\
-&\begin{aligned}
-\ltwo{\wstar-\proj\W\wstar}
-&\le
-\ltwo{\wstar-\proj{(\V+\E\wave)}\wstar}
-\\&~~~~
-+
-\ltwo{\proj{(\V+\E\wave)}\wstar-\proj{(\V+\wave)}\wstar}
-\end{aligned}
-%\\&+
-%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
-\end{align}
-The left term of the inequality above can be thought of as akin to the bias of the ``estimator'' $\proj\Waff\wstar$,
-and the right term can be thought of as akin to the variance.
-
-We begin by bounding the right term.
-We have that with probability at least $1-\exp(-t)$,
-\begin{align}
-&\ltwo{\proj{(\V+\wtave)}\wstar-\proj{(\V+\wave)}\wstar}
-\\
-&=
-\ltwo{(\proj{(\V+\wave)}\wstar + \wave - \wtave) -\proj{(\V+\wave)}\wstar}
-%\ltwo{(\proj\Waff\wstar+\wave-\wtave)-\proj\Waff\wstar}
-\\&=
-\ltwo{\wave-\wtave}
-\\&\le
-\sqrt\frac{v_t}{mn}
-\end{align}
-%The first equality follows because the space $\WaffE$ is a translation of $\Waff$ by the vector $(\wave-\wtave)$.
-The probabilistic inequality is due to Equation \ref{eq:ct1} in the proof of Theorem \ref{thm:wave}.
-This shows that the variance of our estimator is the same as the variance of the averaging estimator,
-which we already argued is essentially optimal.
-
-%To decompose the error, we need to introduce the following two affine spaces:
-%\begin{align}
-%\Waff&=\affspan \{\wmle_i\}_{i=1}^m
-%,
-%\\
-%\WaffE&=\affspan \{\wmle_i + (\wave - \E\wave)\}_{i=1}^m
-%.
-%\end{align}
-%We now have that
-%\begin{align}
-%\ltwo{\wstar-\proj\W\wstar}
-%&\le
-%\ltwo{\wstar-\proj\Waff\wstar}
-%\\
-%&\le
-%\ltwo{\wstar-\proj\WaffE\wstar}
-%%\\&
-%+
-%\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
-%%\\&+
-%%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
-%\end{align}
-%The first line above follows because $\Waff\subseteq\W$,
-%and the second by the triangle inequality.
-%The left term of the inequality above can be thought of as akin to the bias of the ``estimator'' $\proj\Waff\wstar$,
-%and the right term can be thought of as akin to the variance.
-
-%We begin by bounding the right term.
-%We have that with probability at least $1-\exp(-t)$,
-%\begin{align}
-%&\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
-%\\
-%&=
-%\ltwo{(\proj\Waff\wstar+\wave-\wtave)-\proj\Waff\wstar}
-%\\&=
-%\ltwo{\wave-\wtave}
-%\\&\le
-%\sqrt\frac{v_t}{mn}
-%\end{align}
-%The first equality follows because the space $\WaffE$ is a translation of $\Waff$ by the vector $(\wave-\wtave)$.
-%The probabilistic inequality is due to Equation \ref{eq:ct1} in the proof of Theorem \ref{thm:wave}.
-
-We now bound the left term.
-Define the vector space
-\begin{equation}
-\V=\vecspan \{\wmle_i - \wtave\}_{i=1}^m
-\end{equation}
-which is the translation of the affine space $\WaffE$ to include the origin.
-Then we have that
-%Note that $\proj\Wtave$ is not a linear projection but an affine projection.
-%We first apply a translation so that we are working with the linear projection $\proj\W$.
-\begin{align}
-\ltwo{\wstar-\proj\WaffE\wstar}
-&=
-\ltwo{(\wstar-\wtave)-\proj\V(\wstar-\wtave)}
-\\&=
-\ltwo{(I-\proj\V)(\wstar-\wtave)}
-\end{align}
-In the first step, we translate the vectors so that the affine projection onto $\Wtave$ becomes a linear projection onto $\W$.
-Define $\matV$ to be the $d \times m$ matrix whose $i$th column is the vector $(\wmle_i - \E\wave)$.
-Then $\proj{V} = \matV\matV^+$, where the superscript $^+$ indicates the Moore-Penrose pseudoinverse.
-Each entry of $\proj{V}$ is distributed as the sum of products of Gaussian distributions,
-and is therefore subGaussian.
-}
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Our next lemma bounds the error of $\wowa$ in terms of the error of other estimators whose error we know.
+%Our next lemma bounds the error of $\wowa$ in terms of the error of other estimators whose error we know.
+Our next lemma shows that $\proj{\W}\wstar \approx \wowa$.
 %We introduce the following notation.
@@
 \begin{lemma}
-%Assume that $\wmle_i$ satisfies Condition 1.
-%Let $\delta$ be defined as in Equation \ref{eq:delta} from Lemma \ref{lem:proj}.
-%Then with probability at least $1-\delta$,
-%\begin{equation}
-%\ltwo{\wstar-\wowa}
-%\le
-%\ltwo{\wowastar-\wowa}
-%+
-%\qlo^{-1} \qhi (
-%)
-%.
-%\end{equation}
 Let $\qhi,\qlo : \mathbb{R}^+ \to \mathbb{R}^+$ be monotonically increasing functions such that for all points $\w\in\Theta$,
@@
 \begin{equation}
-\ltwo {\wstar-\wowastar}
-\le
+\begin{aligned}
+\ltwo {\wstar-\wowa}
+&\le
+\ltwo{\wowastar-\wowa}
+\\
+&~~~+
 \qlo^{-1} \left(
@@
 .
+\label{eq:lemma2res}
+\end{aligned}
 \end{equation}
@@
 
-\begin{proof}
-By the triangle inequality, we have that
-\begin{align}
-\ltwo{\wstar-\wowa}
-\le
-\ltwo{\wstar-\wowastar} + \ltwo{\wowastar-\wowa}
-.
-\end{align}
-We bound the left hand term as follows.
-By the definition of $\qhi$ and $\qlo$, we have
-\begin{align}
-\qlo \left( \ltwo {\wstar-\wowastar} \right)
-&\le F(\wstar) - F(\wowastar)
-\\
-&\le F(\wstar) - F(\proj\W\wstar)
-\\
-&\le \qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
-.
-\end{align}
-%and so
-%\begin{equation}
-%\ltwo {\wstar-\wowastar}
-%\le
-%\qlo^{-1} \left(
-    %\qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
-%\right)
-%.
-%\end{equation}
-%Applying Lemma \ref{lem:proj} gives the stated result.
-Applying $\qlo^{-1}$ to both sides gives the stated result.
-\end{proof}
+The proof of Lemma 2 is a straightforward application of the triangle inequality.
+It is contained in Appendix B of the supplement.
+We are now ready to state a formal version of the informal Theorem 2.
 
@@
 
-We are now ready to state a formal version of the informal Theorem \ref{thm:informal}.
-
-\newtheorem*{theoremformal}{Theorem 2 (Formal)}
-\begin{theoremformal}
-Assume that both $\wmle_i$ and $\alpha$ satisfy the SGT Condition.
-Let $t>0$ and
-\begin{equation}
-\delta = (1-\exp((d-m)(-t+\ln(t+1))))(1-\exp(-\sigma^2t^2/2))
-.
-\end{equation}
-Then, with probability at least $\delta$,
-\begin{equation}
-\begin{aligned}
-%\ltwo{\wstar-\wowa}
-%&\le
-%{\wstar-\E\w}
-\end{aligned}
-\end{equation}
-\end{theoremformal}
-
-\clearpage
-
-\newpage
-\ignore{
-\section{ANALYSIS}
-
-First we justify why only a small number of data points is needed for the second optimization.
-For simplicity, we'll restrict our discussion to the case of binary classification.
-This lets us use standard VC-dimension arguments.
-
-\subsection{Few datapoints are needed in the optimization}
-
-%Recall the following theorem from standard VC theory.
-%\begin{theorem}
-%Let $Z$ be a dataset of $n$ points sampled i.i.d. from distribution $\D$.
-%Let $\wmle$ be the parameters of the maximum likelihood estimator of $Z$.
-%Let $L : \Theta \to \mathbb{R}$ be the true 0-1 loss on the distribution $\D$.
-%Let $d$ be the VC dimension of the hypothesis class $\Theta$.
-%Then with probability at least $1-\delta$,
-%\begin{equation}
-%L(\wmle) \le L(\wstar) + O\left(\sqrt{\frac{d + \log(1/\delta)}{n}}\right)
-%\end{equation}
-%\end{theorem}
-%For an extended discussion of the above theorem,
-%see for example Chapter 6 of \cite{shalev2014understanding}.
-\begin{theorem}
-%Let $\Y=\{0,1\}$.
-Define $\alpha^{*}$ to be the optimal parameter constrained to the space spanned by the columns of $W$.
-In notation,
-\begin{equation}
-\alpha^{*} = \argmax_{\alpha} \int_{\Y\times\X} f(y,\trans\x W\alpha) d(y,\x)
-\end{equation}
-Define $L : \mathbb{R} \to \mathbb{R}$ to be the 0-1 loss over the true data distribution.
-Then with probability at least $1-\delta$,
-\begin{equation}
-L(\alpha^*) \le L(\alpha) + O\left(\sqrt{\frac{m + \log(1/\delta)}{n^{reopt}}}\right)
-\end{equation}
-\end{theorem}
-\begin{proof}
-For each machine $i$, fix the set of data points $Z_i$
-(which in turn fixes the $\wmle_i$).
-Define $\A$ to be the space of parameters
-\begin{equation}
-\A=\{\trans W\w : \w \in \Theta\}
-\end{equation}
-The linear hypothesis class defined by $\A$ has VC dimension $m$.
-Since each of the data points in $Z^{reopt}$ is independent of $W$,
-the result follows immediately from standard VC theory.
-See for example Chapter 6 of \cite{shalev2014understanding}.
-\end{proof}
-
-\begin{theorem}
-
-\end{theorem}
-
-\ignore{
-\subsection{Overall growth}
-
-In this section we assume the following regularity conditions.
-
-\begin{theorem}
-Assume conditions (1-5) above hold.
-%Let $m$ be arbitrary.
-Then in the limit as $n$ approaches $\infty$:
-\begin{enumerate}
-\item
-If $f$ is locally $\sigma$-strongly convex around $\wstar$
-\begin{equation}
-\Pr{a}
-\end{equation}
-\item
-If $f$ is locally $\ell$-Lipschitz around $\wstar$
-(i.e. ),
-then
-\begin{equation}
-\end{equation}
-\end{enumerate}
-\end{theorem}
-
-\begin{lemma}
-\label{lem:normaff}
-Let $W : \mathbb{R}^{d\times m}$ be a random matrix where each entry is distributed as a standard Gaussian.
-Let $\Wave$ be the affine hull of the columns of $W$;
-that is,
-\begin{equation}
-\Wave = \left\{\sum_{i=1}^m c_i\wmle_i : c\in\mathbb{R}, \sum_{i=1}^m c_i = 1 \right\}
-\end{equation}
-Then for all $t$,
-\begin{equation}
-\Pr\left[\ltwo{\proj{W}0}^2 > O(t(1-m/d))\right] \le 1 - \exp(-t)
-\end{equation}
-\end{lemma}
-}
-}
-
-\begin{figure*}[ht!]
+\begin{figure*}[t]
 %\begin{tikzpicture}
@@
 %\node at (0,-2) {};
-\begin{tabular}{ccc}
-  \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-autoreg-laplace-log-1000-100-20}
-& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-laplace-log-1000-100-100}
-& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-laplace-log-1000-1000-100}
-\\
-  \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-autoreg-spike-log-1000-100-20}
-& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-spike-log-1000-100-100}
-& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-spike-log-1000-1000-100}
+%\begin{tabular}{ccc}
+  %\includegraphics[width=0.3\textwidth]{img/logl2-star-logl2-auto-normal-log-1000-100-20}
+%& \includegraphics[width=0.3\textwidth]{img/logl2-star-logl2-auto-normal-log-1000-100-100}
+%& \includegraphics[width=0.3\textwidth]{img/logl2-star-logl2-auto-normal-log-1000-1000-100}
+%\\
+  %\includegraphics[width=0.3\textwidth]{img/logl2-star-logl2-auto-uniform-log-1000-100-20}
+%& \includegraphics[width=0.3\textwidth]{img/logl2-star-logl2-auto-uniform-log-1000-100-100}
+%& \includegraphics[width=0.3\textwidth]{img/logl2-star-logl2-auto-uniform-log-1000-1000-100}
+%\\
+  %\includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-autoreg-laplace-log-1000-100-20}
+%& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-laplace-log-1000-100-100}
+%& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-laplace-log-1000-1000-100}
+%\\
+  %\includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-autoreg-spike-log-1000-100-20}
+%& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-spike-log-1000-100-100}
+%& \includegraphics[width=0.3\textwidth]{img/logl1-star-logl2-auto-spike-log-1000-1000-100}
+%\end{tabular}
+%\end{tikzpicture}
+\plots{
+\newcommand{\mklambdaplot}[3]{
+\begin{tikzpicture}
+    [ yscale=0.8
+    ]
+\tiny
+#3
+\begin{axis}
+    [ width=0.35\textwidth
+    , xmode=log
+    , ymode=log
+    , xmin=10^-4
+    , xmax=10^4
+    , xtick={0.0001,0.01,1,100,10000}
+    , enlarge y limits=0
+#2
+    ]
+\addplot[black,no marks] table [x index=3,y index=5] {#1};
+\addplot[brown,no marks] table [x index=3,y index=7] {#1};
+\addplot[blue,no marks] table [x index=3,y index=9] {#1};
+\addplot[red,thick,no marks] table [x index=3,y index=17] {#1};
+\addplot[darkgreen,very thick,no marks] table [x index=3,y index=55] {#1};
+\addplot[dotted,darkgreen,very thick,no marks] table [x index=3,y index=56] {#1};
+\addplot[darkgreen,thin,no marks] table [x index=3,y index=56] {#1};
+%\addplot[thick,red,no marks] table [x=n,y=bootll] {dat/kdd-scaling.dat};
+%\addplot[very thick,darkgreen,no marks] table [x=n,y=owall] {dat/kdd-scaling.dat};
+\end{axis}
+%\node at (2.2,-0.8) {\small regularization strength ($\lambda$)};
+\end{tikzpicture}
+}
+\begin{tabular}{cccc}
+& $d=100,m=20$
+& $d=100,m=100$
+& $d=1000,m=100$
+\\
+{\small\rotatebox{90}{\hspace{0.05cm}squared error $\ltwo{\wstar-\w}^2$}}
+&\hspace{-0.5cm}\mklambdaplot
+    {dat/logl1-star-logl2-autoreg-spike-log-1000-100-20.pdf.csv}
+    {,ymin=10^-2,ymax=10^2,ytick={0.01,0.1,1,10,100}}{
+    \node at (2,2.4) {\textcolor{brown}{$\wmle_i$}};
+    \draw[->,brown] (2,2.3) -- (2.1,2);
+    \node at (3.5,1.95) {\textcolor{blue}{$\wave$}};
+    \draw[->,blue] (3.5,1.9) -- (3.55,1.75);
+    \node at (4,1.1) {\textcolor{red}{$\wboot$}};
+    \draw[->,thick,red] (3.8,1.3) -- (3.75,1.45);
+    \node at (0.5,1.85) {$\wmle$};
+    \draw[->] (0.5,2.0) -- (0.6,2.2);
+    \node at (0.7,0.9) {\textcolor{darkgreen}{$\wowa$}};
+    \draw[->,thick,darkgreen] (0.9,0.8) -- (1.2,0.8);
+    \node at (0.6,0.4) {\textcolor{darkgreen}{$\wowafull$}};
+    \draw[->,thick,dotted,darkgreen] (0.9,0.3) -- (1.2,0.3);
+    \draw[->,very thin,darkgreen] (0.9,0.3) -- (1.2,0.3);
+    }
+&\hspace{-0.5cm}\mklambdaplot
+    {dat/logl1-star-logl2-auto-spike-log-1000-100-100.pdf.csv}
+    {,ymin=10^-3,ymax=10^2,ytick={0.001,0.01,0.1,1,10,100}}
+    {}
+&\hspace{-0.5cm}\mklambdaplot
+    {dat/logl1-star-logl2-auto-spike-log-1000-1000-100.pdf.csv}
+    {,ymin=10^-1,ymax=10^3,ytick={0.1,1,10,100,1000}}
+    {}
+\\
+& \hspace{0.2cm} {\small regularization strength ($\lambda$)}
+&
+&
 \end{tabular}
-%\end{tikzpicture}
-\caption{test}
+}
+\caption{
+    OWA is robust to the regularization strength.
+    Surprisingly, additional regularization introduced by OWA lets it outperform the oracle estimator $\wmle$ in some cases.
+    Our theory states that as $m\to d$, $\wowa\to\wmle$.
+    This is confirmed in the middle experiment.
+    In the leftmost experiment, $m<d$, but $\wowa$ still behaves similarly to $\wmle$.
+    In the rightmost experiment, $\wowa$ has similar performance as $\wave$ and $\wboot$ but is less sensitive to $\lambda$.
+    }
 \label{fig:lambda}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newtheorem*{theoremformal}{Theorem 2 (Formal)}
+\begin{theoremformal}
+Assume that $\wmle_i$ and $\alpha$ satisfy the SGT condition.
+Let $t>0$ and
+\begin{equation}
+%\delta = 1-\exp((d-m)(-t+\ln(t+1)))
+\delta = (1-\exp((d-m)(-t+\ln(t+1))))(1-\exp(-t))
+.
+\end{equation}
+Then, with probability at least $\delta$,
+\begin{equation}
+\begin{aligned}
+&\ltwo{\wstar-\wowa}
+\\
+&\le
+%\ltwo{\wowastar-\wowa}
+\ltwo{\wowastar-\E\wowa}
++ \sqrt{\frac{v_t}{m\nowa}}
+\\
+&+
+\qlo^{-1}\bigg(\qhi\bigg((t+1)\sqrt{\frac{\smin}{\smax}\left(1-\frac{m}{d}\right)}\ltwo{\wstar-\wave}\bigg)\bigg)
+\end{aligned}
+\end{equation}
+\end{theoremformal}
+
+\begin{proof}
+Bound the leftmost term of Equation \ref{eq:lemma2res} using the same procedure we used for Theorem 1.
+Then use Lemma 1 to bound the rightmost term.
+%Define
+%\begin{equation}
+%\astar = \argmin_{\w\in\W} \int_{\X\times\Y} \left(f(y,\trans\x\w) + \lambda R(\w)\right) d(\x,y)
+%\end{equation}
+%The result follows by substituting Theorem 1 and Lemma 1 into Lemma 2.
+\end{proof}
+
+\newpage
+
+%There are two differences between the formal and informal version of Theorem 2 worth highlighting.
+Notice that the formal version of Theorem 2 contains the term $\ltwo{\wowastar-\E\wowa}$, which doesn't appear in the informal version.
+This is the bias of the second round of optimization in the subspace $\Wowa$ due to $\nowa$ being finite.
+This bias will be less than the bias of the full problem because $\Wowa\subset\Theta$.
+As $m\to\infty$, $\wowastar\to\wstar$ and $\E\wowa\to\W\wmle$, so $\ltwo{\wowastar-\E\wowa}\to\ltwo{\wstar-\E\wmle}$.
+%Furthermore, this bias will reduce at the rate of $O(1/\sqrt{m\nowa})$ under standard assumptions.
+%The next difference is the application of the functions $\qlo$ and $\qhi$.
+
 \section{EXPERIMENTS}
@@
 
-We test our distributed estimator on two logistic regression tasks.
+We test OWA on two logistic regression tasks.
 The first task uses synthetic data.
 The second task uses real world add click data from the Tencent search engine.
-
 In each experiment, we compare our $\wowa$ estimator with four baseline estimators:
@@
 The $\wboot$ estimator has a parameter $r$ that needs to be tuned.
-We make an effort to present the $\wboot$ estimator in the best light possible.
-In all experiments we evaluate $\wboot$ with $r \in \{0.005,0.01,0.02,0.04,0.1,0.2,0.3\}$,
+%We make an effort to present the $\wboot$ estimator in the best light possible.
+In all experiments we evaluate $\wboot$ with $r \in \{0.005,0.01,0.02,0.04,0.1,0.2\}$,
 which is a set recommended in the original paper \citep{zhang2012communication},
@@
 %\begin{tikzpicture}
-\begin{tabular}{ccc}
+%\begin{tabular}{ccc}
   %\includegraphics[width=0.3\textwidth]{img/logl2-auto-logl2-auto-normal-log-1000-100-star.pdf}
@@
 %\\
-  \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-100-star.pdf}
-& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-1000-star.pdf}
-& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-10000-star.pdf}
-\\
-  \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-spike-log-1000-100-star.pdf}
-& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-spike-log-1000-1000-star.pdf}
-& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-spike-log-1000-10000-star.pdf}
+  %\includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-100-star.pdf}
+%& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-1000-star.pdf}
+%& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-laplace-log-1000-10000-star.pdf}
+%\\
+  %\includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-spike-log-1000-100-star.pdf}
+%& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-spike-log-1000-1000-star.pdf}
+%& \includegraphics[width=0.3\textwidth]{img/logl1-auto-logl2-auto-spike-log-1000-10000-star.pdf}
+%\end{tabular}
+%\end{tikzpicture}
+\plots{
+\newcommand{\mklambdaplot}[4]{
+\begin{tikzpicture}
+    [ yscale=0.8
+    ]
+\tiny
+#3
+\begin{axis}
+    [ width=0.35\textwidth
+    %, xmode=log
+    , ymode=log
+    , xmin=1
+    , xmax=100
+#2
+    ]
+\addplot[black,no marks] table [x index=0,y index=5] {#1};
+\addplot[brown,no marks] table [x index=0,y index=7] {#1};
+\addplot[blue,no marks] table [x index=0,y index=9] {#1};
+\addplot[red,thick,no marks] table [x index=0,y index=#4] {#1};
+\addplot[darkgreen,very thick,no marks] table [x index=0,y index=55] {#1};
+\addplot[dotted,darkgreen,very thick,no marks] table [x index=0,y index=56] {#1};
+\addplot[darkgreen,very thin,no marks] table [x index=0,y index=56] {#1};
+%\addplot[thick,red,no marks] table [x=n,y=bootll] {dat/kdd-scaling.dat};
+%\addplot[very thick,darkgreen,no marks] table [x=n,y=owall] {dat/kdd-scaling.dat};
+\end{axis}
+\end{tikzpicture}
+}
+\begin{tabular}{cccc}
+& $d=100$
+& $d=1000$
+& $d=10000$
+\\
+{\small\rotatebox{90}{\hspace{0.05cm}squared error $\ltwo{\wstar-\w}^2$}}
+&\hspace{-0.5cm}\mklambdaplot
+    {dat/logl1-auto-logl2-auto-spike-log-1000-100-star.pdf.csv}
+    {,ymin=10^-2,ymax=10^1}
+    { \node at (3,3) {\tiny\textcolor{brown}{$\wmle_i$}};
+      \draw[->,brown] (2.7,2.85) -- (2.6,2.7);
+      \node at (3.5,2.5) {\tiny\textcolor{blue}{$\wave$}};
+      \draw[->,blue] (3.4,2.45) -- (3.45,2.35);
+      \node at (2.5,1.9) {\tiny\textcolor{red}{$\wboot$}};
+      \draw[->,thick,red] (2.5,2.0) -- (2.6,2.2);
+      \node at (3.3,1.4) {\tiny$\wmle$};
+      \draw[->] (3.2,1.3) -- (3.3,1.15);
+      \node at (0.8,0.5) {\tiny\textcolor{darkgreen}{$\wowafull$}};
+      \draw[->,darkgreen,dotted,thick] (0.8,0.65) -- (0.9,0.8);
+      \draw[->,darkgreen,very thin] (0.8,0.65) -- (0.9,0.8);
+      \node at (2.7,0.8) {\tiny\textcolor{darkgreen}{$\wowa$}};
+      \draw[->,darkgreen,thick] (2.6,0.7) -- (2.7,0.55);
+    }
+    {19}
+&\hspace{-0.5cm}\mklambdaplot
+    {dat/logl1-autoreg-logl2-auto-spike-log-1000-1000-star.pdf.csv}
+    {,ymin=10^0,ymax=10^2}
+    {}
+    {15}
+&\hspace{-0.5cm}\mklambdaplot
+    {dat/logl1-auto-logl2-auto-spike-log-1000-10000-star.pdf.csv}
+    {}
+    {}
+    {11}
+\\
+& \hspace{0.2cm} {\small number of machines ($m$)}
+&
+&
 \end{tabular}
-%\end{tikzpicture}
-\caption{}
+}
+\vspace{-0.15in}
+\caption{
+    $\wowa$ scales well with the number of machines.
+    Surprisingly, it outperforms the oracle estimator trained on all of the data $\wmle$ in some situations.
+    This is likely due to the additional regularization introduced by the OWA algorithm,
+    as seen in Figure \ref{fig:lambda}.
+    }
 \label{fig:synscale}
@@
 \end{equation}
-In all experiments, we use the $L1$ regularizer to induce sparsity in our estimates of $\wstar$.
+In all experiments, we use the L1 regularizer to induce sparsity in our estimates of $\wstar$.
 Our estimators will be biased because the model is misspecified.
-$L1$ regularization actually corresponds to a Laplace prior on the parameter vector.
-Results are qualitatively similar when using a uniform, normal, and Laplace prior on $\wstar$.
-These results are not shown due to space limitations.
+%(L1 regularization corresponds to a Laplace prior on the parameter vector.)
+The true model for L1 regularization has a Laplace prior on $\wstar$.
+%Results are qualitatively similar when using a uniform, normal, and Laplace prior on $\wstar$.
+%These results are not shown due to space limitations.
 
@@
 In this experiment, $\lambda$ was allowed to vary from $10^{-4}$ to $10^4$.
-For each value of $\lambda$, we randomly generated 50 datasets and calculated the corresponding estimators.
-Figure \ref{fig:lambda} shows the average of the results.
-Our $\wowa$ estimator exhibits significantly less sensitivity to the selection of $\lambda$.
+For each value of $\lambda$, we randomly generated 50 datasets (with $n=1000$) and then calculated the corresponding estimators.
+Figure \ref{fig:lambda} shows the average of the results for three choices of $m$ and $d$.
+Our $\wowa$ estimator is significantly less sensitive to the choice of $\lambda$ than the other distributed estimators.
+Surprisingly, $\wowa$ even outperforms the oracle $\wmle$ in some regimes.
+This is likely due to additional regularization induced by the approximation in Equation \ref{eq:approxreg}.
 
 Our second experiment shows how the parallel algorithms scale as the number of machines $m$ increases.
-We fix $n=1000$ data points per machine, so the size of the dataset $mn$ grows as we add more machines.
-This simulates the typical ``big data'' scenario where data is abundant,
+We fix $n=1000$ data points per machine,
+so the size of the dataset $mn$ grows as we add more machines.
+This simulates the typical ``big data'' regime where data is abundant,
 but processing resources are scarce.
-%We further consider three regimes:
-%more data than dimensions ($d=100$),
-%one data point per dimension ($d=1000$),
-%and less data than dimensions ($d=10000$).
 Each machine independently uses cross validation to select the $\lambda$ that best fits the data locally.
-There are three advantages to this method of parameter selection.
-First, there is no additional communication overhead for model selection because model selection is a completely local task.
+There are three advantages to this model selection procedure.
+First, there is no additional communication because model selection is a completely local task.
 Second, existing optimizers have built-in model selection routines which make the process easy to implement.
-In our experiments, we use the default model selection procedure from Python's SciKit-Learn \citep{scikit-learn}.
-Third, the data may be best fit using different regularization strengths for each machine,
-and this method allows that to happen.
+We used the default model selection procedure from Python's SciKit-Learn \citep{scikit-learn}.
+Third, the data may be best fit using different regularization strengths for each machine.
 The results are shown in Figure \ref{fig:synscale}.
-The performance of our method scales well with the number of machines,
-and in some cases even outperforms the oracle on all of the data.
+%The performance of our method scales well with the number of machines,
+%and in some cases even outperforms the oracle trained on all of the data.
+The performance of $\wowa$ scales much better than $\wave$ and $\wboot$.
+Because of the regularization induced by Equation \ref{eq:approxreg} observed in the previous experiment,
+$\wowa$ even scales better than $\wmle$ in some regimes.
+%Surprisingly, matches (or exceeds) the performance of the oracle $\wmle$ as the number of machines $m$ increases.
 
@@
 
-In this section,
-we evaluate our estimator on real world data from the KDD 2012 Cup \citep{kddcup2012}.
+We now evaluate our estimator on real world data from the KDD 2012 Cup \citep{kddcup2012}.
 The goal is to predict whether a user will click on an add from the Tencent internet search engine.
 This dataset was previously used to evaluate the performance of $\wboot$ \citep{zhang2012communication}.
-
 This dataset is too large to fit on a single machine.
@@
 each of dimension 741,725.
+The data points are sparse, so we use the L1 norm to encourage sparsity in our final solution.
+The regularization strength was set using cross validation in the same manner as for the synthetic data.
 For each test, we split the data into 80 percent training data and 20 percent test data.
@@
 
+Our first experiment tests the sensitivity of the $\nowa$ parameter on large datasets.
+We fix $m=128$, and allow $\nowa$ to vary from $2^0$ to $2^{20}$,
+which is approximately the size of the full dataset.
+We repeated the experiment 50 times, each time using a different randomly selected set $\Zowa$ for the second optimization.
+Figure \ref{fig:nowa} shows the results.
+Our $\wowa$ estimator has lower loss the $\wave$ using only 16 data points per machine (approximately $4\times10^{-8}$ percent of the full training set)
+and $\wowa$ has converged to its final loss value with only 1024 data points per machine (approximately $2.7\times10^{-6}$ percent of the full training set).
+This justifies our claim that only a small number of data points are needed for the second round of optimization,
+and so the communication complexity of $\wowa$ is essentially the same as $\wave$.
+
+\begin{figure}[h!]
+\plots{
+\begin{tabular}{cc}
+\rotatebox{90}{\hspace{1cm}log-loss}
+&
+\hspace{-0.5cm}
+\begin{tikzpicture}
+\small
+\node at (2.5,1.5) {\tiny\textcolor{darkgreen}{$\wowa$}};
+\draw[->,darkgreen,thick] (2.1,1.5) -- (1.85,1.5);
+\node at (1,0.2) {\tiny\textcolor{red}{$\wboot$}};
+\draw[->,red,thick] (0.6,0.25) -- (0.5,0.45);
+\node at (1,1.1) {\tiny\textcolor{blue}{$\wave$}};
+\draw[->,blue] (0.9,1) -- (1.0,0.85);
+\begin{axis}
+    [ width=0.45\textwidth
+    , height=1.5in
+    , xmin=1
+    , xmax=1048576
+    , ymin = 0.137
+    , ymax = 0.14
+    , y tick label style={
+        /pgf/number format/.cd,
+            fixed,
+            fixed zerofill,
+            precision=3,
+        /tikz/.cd
+    },
+    , xtick={1,32,1024,32768,1048576}
+    , log basis x={2}
+    , xmode=log
+    ]
+%\addplot[blue,no marks] coordinates {(1,0.1376699823) (1048576,0.1376699823)};
+%\addplot[thick,red,no marks] coordinates {(1,0.137198363008) (1048576,0.137198363008)};
+%\addplot[very thick,darkgreen,no marks] table [x=nowa,y=128ll] {dat/kdd-nowa.dat};
+\addplot[blue,no marks] coordinates {(1,0.138045477888) (1048576,0.138045477888)};
+\addplot[thick,red,no marks] coordinates {(1,0.137682508908) (1048576,0.137682508908)};
+\addplot[very thick,darkgreen,no marks] table [x=nowa,y=16ll] {dat/kdd-nowa.dat};
+\end{axis}
+\end{tikzpicture}
+\\
+&
+\hspace{-0.15cm} data points used in second optimization $(\nowa)$
+\end{tabular}
+}
+\vspace{-0.15in}
+\caption{
+    Relatively few data points are needed in the second round of optimization for $\wowa$ to converge.
+    %Shown is the case when $m=16$.
+    %Other values of $m$ are similar.
+    }
+\label{fig:nowa}
+\end{figure}
+
+\begin{figure}[h!]
+\plots{
+\begin{tabular}{cc}
+\rotatebox{90}{\hspace{1cm}log-loss}
+&
+\hspace{-0.25cm}
+\begin{tikzpicture}
+\node at (5.3,0.75) {\tiny\textcolor{blue}{$\wave$}};
+\draw[->,blue] (5.1,0.6) -- (5,0.4);
+\node at (3.5,0.85) {\tiny\textcolor{red}{$\wboot$}};
+\draw[->,red,thick] (3.2,0.7) -- (3,0.35);
+\node at (0.5,0.35) {\tiny\textcolor{darkgreen}{$\wowa$}};
+\draw[->,darkgreen, very thick] (0.5,0.5) -- (0.7,0.9);
+\small
+\begin{axis}
+    [ width=0.45\textwidth
+    , height=1.5in
+    , xmin=2
+    , xmax=128
+    , ymin = 0.137
+    , ymax = 0.142
+    , ytick={0.137,0.138,0.139,0.14,0.141,0.142}
+    , y tick label style={
+        /pgf/number format/.cd,
+            fixed,
+            fixed zerofill,
+            precision=3,
+        /tikz/.cd
+    },
+    %, xtick={2,4,128}
+    , log basis x={2}
+    , xmode=log
+    ]
+\addplot[blue,no marks] table [x=n,y=avell] {dat/kdd-scaling.dat};
+\addplot[thick,red,no marks] table [x=n,y=bootll] {dat/kdd-scaling.dat};
+\addplot[very thick,darkgreen,no marks] table [x=n,y=owall] {dat/kdd-scaling.dat};
+\end{axis}
+\end{tikzpicture}
+\\
+&
+\hspace{0.5cm}number of machines ($m$)
+\end{tabular}
+}
+\vspace{-0.15in}
+\caption{
+    Performance of the parallel estimators on advertising data as the number of machines $m$ increases.
+    }
+\label{fig:kdd-scaling}
+\end{figure}
+
+Our last experiment shows the performance as we scale the number of machines $m$.
+The results are shown in Figure \ref{fig:kdd-scaling}.
+Here, our $\wowa$ performs especially well in the low $m$ setting.
+For large $m$, $\wowa$ continues to slightly outperform $\wboot$ without the need for an expensive model selection procedure to determine the $r$ parameter.
+
 \section{CONCLUSION}
+
+\vspace{-0.1in}
+We introduced a new distributed estimation algorithm called OWA.
+OWA has the speed advantages of non-interactive distributed estimators,
+but has better accuracy due to a (cheap) second round of optimization.
+Unlike other algorithms, OWA does not require cross validation to tune any hyperparameters.
+Furthermore, our analysis is more general than the analysis of similar algorithms.
+As part of our analysis, we also provided a more general analysis of the averaging estimator.
+
+%We do not require the likelihood to be convex.
+%OWA uses two rounds of communication,
+%so it is not subject to existing regret bounds for non-interactive learners.
+%The second round of communication requires little communication, however,
+%so we maintain the advantages of non-interactive distributed learning.
 
@@
 
+\clearpage
+
+\section*{Appendix A: Proof of Lemma 1}
+
+Define the space $\Gamma=\vecspan\{\wmle_i-\E\wave\}_{i=1}^m$.
+Our proof strategy is to bound the distance $\ltwo{\wstar-\proj\Gamma\wstar}$ and then show that $\Gamma\approx\W$.
+Specifically, we have
+\begin{align}
+~~~~~~&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!%\!\!\!\!\!
+\ltwo{\wstar-\proj\W\wstar}
+\nonumber
+\\
+&=
+\ltwo{(\wstar-\wave) - (\proj\W\wstar - \wave)}
+%\nonumber
+\\
+&=
+\ltwo{(\wstar-\wave) - \proj\W(\wstar - \wave)}
+\label{eq:waveinW}
+%\nonumber
+\\
+&\le
+\ltwo{(\wstar-\wave) - \proj\Gamma(\wstar-\wave)}
+\nonumber
+\\
+&~~~+
+\ltwo{\proj\Gamma(\wstar-\wave) - \proj\W(\wstar-\wave)}
+\label{eq:gammatriangle}
+\\
+&=
+\ltwo{(I-\proj\Gamma)(\wstar - \wave)}
+\nonumber
+\\
+&~~~+
+\ltwo{(\proj\Gamma-\proj\W)(\wstar - \wave)}
+.
+\label{eq:i-projW}
+\end{align}
+Equation \ref{eq:waveinW} follows because $\wave\in\W$,
+and Equation \ref{eq:gammatriangle} is due to the triangle inequality.
+%Recall that we defined $\matW$ to be the $d\times m$ matrix with $i$th column equal $\wmle_i$.
+%The column space of $\matW$ is $\W$.
+We now bound each of the terms in Equation \ref{eq:i-projW} separately,
+beginning with the leftmost.
+
+%We can write
+%\begin{equation}
+%\proj\Gamma = \Delta_{\wmle_i}
+%\end{equation}
+Define the matrix $G$ with $i$th column equal to the vector $(\wmle_i-\E\wmle_i)$.
+The column space of $G$ is $\Gamma$,
+and we have that
+%$\proj\Gamma=G\pinv G$.
+\begin{align}
+\proj\Gamma
+&=
+G\pinv G
+\\
+&= \I_\wstar^{-1/2}\bigg(\sqrt{n}\I_\wstar^{1/2}G\bigg)
+             \pinv{\bigg(\sqrt{n}\I_\wstar^{1/2}G\bigg)}
+   \I_\wstar^{1/2}
+,
+\label{eq:ggpinv}
+\end{align}
+where the superscript $\pinv{}$ denotes the Moore-Penrose pseudoinverse.
+The $i$th column of matrix $\sqrt{n}\I^{1/2}_{\wstar}G$ is equal to $\Delta_{\wmle_i}$ by definition.
+%
+%So the projection $\proj\W = \matW\pinv\matW$, where the superscript $\pinv{}$ is the Moore-Penrose pseudoinverse.
+%Define the matrix $\matV=\sqrt{n}I^{1/2}\matW$.
+%Then the columns of $\matV$ are sub-Gaussian by the SGT condition,
+%and we can write
+%\begin{equation}
+%\proj\W
+%= \I^{-1/2} \matV \pinv\matV \I^{1/2}
+%.
+%\label{eq:vvp}
+%\end{equation}
+Let $U\Sigma \trans V$ be the singular value decomposition of $\sqrt{n}\I^{1/2}_{\wstar}G$,
+where $U$ and $V$ are orthogonal matrices and $\Sigma$ is the diagonal matrix of singular values.
+By the rotational invariance of the sub-Gaussian distribution (Lemma 5.9 of Vershynin, 2011),
+$U$ and $V$ are distributed uniformly over the orthogonal group of matrices.
+Substituting into Equation \ref{eq:ggpinv} gives
+\begin{align}
+\proj\Gamma
+&= \I^{-1/2} (U\Sigma \trans V)\pinv{(U \Sigma \trans V)} \I^{1/2}
+\\
+&= \I^{-1/2} (U\Sigma \trans V)(V \pinv\Sigma \trans U) \I^{1/2}
+\\
+&= \I^{-1/2} U\Sigma \pinv\Sigma \trans U \I^{1/2}
+.
+\label{eq:vvp2}
+\end{align}
+%Substituting Equation \ref{eq:vvp2} into \ref{eq:i-projW} gives
+This implies that
+\begin{align}
+%~~~~~&\!\!\!\!\!\!\!\!\!\!\!\ltwo{\wstar-\proj\Gamma\wstar}
+~~~~~&\!\!\!\!\!\!\!\!\!\!\!\ltwo{(I-\proj\Gamma)(\wstar-\wave)}
+\nonumber
+\\
+&=
+\ltwo{(I-\I^{-1/2} U\Sigma \pinv\Sigma \trans U \I^{1/2})(\wstar - \wave)}
+\\
+&=
+\ltwo{\I^{-1/2} U(I-\Sigma \pinv\Sigma )\trans U \I^{1/2}(\wstar - \wave)}
+\\
+&=
+\sqrt{\smin}\ltwo{U(I-\Sigma \pinv\Sigma )\trans U \I^{1/2}(\wstar - \wave)}
+\label{eq:fromdefsmin}
+\\
+&=
+\sqrt{\smin}\ltwo{(I-\Sigma \pinv\Sigma )\trans U \I^{1/2}(\wstar - \wave)}
+\label{eq:fromrotinvl2}
+.
+\end{align}
+Equation \ref{eq:fromdefsmin} follows from the definition of $\smin$ as the smallest eigenvector of $\I$,
+and Equation \ref{eq:fromrotinvl2} follows from the rotational invariance of the L2 norm.
+The matrix $(I-\Sigma\pinv\Sigma)$ contains 1s in the first $d-m$ diagonal entries, and zeros everywhere else.
+Because $U$ is distributed uniformly over the orthogonal group,
+the operator $(I-\Sigma\pinv\Sigma)\trans U$ projects a vector onto a subspace of dimension $d-m$ chosen uniformly at random.
+Lemma 2.2 from \cite{dasgupta2003elementary} provides a bound for this situation.
+The lemma states that for any vector $\x$, for all $t>0$, with probability at least $1-\exp((d-m)(-t+\ln (t+1)))$,
+\begin{align}
+\ltwo{(I-\Sigma\pinv\Sigma)\trans U \x} \le (t+1) \sqrt{1-\frac{m}{d}}
+.
+\label{eq:dasgupta}
+\end{align}
+%Substituting gives
+%Applying Equation \ref{eq:dasgupta} to \ref{eq:blurby} gives
+Setting $\x=\I^{1/2}(\wstar-\wave)$ gives
+\begin{align}
+~~~~~&\!\!\!\!\!\!\!\!\!\!\!\ltwo{(I-\proj\Gamma)(\wstar-\wave)}
+\nonumber
+\\
+&\le
+(t+1)\sqrt{\smin\left(1-\frac{m}{d}\right)}\ltwo{\I^{1/2}(\wstar - \wave)}
+\\
+&\le
+(t+1)\sqrt{\frac{\smin}{\smax}\left(1-\frac{m}{d}\right)}\ltwo{\wstar - \wave}
+,
+\end{align}
+which bounds the leftmost term of Equation \ref{eq:i-projW}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section*{Appendix B: Proof of Lemma 2}
+By the triangle inequality, we have that
+\begin{align}
+\ltwo{\wstar-\wowa}
+\le
+\ltwo{\wowastar-\wowa}
++
+\ltwo{\wstar-\wowastar}
+.
+\label{eq:wstarwowa}
+\end{align}
+We bound the right hand term as follows.
+By the definition of $\qhi$ and $\qlo$, we have
+\begin{align}
+\qlo \left( \ltwo {\wstar-\wowastar} \right)
+&\le F(\wstar) - F(\wowastar)
+\\
+&\le F(\wstar) - F(\proj\W\wstar)
+\\
+&\le \qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
+.
+\end{align}
+%and so
+%\begin{equation}
+%\ltwo {\wstar-\wowastar}
+%\le
+%\qlo^{-1} \left(
+    %\qhi \left( \ltwo {\wstar - \proj\W\wstar} \right)
+%\right)
+%.
+%\end{equation}
+%Applying Lemma \ref{lem:proj} gives the stated result.
+Applying $\qlo^{-1}$ to both sides and substituting into Equation \ref{eq:wstarwowa} gives the stated result.
+
 \end{document}
